---
title: "Roughy 2019 Tables and Figures"
author: "Caroline Sutton"
date: "03/12/2019"
output:
  html_document:
    code_folding: hide
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE,
  warning = FALSE,
  message = FALSE,
  results = 'asis',
  fig.align = 'center')

#Load libraries
library(tidyverse)
library(knitr)
library(kableExtra)


#read in tidy data from results folder
Roughy <- read_csv("Results/OREastern_Tidy_87to2019.csv", col_types = list(ShotWeight_kg = col_double(), mark = col_character()))


#remame and select columns
Roughy <- Roughy %>% 
  mutate(Length =round(SLadj_cm)) %>% #round values for length in cm 2019 was to nearest mm
  mutate(Ground = Area_MS) %>% 
  drop_na(Ground) %>% 
  #drop_na(Length) %>% 
  drop_na(sex) %>% 
  select(Ground, year, Shot, Date,sex,stage, SLadj_cm,Length,ShotWeight_kg, weight_kg, Otoliths)
```

```{r}
  
Roughyminmax <- Roughy %>% 
  dplyr::group_by(Ground,sex) %>% 
  dplyr::summarise(round(mean(Length),digits = 2))

Roughyminmax
```
# Introduction
Summary Data, Tables and Figures for Eastern Zone Roughy Survey 2019


## Table and Figures

---
### summary table

```{r mytable, out.width='100%', echo = T}
Roughy_DataSummary <- Roughy %>% 
  dplyr::group_by(year, Ground,sex) %>% 
  dplyr::summarise(n= n())

Roughy_DataSummary <- spread(Roughy_DataSummary, key = year,value = n)


knitr::kable(tail(Roughy_DataSummary, n = 5),caption = "Summary data: Number of Fish measured for orange roughy eastern zone spawning surveys by, year, Area, sex") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width = "100%")
```
## Table 2.3
```{r table 2.3, echo=FALSE}
# fitler data for 2019
Roughy_year <- Roughy %>% 
  filter(year==2019)

Roughy_year_sum <- Roughy %>% 
  filter(year==2019) %>% 
  dplyr::group_by(Ground,Shot,Date,ShotWeight_kg,sex) %>% 
  dplyr::summarise(num_rows = n()) %>% 
  spread(sex,num_rows) %>% 
  mutate(`Total` = F + M) 

# creating ground/ and survey totals for:

#St Helens
TSH <- Roughy_year_sum %>% 
  filter(Ground == "St Helens") %>% 
  dplyr::group_by(Ground) %>% 
  dplyr::summarise(ShotWeight_kg =sum(ShotWeight_kg), F= sum(F), M = sum(M), Total =sum(Total)) %>% 
  mutate(Ground = "St Helens Total")
#St Patricks
TSP <- Roughy_year_sum %>% 
  filter(Ground == "St Patricks") %>% 
  dplyr::group_by(Ground) %>% 
  dplyr::summarise(ShotWeight_kg =sum(ShotWeight_kg, na.rm = TRUE), F= sum(F), M = sum(M), Total =sum(Total)) %>% 
  mutate(Ground = "St Patricks Total")

# totals for SH and SP
TotalS <- Roughy_year_sum %>% 
  summarise(ShotWeight_kg = sum(ShotWeight_kg, na.rm = TRUE), F= sum(F), M = sum(M), Total =sum(Total)) %>% 
  mutate(Ground = "Survey Total")

#bind the shot and totals 
Roughy_year_sumtotals <- bind_rows(Roughy_year_sum, TSH, TSP,TotalS)  # put all together for totals
 # dplyr::rename("Shot Weight (kg)" = ShotWeight_kg, "Total Nos Fish" = Total)
  
Roughy_year_sumtotals

```
## Summary for otoliths data for updating Table 2.3
```{r}
Roughy_year <- Roughy %>% 
  filter(year == 2019)
Roughy_otoliths <- Roughy_year %>% 
  dplyr::group_by(Ground,Shot, sex) %>% 
  dplyr::summarise(o =sum(Otoliths, na.rm = TRUE)) %>% 
  spread(sex,o) %>% 
  mutate('T_o' = F + M)

Roughy_otoliths <- Roughy_year %>% 
  dplyr::group_by(Ground,sex) %>% 
  dplyr::summarise(o =sum(Otoliths, na.rm = TRUE)) %>% 
  spread(Ground,o) 
Roughy_otoliths



```
### Table 2.4
* need to update code to include shot weights for shots where fish were not measured **

Summary of shot dates and weights for St Helens and St Patricks, Eastern Zone, 2019 (* no fish collected from these shots. Note that bycatch data were not reported

```{r table2.4, out.width='100%', echo = T} 
Roughy_2019 <- Roughy %>% 
  filter(year == 2019)
Roughy_2019_shotsum <- Roughy_2019 %>% 
  dplyr::group_by(Ground,Shot,Date, 'Shot Weight (kg)' = ShotWeight_kg,sex) %>% 
  dplyr::summarise(nos = n()) %>% 
  spread(sex,nos) %>% 
  mutate(`Total fish` = F + M)

write.csv(Roughy_2019_shotsum,file = "Figures/t2.4.csv")

#tab_df(OR2019_sum,file = "Figures/t2.3.doc") # create and save word doc . Must # as once created

knitr::kable(head(Roughy_2019_shotsum, n=14), format = "html", caption = "Table 2.4 Summary of shot dates and weights for St Helens and St Patricks, Eastern Zone, 2019") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width  = "100%")

```


### Figure 2.1 
Average Standard Length (cm) of orange roughy Eastern Zone spawning population, showing males and females at St Helens and St Patricks for 1987 to 2019. Error bars are 95% confidence limits (updated from Kloser et al. 2016)
```{r myfig2.1}
#average SL by Area and sex

library(Rmisc)
library(ggplot2)

roughy_sum <- summarySE(Roughy, measurevar = "Length", groupvars = c("year","Ground","sex")) 


  
pd <- position_dodge(0.6) # move position horizontally b/c of overlap

f2.1 <- ggplot(roughy_sum, aes(x = year,y = Length, shape = Ground, colour = sex)) +
  geom_errorbar(aes(ymin = Length - ci, ymax= Length + ci),
                width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_line(position = pd) +
  xlab("Eastern Zone Spawning Population by year for July")+
  ylab("Average Standard Length (cm)")+
  scale_y_continuous(breaks = seq(26,38,1),limits = c(26,38))+
  scale_x_continuous(limits = c(1986,2020), breaks = seq(1986,2020,4))+
  theme_bw()+
  theme(legend.position = c(.95, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'),
         plot.background = element_rect(linetype =  "solid") )

  
plot(f2.1)
ggsave("Figures/f2.1.emf")
ggsave("Figures/f2.1.jpg")
```

## length data table
```{r}
Roughy_2019_length_GroundbySex <- summarySE(Roughy_2019, measurevar = "Length", 
                                      groupvars = c("Ground","sex"))
Roughy_2019_length_GroundbySex

Roughy_2019_length_Ground <- summarySE(Roughy_2019, measurevar = "Length", 
                                      groupvars = c("Ground"))
Roughy_2019_length_Ground
```
## Figure 2.2
Standard Length by sex and ground
```{r myfig2.2}
OR_LF <- Roughy_year %>% # This is the ONE
  group_by(Ground, sex) %>% 
  add_tally() %>% 
  group_by(Ground, sex, Length) %>% 
  dplyr::summarise(prop = n() / n[1])

f2.2 <- ggplot(OR_LF, aes(Length, prop, fill = sex)) +
  geom_col(position = "identity", alpha = 0.6) +
  scale_x_continuous(breaks = seq(28,46,2),limits = c(28,46))+
  facet_wrap(~Ground)+
   theme_bw()+
  ylab("proportion")+
  xlab("Length (cm)")+
  theme(legend.position = "bottom")
plot(f2.2)
ggsave(filename = "Figures/f2.2.emf")
ggsave(filename = "Figures/f2.2.jpg")
```

## Figure 2.3
Cummunlative Frequency
```{r myfig2.3}

#using stat_ecdf to calculate the cummulative frequency

f2.3 <- ggplot(Roughy_year,aes(x=Length, colour= sex)) +
   stat_ecdf(geom = "smooth")+
  scale_x_continuous(breaks = seq(28,44,2),limits = c(28,44))+
  scale_y_continuous(breaks = seq(0,1,0.2))+
  facet_wrap(~Ground)+
  ylab("cummulative frequency")+
  xlab("Length (cm)")+
  theme_bw()+
  theme(legend.position = "bottom")
plot(f2.3)
ggsave(filename = "Figures/f2.3.emf")
```

## Table 2.5 
Mode and Numbers of fish by Year, Ground, Sex
```{r Table2.5}
Roughy_1 <- Roughy %>% 
  mutate(GbyS= paste(Ground,sex,sep = "_")) %>%  #new variable to combine Ground and sex 
  group_by(GbyS) %>% 
  mutate(sumsum = cumsum(Length))


#create function to find the mode
Mode <- function(x) {
    if (is.numeric(x)) {
        x_table <- table(x)
        return(as.numeric(names(x_table)[which.max(x_table)]))
        }
}

#calcuate the mode
Roughy_mode <- Roughy_1 %>% 
  dplyr::group_by(year,GbyS) %>%
  dplyr::summarise(Mode = Mode(x = Length))%>%
  spread(GbyS, Mode) 
#calcuate the  counts
Roughy_nos <- Roughy_1 %>% 
  dplyr::group_by(year, GbyS) %>%
  dplyr::summarise(Numbers = n())%>%
  spread(GbyS, Numbers)

#bind the mode and counts
R_Mode_Count <-  cbind(Roughy_mode,Roughy_nos)%>%
  select(-6)#remove the year1 colomn

t2.5<-  R_Mode_Count_order <- R_Mode_Count[,c(1,2,6,3,7,4,8,5,9)]  # order the columns
  
  write.csv(t2.5,"Figures/t2.5.csv")
  #note that this table is also saved as t2.4.doc in "OR Tables.Rmd"
 
knitr::kable(R_Mode_Count_order, col.names = c("", "Nos. Fish","Mode", "Nos. Fish","Mode", "Nos. Fish","Mode", "Nos. Fish","Mode"), format = "html", caption = " Figure 2.5. Number of Fish measured and mode of standard lengths (cm) for male and femailes measured at St Helens and St Particks from 1987 to 2019") %>% 
  kableExtra::kable_styling("striped") %>%
  add_header_above(c(" ", "Female" = 2, "Male" = 2, "Female" = 2, "Male"=2)) %>%
    add_header_above(c(" ", "St Helens" = 4, "St Patricks" = 4)) %>%
    kableExtra::scroll_box(width = "100%")

```


### Figure 2.4
```{r myfig2.4, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}

library(gridExtra) #for function grid.arrange

#average SL by Area and sex
Roughy_year_lengthsum <- summarySE(Roughy_year, measurevar = "Length", groupvars = c("Shot","Date","Ground","sex")) %>% 

  mutate(Shot = as.character(Shot)) %>% 
  arrange(Length)

pd <- position_dodge(0) # in case need to experiment with operlap of values - this was not required so set 0

SH <- ggplot(Roughy_year_lengthsum %>% 
               filter(Ground == "St Helens"),
             aes(x = reorder(Shot, sort(as.numeric(Shot))),y = Length, colour = sex)) +
    geom_errorbar(aes(ymin = Length - se, ymax= Length + ci),
                width =.7, position = pd) +
   geom_point( position = pd, size = 2, width = 1)+
 facet_grid(.~Date, 
            scales = "free_x")+ # ony dates with values are included
  theme_bw()+
  theme(panel.spacing = unit(0, "mm"), # remove spacing between facets
        
        strip.background = element_rect(size = 0.1), # match default line size of theme_classic)
        strip.text.x = element_text(size =8))  +       # change font size
  theme(legend.position = c(.97, .05), #position legend bottom right
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'))+
   scale_y_continuous(breaks = seq(30,38,1),limits = c(30,38))+ 
  xlab("Shot Number, St Helens")+
  ylab("Average Standard Length (cm)")

SP <- ggplot(Roughy_year_lengthsum %>% 
               filter(Ground =="St Patricks"),
             aes(x = reorder(Shot, sort(as.numeric(Shot))),y = Length, colour = sex)) +
    geom_errorbar(aes(ymin = Length - ci, ymax= Length + ci),
                width =.7, position = pd) +
   geom_point( position = pd, size = 2)+
    facet_grid(.~Date, scales = "free_x")+
  theme_bw()+
  theme(legend.position = c(.97, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'))+
     theme(panel.spacing = unit(0, "mm"),                      
        strip.background = element_rect(size = 0.1),
        strip.text.x = element_text(size =8))  +
  scale_y_continuous(breaks = seq(30,38,1),limits = c(30,38))+
  xlab("Shot Number, st Patricks")+
  ylab("") #removes y axis


grid.arrange(SH,SP,ncol=2, widths= c(5,4)) #arrange plot into a grid (this does not save it only plots to view)

f2.4 <- arrangeGrob(SH, SP,ncol = 2, widths = c(5,4)) # creates plot as an object

                 
ggsave("Figures/f2.4.emf",f2.4) # saves the plot to file :)
```
### Table 2.6

```{r table2.6, out.width='100%', echo = T} 


#spread the weight by sex


Roughy_year_shotweight <- Roughy_year %>% 
  dplyr::group_by(Ground,sex,Shot) %>% 
  dplyr::summarise('Average weight' = round(mean(weight_kg, na.rm = TRUE),digits = 2)) %>% 
  spread(sex,'Average weight') %>% 
  rename(c("F"="fw","M"="mw"))

#spread the SD by sex
Roughy_year_ShotweightSD <- Roughy_year %>% 
  dplyr::group_by(Ground,sex,Shot) %>% 
  dplyr::summarise(SD = round(sd(weight_kg, na.rm = TRUE),digits = 2)) %>% 
  spread(sex, SD) %>% 
   rename(c("F"="fsd","M"="msd"))

#spread the nos by sex
Roughy_year_Shotnos <-Roughy_year %>% 
  dplyr::group_by(Ground,sex,Shot) %>% 
  dplyr::summarise(nos = n()) %>% 
  spread(sex, nos)

#bind weight, sd and nos, remove dupicate columns

t2.6 <- cbind(Roughy_year_shotweight,Roughy_year_ShotweightSD, Roughy_year_Shotnos) %>% 
  select(-5,-6,-9,-10,-13,-14)

t2.6 <- t2.6[ ,c(1,2,3,5,7,4,6,8)] 
write.csv(t2.6,"Figures/t2.6.csv")


#tab_df(OR2019_sum,file = "Figures/t2.3.doc") # create and save word doc . Must # as once created

knitr::kable(head(t2.6, n=14), format = "html", caption = "Table 2.6", 
             col.names = c("Ground","Shot","Avg weight (kg)", "SD","nos","Avg weight (kg)", "SD", "nos")) %>% 
  add_header_above(c("","", "Female fish "= 3, "Male Fish" = 3)) %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width  = "100%")

```

### Table 2.7

```{r t2.6}

t2.7a <- Roughy %>% 

  filter(year == 2019) %>% 

  dplyr::group_by(Ground,sex,Shot, ShotWeight_kg) %>% 

  dplyr::summarise(nos = n()) %>% 

  spread(sex, nos) %>% 

  mutate(Total= F + M) %>% 

  mutate(Fratio = round(F/Total,digits = 2)) %>% 

  mutate(Mratio = round(M/Total,digits = 2)) %>% 

  select(-c(F,M))



t2.7b <- Roughy_year %>% 

   dplyr::group_by(Ground,sex)%>% 

  dplyr::summarise(nos = n()) %>% 

  spread(sex,nos) %>% 

  mutate(Total = F + M) %>%

  mutate(Fratio = round(F/Total, digits = 2)) %>% 

  mutate(Mratio = round(M/Total, digits = 2)) %>% 

    select(-c(F,M))



t2.7c <- Roughy_year%>% 

   dplyr::group_by(sex)%>% 

  dplyr::summarise(nos = n()) %>% 

  spread(sex,nos) %>% 

  mutate(Total = F + M) %>%

  mutate(Fratio = round(F/Total, digits = 2)) %>% 

  mutate(Mratio = round(M/Total, digits = 2)) %>% 

    select(-c(F,M))

  
t2.7 <- bind_rows(t2.7a,t2.7b,t2.7c) %>% 

 rename(c("Total" = "nos fish", "Fratio"= "F ratio", "Mratio" = "M ratio"))

  

knitr::kable((t2.7), caption = "Figure 2.7")  

 # kableExtra::kable_styling("striped") %>%

  #kableExtra::scroll_box(width = "100%")



write.csv(t2.7,"Figures/t2.7.csv")

```

### Figure 2.5
```{r myfig2.5}

Roughy_year <- Roughy_year %>%
  mutate(stage = as.character(stage))%>% 
  mutate(Shot = as.character(Shot)) %>%
  mutate(Date = as.Date.character(Date,"%d/%m")) %>% 
  group_by(sex)

ORyear_sumFH <- Roughy_year_lengthsum %>% 
  filter(Ground == "St Helens", sex == "F")
ORyear_sumMH <- Roughy_year_lengthsum %>% 
  filter(Ground == "St Helens", sex == "M")
ORyear_sumFP <- Roughy_year_lengthsum %>% 
  filter(Ground == "St Patricks", sex == "F")
ORyear_sumMP <- Roughy_year_lengthsum %>% 
  filter(Ground == "St Patricks", sex == "M")

Fsh <- ggplot(Roughy_year %>% 
  filter(Ground == "St Helens", sex == "F"),
  aes(x = Date, fill = stage))+
  geom_bar(position = "fill")+
  scale_fill_manual(values = c("yellow","orange","red","grey"))+
  theme(axis.text=element_text(size=9))+
 facet_grid(.~Shot, scales = "free_x", switch ="x")+
  theme(strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"))+ # remove spacing between facets
  theme(legend.position = "none")+
  theme(axis.title.y = element_text(size = 11))+
   theme(axis.title.x = element_text(size = 11),  
        axis.ticks = element_blank()) +
  ylab("Proportion female fish")+
  xlab("")


Msh <- ggplot(Roughy_year %>% 
  filter(sex == "M",Ground =="St Helens"),
  aes(x = Date, fill = stage))+
  geom_bar(position = "fill")+
  scale_fill_manual(values = c("yellow","orange","black","black"))+
  theme(axis.text=element_text(size=9))+
  facet_grid(.~Shot, scales = "free_x", switch ="x")+
  theme(legend.position = "none",
        strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"), # remove spacing between facets
        axis.title.x = element_text(size = 11),  
        axis.ticks = element_blank(),
        axis.title.y = element_text(size = 11))+
  xlab("Shot Number and Date, St Helens")+
  ylab("Proportion male fish")
  

Fsp <- ggplot(Roughy_year %>% 
  filter(sex == "F",Ground =="St Patricks"),
  aes(x = Date, fill = stage, ))+
  geom_bar(position = "fill")+
  labs(fill = "stage")+
  scale_fill_manual(values = c("yellow","orange","red","grey"))+
  theme(axis.text=element_text(size=9))+
  facet_grid(.~Shot, scales = "free_x", switch ="x")+
  theme(strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"))+ # remove spacing between facets
   theme(axis.title.x = element_text(size = 11),  
        axis.ticks = element_blank()) +
   ylab("")+
   xlab("")

Msp <- ggplot(Roughy_year %>% 
  filter(sex == "M",Ground =="St Patricks"),
  aes(x = Date, fill = stage))+
  geom_bar(position = "fill")+
    labs(fill = "stage")+
    scale_fill_manual(values = c("yellow","orange","black","black"))+
    theme(axis.text=element_text(size=9))+
  facet_grid(.~Shot, scales = "free_x", switch ="x")+
    theme(strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"), # remove spacing between facets
        axis.title.x = element_text(size = 11),  
        axis.ticks = element_blank())+
  xlab("Shot Number and Date, St Patricks")+
  ylab("")

#plot(Fsh)
#plot(Msh)
#plot(Fsp)
#plot(Msp)

grid.arrange(Fsh,Fsp,Msh, Msp, ncol=2, nrow=2, widths= c(5,6), heights= c(7,7)) #arrange plot into a grid (this does not save it only plots to view)

f2.5 <- arrangeGrob(Fsh,Fsp,Msh, Msp, ncol=2, nrow=2, widths= c(5,6), heights= c(7,7)) # creates plot as an object

                 
ggsave("Figures/f2.5.emf",f2.5) # saves the plot to file :)
```

### Figure 2.5b
```{r my fig2.5b}
# need to replace 2019 with "year"
Roughy_2019_2 <- Roughy_2019 %>%
  mutate(stage = as.character(stage))%>% 
  mutate(Shot = as.character(Shot)) %>% 
  group_by(sex)

OR2019_sumFH <- Roughy_2019_lengthsum %>% 
  filter(Ground == "St Helens", sex == "F")
OR2019_sumMH <- Roughy_2019_lengthsum %>% 
  filter(Ground == "St Helens", sex == "M")
OR2019_sumFP <- Roughy_2019_lengthsum %>% 
  filter(Ground == "St Patricks", sex == "F")
OR2019_sumMP <- Roughy_2019_lengthsum %>% 
  filter(Ground == "St Patricks", sex == "M")

Fsh1 <- ggplot(Roughy_2019_2 %>% 
  filter(Ground == "St Helens", sex == "F"),
  aes(x = reorder(Shot, sort(as.numeric(Shot))),fill = stage))+
  geom_bar()+
  scale_fill_manual(values = c("yellow","orange","red","grey"))+
  geom_text(aes(Shot, N +20, label = N, fill = NULL),size =3, data = OR2019_sumFH)+
 ylab("Female fish")+
  xlab("")


Msh1 <- ggplot(Roughy_2019_2 %>% 
  filter(sex == "M",Ground =="St Helens"),
    aes(x = reorder(Shot, sort(as.numeric(Shot))),fill = stage))+
  geom_bar()+
  scale_fill_manual(values = c("yellow","orange","black","grey"))+
  geom_text(aes(Shot, N +20, label = N, fill = NULL),size=3, data = OR2019_sumMH)+
  xlab("Shot Number, St Helens")+
  ylab("Male fish")
  

Fsp1 <- ggplot(Roughy_2019_2 %>% 
  filter(sex == "F",Ground =="St Patricks"),
  aes(x = reorder(Shot, sort(as.numeric(Shot))),fill = stage))+
  geom_bar()+
  scale_fill_manual(values = c("yellow","orange","red","grey"))+
  geom_text(aes(Shot, N +20, label = N, fill = NULL), size=3, data = OR2019_sumFP)+
   ylab("")+
   xlab("")

Msp1 <- ggplot(Roughy_2019_2 %>% 
  filter(sex == "M",Ground =="St Patricks"),
    aes(x = reorder(Shot, sort(as.numeric(Shot))),fill = stage))+
  geom_bar()+
  scale_fill_manual(values = c("yellow","orange","black","black"))+
  geom_text(aes(Shot, N +20, label = N, fill = NULL), size=3, data = OR2019_sumMP)+
  xlab("Shot Number, St Patricks")+
  ylab("")

#plot(Fsh)
#plot(Msh)
#plot(Fsp)
#plot(Msp)

grid.arrange(Fsh1,Fsp1,Msh1, Msp1, ncol=2, nrow=2, widths= c(5,4.5), heights= c(7,7)) #arrange plot into a grid (this does not save it only plots to view)

f2.5b <- arrangeGrob(Fsh1,Fsp1,Msh1, Msp1, ncol=2, nrow=2, widths= c(5,4.5), heights= c(7,7)) # creates plot as an object

                 
ggsave("Figures/f2.5b.emf",f2.5b) # saves the plot to file :)

```



### Table 2.6b
```{r}


roughy_sum_W <- filter(Roughy, year == 2019) %>% 
  drop_na(weight_kg)
WeightbySandG <- summarySE(roughy_sum_W, measurevar = "weight_kg", groupvars = c("year","Ground","sex"))
WeightbySandG


WeightbyG <- summarySE(roughy_sum_W, measurevar = "weight_kg", groupvars = c("year","Ground"))
WeightbyG

WeightSurv <- summarySE(roughy_sum_W, measurevar = "weight_kg", groupvars = c("year"))
WeightSurv

t2.6b <- bind_rows(WeightbySandG,WeightbyG,WeightSurv) %>% 
  select(1,2,3,4,5,6)

t2.6bb1<- Roughy %>% 
  filter(year==2019) %>% 
  dplyr::group_by(year,Ground, sex) %>% 
  dplyr::summarise(weight = round(mean(weight_kg,na.rm = TRUE),digits = 2),
                   SD = round(sd(weight_kg,na.rm = TRUE),digits = 2))

t2.6bb2<- Roughy %>% 
  filter(year==2019) %>% 
  dplyr::group_by(year,Ground) %>% 
  dplyr::summarise(weight = round(mean(weight_kg,na.rm = TRUE),digits = 2),
                   SD = round(sd(weight_kg,na.rm = TRUE),digits = 2))

t2.6bb3<- Roughy %>% 
  filter(year==2019) %>% 
  dplyr::group_by(year) %>% 
  dplyr::summarise(weight = round(mean(weight_kg,na.rm = TRUE),digits = 2),
                   SD = round(sd(weight_kg,na.rm = TRUE),digits = 2))

t2.6bb <- bind_rows(t2.6bb1,t2.6bb2,t2.6bb3)
t2.6bb
write.csv(t2.6bb,file = "Figures/t2.6bb.csv")

roughy_sum_W %>% 
   dplyr::group_by(sex) %>% 
   dplyr::summarise(max(weight_kg), min(weight_kg),mean(weight_kg)) 



 1.67-1.32
 
 
```

### Table 2.8
```{r mytable2.8}
ORyear_p <- Roughy %>% 
  filter(year == 2019) 



  
  OR2019_sHg <- ORyear_p %>%                                        # table for SH both sexes
    filter(Ground == "St Helens") %>%
    dplyr::group_by(Ground, sex) %>%
    dplyr::summarise(n = n(), mean = mean(Length)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex = c("F","M"))
  
  OR2019_sPg <- ORyear_p %>%                                        #talbe for SP both sexes
    filter(Ground == "St Patricks") %>%
    dplyr::group_by(Ground, sex) %>%
    dplyr::summarise(n = n(), mean = mean(Length)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex = c("F","M"))
  
 OR2019_SHSPbg <- ORyear_p %>%                                      # table for Grounds combined sexes
    dplyr::group_by(Ground) %>%
    dplyr::summarise(n = n(), mean = mean(Length)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex = "M & F")
 
 OR2019_sex <- ORyear_p %>%                                      # table for sex combined grounds
    dplyr::group_by(sex) %>%
    dplyr::summarise(n = n(), mean = mean(Length)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(Ground = "SH and SP")
 
  OR2019_survey <- ORyear_p %>%     # talbe for survey
    dplyr::summarise(n = n(), mean = mean(round(Length))) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex ="all fish",
           Ground = "Survey total")

    
t2.8 <- bind_rows(OR2019_sHg, OR2019_sPg, OR2019_SHSPbg,OR2019_sex,OR2019_survey) %>% 
  mutate("Estimated weight (kg) based on Lyle et al(1991)" = (0.037*(mean)^2.942)/1000) %>% 
  dplyr::rename("No of fish" = n, "Avg length (cm)" = mean, "Spawning Ground"= Ground, "Prop"=Proportion) %>% 
  dplyr::mutate_at(4:6, funs(round(.,2)))  
t2.8 <- t2.8[,c(1,2,3,5,4,6)]
  


knitr::kable(tail(t2.8, n = 9), format = "html", caption = "t2.8 The proportion of male and female fish at each ground and their associated weighted averages for fish length (cm) for the 2019 Eastern Zone spawning population") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width = "100%")
write.csv(t2.8,"Figures/t2.8.csv")

```

### Table 2.9
```{r mytable2.9}
OR2019w <- Roughy %>% 
  filter(year == 2019) %>% 
  drop_na(weight_kg)


  OR2019_sHgw <- OR2019w %>%                                        # table for SH both sexes
    filter(Ground == "St Helens") %>%
    dplyr::group_by(Ground, sex) %>%
    dplyr::summarise(n = n(), mean = mean(weight_kg,na.rm = TRUE)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex = c("F","M"))
  
  OR2019_sPgw <- OR2019w %>%                                        #talbe for SP both sexes
    filter(Ground == "St Patricks") %>%
    dplyr::group_by(Ground, sex) %>%
    dplyr::summarise(n = n(), mean =  mean(weight_kg,na.rm = TRUE)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex = c("F","M"))
  
 OR2019_SHSPbgw <- OR2019w %>%                                      # table for Grounds combined sexes
    dplyr::group_by(Ground) %>%
    dplyr::summarise(n = n(), mean = mean(weight_kg,na.rm = TRUE)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex = "M & F")
 
 OR2019_sexw <- OR2019w %>%                                      # table for sex combined grounds
    dplyr::group_by(sex) %>%
    dplyr::summarise(n = n(), mean =  mean(weight_kg,na.rm = TRUE)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(Ground = "SH & SP")
 
  OR2019_surveyw <- OR2019w %>%     # talbe for survey
    dplyr::summarise(n = n(), mean =  mean(weight_kg,na.rm = TRUE)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex ="all fish",
           Ground = "Survey total")

    
t2.9 <- bind_rows(OR2019_sHgw, OR2019_sPgw, OR2019_SHSPbgw,OR2019_sexw,OR2019_surveyw) %>% 
  dplyr::rename("Nos of fish weighed" = n, "Avg of measured weights (kg)" = mean, "Spawning Ground"= Ground) %>% 
  dplyr::mutate_at(4:5, funs(round(.,2))) %>% 
  bind_cols(t2.8) %>% 
  select(-6,-7,-8,-9,-10)

t2.9 <- t2.9[,c(1,2,3,5,4,6)]
  


knitr::kable(tail(t2.9, n = 9), format = "html", caption = "Table 2.9 The proportion of male and female fish at each ground and their associated weighted averages for fish wieght (kg) for the 2019 Eastern Zone spawning population") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width = "100%")
write.csv(t2.9,"Figures/t2.9.csv")
```


##Figure 2.6

```{r myfig2.6}, out.width='100%', fig.align='center', fig.cap= "Length vs Weight, showing the predicted values (Lyle 1991)"}

Lyle_com <- function(x) {(0.037 * (x) ^ 2.942)}
S2019_fun <- function(x) {(0.056 * (x) ^ 2.8502)}
S2019M_fun <-function(x) {(0.073 * (x) ^ 2.7685)}
S2019F_fun <-function(x) {(0.071 * (x) ^ 2.7945)}

  RoughyWeight<- Roughy%>%
    filter(year==2019) %>%
    drop_na(weight_kg) %>%
    filter(weight_kg != 0) %>% 
    mutate(weight_g = weight_kg *1000) %>%
    mutate(year=as.factor(year))


f2.6<- ggplot(data = RoughyWeight,
       mapping = aes(x  = SLadj_cm,
                     y = weight_g, colour = sex)) +
  scale_colour_manual(values = c("red", "blue"))+
  geom_point(size = .3)+
  scale_x_continuous(breaks = seq(20,50,4),limits = c(20,50))+
  ylab("weight (g)")+
  xlab("Standard length (cm)")+ 
  stat_function(fun = S2019F_fun, colour="red")+
  stat_function(fun = S2019M_fun, colour = "blue") +
  stat_function(fun = Lyle_com, colour ="black")+
  stat_function(fun = S2019_fun, colour ="orange")+
  theme_bw()+
  annotate('text', x = 22, y = 3800, size =4,hjust =0,
           label = expression(paste("Weight"["Lyle sex combined"]^{}, 
                                    " = 0037 * Length"^{2.942})))+
  
  
  annotate('text', x = 22, y = 3500, size = 4, hjust = 0,colour = "orange",
           label = expression(paste("Weight"["2019 sex combined"]^{},
                                    " = 0056 * Length"^{2.850})))+
  
  annotate('text', x = 22, y = 3200, size =4,hjust =0,colour = "red",
           label = expression(paste("Weight"["2019 Females"]^{}, " = 0.071 * Length"^{2.769}))) +
  
  annotate('text', x = 22, y = 2900, size =4, hjust =0, colour = "blue",
           label = expression(paste("Weight"["2019 Males"]^{}, " = 0.073 * Length"^{2.794})))
  
           




f2.6
  
 # facet_wrap(~year)

#Your figure and table captions are automatically numbered and can be referenced in the text
#if needed: see eg. Table \@ref(tab:mytable) and Figure \@ref(fig:standard-plot)



ggsave("Figures/f2.6.emf",f2.6) # saves the plot to file :)
```

```{r myfig2.7}, out.width='100%', fig.align='center', fig.cap= "Length vs Weight, showing the predicted values (Lyle 1991)"}

Lyle_com <- function(x) {(0.037 * (x) ^ 2.942)/1000}


  RoughyWeight<- Roughy%>%
    #filter(year==2019) %>%
    drop_na(weight_kg) %>%
    filter(weight_kg != 0) %>% 
    mutate(year=as.factor(year))



f2.7<- ggplot(data = RoughyWeight,
       mapping = aes(x  = Length,
                     y = weight_kg, colour = year)) +
  geom_point()+
  #geom_smooth(method="glm",
  #          method.args=list(family=gaussian(link="log")), colour = "blue", linetype = 2, size = .5)+
   ylab("weight (kg)")+
  xlab("Standard length (cm)")+            
  stat_function(fun = Lyle_com, colour ="black")+
 facet_wrap(~year)

#Your figure and table captions are automatically numbered and can be referenced in the text
#if needed: see eg. Table \@ref(tab:mytable) and Figure \@ref(fig:standard-plot)

f2.7

ggsave("Figures/f2.7.emf",f2.7) # saves the plot to file :)
```