---
title: 'Orange Roughy Eastern Zone: Biological Summary from the Acoustic Survey 2024'
author: "sutton"
date: "2025-01-28"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(scales)
library(ggpubr)
library(Rmisc)
library(dplyr)
library(devEMF)
library(tidyr)
library(kableExtra)
library(knitr)
library(cowplot)
library(gridExtra)
```

load data
```{r}
# Biological Index
Roughy <- read_csv("Results/OREastern_Tidy_87to2019.csv", col_types = list(ShotWeight_kg = col_double(), mark = col_character()))


         
# 2024 data
Bio2024 <- read_csv("Data/RoughyBiologicals_EZ_2024.csv")
Shot2024 <- read_csv("Data/RoughyShot_EZ_2024.csv")
discard2024 <- read_csv("Data/Roughy2024_discards.csv")
```


manipulate 2024 data

```{r}
Shot2024_ed <- Shot2024 %>% 
  select(Shot,Location,Date,Lengths,Otoliths,Total_catch_weight_kg)

Bio2024_1 <- Bio2024 %>% 
  left_join(Shot2024_ed, by = "Shot") 
```

# Index data - add new data

explore fields
```{r}
colnames(Roughy)
colnames(Bio2024_1)
```
add relevant columns

```{r}

Bio <- Bio2024_1 %>%
  mutate(zone = "Eastern", year = 2024, Area_MS = Location, Survey = Voyage, stage = Gonad_Stage,
         SLadj_cm = Length_cm)

 Bio_1 <- Bio %>% 
   select(zone, year, Area_MS, Survey, Shot, sex = Sex, stage, SLadj_cm, Otoliths, vessel = Vessel, Date= Date.x, ShotWeight_kg = Total_catch_weight_kg)
```

append to the index data

```{r}
Roughy <- Roughy %>% 
  bind_rows(Bio_1)
```

# Table and Figures

## summarise bio data 2024

```{r}
BioSummary24 <- Bio2024_1 %>% 
  dplyr::group_by(Shot,Sex) %>% 
  dplyr::summarise(n= n())

BioSummary24 <- spread(BioSummary24, key = Sex,value = n)
```


add other required fields

```{r}
BioSummary24 <- BioSummary24 %>% 
  left_join(Shot2024_ed, by= "Shot") %>% 
  select(Ground = Location, Shot, Date, Total_catch_weight_kg, F, M, Total_Fish_Measured = Lengths, Otoliths_taken = Otoliths)
```

add discards

select relevant fields
```{r}
discard2024 <- discard2024 %>% 
  select(Shot,retained,discarded)
```
 
 join to summary
 
```{r}
BioSummary24 <- BioSummary24 %>% 
  left_join(discard2024, by = "Shot")
```

Write data
```{r}
#write.csv(BioSummary24,"C:/Users/sut109/OneDrive - CSIRO/roughy - working/Figures/t2.4_sumshot_24.csv")
```




---
### summary table with additional 2024 data

```{r mytable, out.width='100%', echo = T}


#remame and select columns
Roughy_1 <- Roughy %>% 
  mutate(Length =round(SLadj_cm)) %>% #round values for length in cm 2019 was to nearest mm
  mutate(Ground = Area_MS) %>% 
  drop_na(Ground) %>% 
  #drop_na(Length) %>% 
  drop_na(sex) %>% 
  select(Ground, year, Shot, Date,sex,stage, SLadj_cm,Length,ShotWeight_kg, weight_kg, Otoliths)


Roughy_DataSummary <- Roughy_1 %>% 
  dplyr::group_by(year, Ground,sex) %>% 
  dplyr::summarise(n= n())

Roughy_DataSummary <- spread(Roughy_DataSummary, key = year,value = n)


knitr::kable(tail(Roughy_DataSummary, n = 5),caption = "Summary data: Number of Fish measured for orange roughy eastern zone spawning surveys by, year, Area, sex") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width = "100%")
```

```{r}
  
Roughyminmax <- Roughy_1 %>% 
  dplyr::group_by(Ground,sex) %>% 
  dplyr::summarise(round(mean(Length),digits = 2))

Roughyminmax
```

### Length - index
Average Standard Length (cm) of orange roughy Eastern Zone spawning population, showing males and females at St Helens and St Patricks for 1987 to 2024. Error bars are 95% confidence limits (updated from Kloser et al. 2016)
```{r myfig2.1}
#average SL by Area and sex

library(Rmisc)
library(ggplot2)

roughy_sum <- summarySE(Roughy_1, measurevar = "Length", groupvars = c("year","Ground","sex")) 


  
pd <- position_dodge(0.6) # move position horizontally b/c of overlap

f2.1 <- ggplot(roughy_sum, aes(x = year,y = Length, shape = Ground, colour = sex)) +
  geom_errorbar(aes(ymin = Length - ci, ymax= Length + ci),
                width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_line(position = pd) +
  xlab("Eastern Zone Spawning Population by year for July")+
  ylab("Average Standard Length (cm)")+
  scale_y_continuous(breaks = seq(27,38,1),limits = c(27,38))+
  scale_x_continuous(limits = c(1986,2024), breaks = seq(1986,2024,4))+
  theme_bw()+
  theme(legend.position = c(.95, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'),
         plot.background = element_rect(linetype =  "solid") )

  
plot(f2.1)
ggsave("Figures/LF_index.emf")
ggsave("Figures/LF_index.jpg")
```


## DATA for 2024
```{r}
Roughy_2 <- Roughy_1 %>% 
filter(year == 2024)
```

### length data table
```{r}

Roughy_2_length_GroundbySex <- summarySE(Roughy_2, measurevar = "Length", 
                                      groupvars = c("Ground","sex"))
Roughy_2_length_GroundbySex

Roughy_2_length_Ground <- summarySE(Roughy_2, measurevar = "Length", 
                                      groupvars = c("Ground"))
Roughy_2_length_Ground
```

### summaries

#### Table 2.4 (needs updating for future surveys to include shot without measures in "OR.R")
```{r table 2.4, echo=FALSE}
#need to update code to include shot weights for shots where fish were not measured **
# fitler data for 2019


Roughy_2_sum <- Roughy_2 %>% 
  dplyr::group_by(Ground,Shot,Date,ShotWeight_kg,sex) %>% 
  dplyr::summarise(num_rows = n()) %>% 
  spread(sex,num_rows) %>% 
  mutate(`Total` = F + M) 

# creating ground/ and survey totals for:

#St Helens
TSH <- Roughy_2_sum %>% 
  filter(Ground == "St Helens") %>% 
  dplyr::group_by(Ground) %>% 
  dplyr::summarise(ShotWeight_kg =sum(ShotWeight_kg), F= sum(F), M = sum(M), Total =sum(Total)) %>% 
  mutate(Ground = "St Helens Total")

#St Patricks
TSP <- Roughy_2_sum %>% 
  filter(Ground == "St Patricks") %>% 
  dplyr::group_by(Ground) %>% 
  dplyr::summarise(ShotWeight_kg =sum(ShotWeight_kg, na.rm = TRUE), F= sum(F), M = sum(M), Total =sum(Total)) %>% 
  mutate(Ground = "St Patricks Total")

# totals for SH and SP
TSHSP <- Roughy_2_sum %>% 
  summarise(ShotWeight_kg = sum(ShotWeight_kg, na.rm = TRUE), F= sum(F), M = sum(M), Total =sum(Total)) %>% 
  mutate(Ground = "Survey Total")

#bind the shot and subtotals and total
t2.4 <- bind_rows(Roughy_2_sum, TSH, TSP,TSHSP) %>%   # put all together for totals
  dplyr::rename("Shot Weight (kg)" = ShotWeight_kg, "Total Nos Fish" = Total)
  

write.csv(t2.4,"Figures/t2.4.csv")

knitr::kable(head(t2.4, n=20), format = "html", caption = "Table 2.4 Summary of shot dates and weights for St Helens and St Patricks, Eastern Zone, 2019") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width  = "100%")
```

### Standard Length by sex and ground
```{r myfig2.2}
OR_LF <- Roughy_2 %>% # This is the ONE
  group_by(Ground, sex) %>% 
  add_tally() %>% 
  group_by(Ground, sex, Length) %>% 
  dplyr::summarise(prop = n() / n[1])

f2.2 <- ggplot(OR_LF, aes(Length, prop, fill = sex)) +
  geom_col(position = "identity", alpha = 0.6) +
  scale_x_continuous(breaks = seq(28,46,2),limits = c(28,46))+
  facet_wrap(~Ground)+
   theme_bw()+
  ylab("proportion")+
  xlab("Length (cm)")+
  theme(legend.position = "bottom")
plot(f2.2)
ggsave(filename = "Figures/LF2024.emf")
ggsave(filename = "Figures/LF2024.jpg")
```


### Cummunlative Frequency
```{r myfig2.3}

#using stat_ecdf to calculate the cummulative frequency

f2.3 <- ggplot(Roughy_2,aes(x=Length, colour= sex)) +
   stat_ecdf(geom = "smooth")+
  scale_x_continuous(breaks = seq(28,44,2),limits = c(28,44))+
  scale_y_continuous(breaks = seq(0,1,0.2))+
  facet_wrap(~Ground)+
  ylab("cummulative frequency")+
  xlab("Length (cm)")+
  theme_bw()+
  theme(legend.position = "bottom")
plot(f2.3)
ggsave(filename = "Figures/cumLF2024.emf")
```

## Mode and Numbers of fish by Year, Ground, Sex 

```{r Table2.5}
#need to group by Ground and sex togehter so need to make a combined variable
Roughy_3 <- Roughy_1 %>% 
  mutate(GbyS= paste(Ground,sex,sep = "_")) %>%  #new variable to combine Ground and sex 
  group_by(GbyS) %>% 
  mutate(sumsum = cumsum(Length))


#create function to find the mode
Mode <- function(x) {
    if (is.numeric(x)) {
        x_table <- table(x)
        return(as.numeric(names(x_table)[which.max(x_table)]))
        }
}

#calcuate the mode
Roughy_mode <- Roughy_3 %>% 
  dplyr::group_by(year,GbyS) %>%
  dplyr::summarise(Mode = Mode(x = Length))%>%
  spread(GbyS, Mode) 
#calcuate the  counts
Roughy_nos <- Roughy_3 %>% 
  dplyr::group_by(year, GbyS) %>%
  dplyr::summarise(Numbers = n())%>%
  spread(GbyS, Numbers)

#bind the mode and counts
R_Mode_Count <-  cbind(Roughy_mode,Roughy_nos)%>%
  select(-6)#remove the year1 colomn

t2.5<-  R_Mode_Count_order <- R_Mode_Count[,c(1,2,6,3,7,4,8,5,9)]  # order the columns
  
  write.csv(t2.5,"Figures/Lmode_index.csv")
  #note that this table is also saved as t2.4.doc in "OR Tables.Rmd"
 
knitr::kable(R_Mode_Count_order, col.names = c("", "Nos. Fish","Mode", "Nos. Fish","Mode", "Nos. Fish","Mode", "Nos. Fish","Mode"), format = "html", caption = " Figure 2.5. Number of Fish measured and mode of standard lengths (cm) for male and femailes measured at St Helens and St Particks from 1987 to 2019") %>% 
  kableExtra::kable_styling("striped") %>%
  add_header_above(c(" ", "Female" = 2, "Male" = 2, "Female" = 2, "Male"=2)) %>%
    add_header_above(c(" ", "St Helens" = 4, "St Patricks" = 4)) %>%
    kableExtra::scroll_box(width = "100%")

```

### Figure 2.4
```{r myfig2.4, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}

library(gridExtra) #for function grid.arrange

#average SL by Area and sex
Roughy_2_lengthsum <- summarySE(Roughy_2, measurevar = "Length", groupvars = c("Shot","Date","Ground","sex")) %>% 

  mutate(Shot = as.character(Shot)) %>% 
  arrange(Length)

pd <- position_dodge(0) # in case need to experiment with operlap of values - this was not required so set 0

SH <- ggplot(Roughy_2_lengthsum %>% 
               filter(Ground == "St Helens"),
             aes(x = reorder(Shot, sort((Shot))),y = Length, colour = sex)) +
    geom_errorbar(aes(ymin = Length - se, ymax= Length + ci),
                width =.7, position = pd) +
   geom_point( position = pd, size = 2, width = 1)+
# facet_grid(.~Date, 
 #           scales = "free_x")+ # ony dates with values are included
  theme_bw()+
  theme(panel.spacing = unit(0, "mm"), # remove spacing between facets
        
        strip.background = element_rect(size = 0.1), # match default line size of theme_classic)
        strip.text.x = element_text(size =8))  +       # change font size
  theme(legend.position = c(.97, .05), #position legend bottom right
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'))+
   scale_y_continuous(breaks = seq(32,40,1),limits = c(32,40))+ 
  xlab("Shot Number, St Helens")+
  ylab("Average Standard Length (cm)")

SP <- ggplot(Roughy_2_lengthsum %>% 
               filter(Ground =="St Patricks"),
             aes(x = reorder(Shot, sort((Shot))),y = Length, colour = sex)) +
    geom_errorbar(aes(ymin = Length - ci, ymax= Length + ci),
                width =.7, position = pd) +
   geom_point( position = pd, size = 2)+
    #facet_grid(.~Date, scales = "free_x")+
  theme_bw()+
  theme(legend.position = c(.97, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'))+
     theme(panel.spacing = unit(0, "mm"),                      
        strip.background = element_rect(size = 0.1),
        strip.text.x = element_text(size =8))  +
  scale_y_continuous(breaks = seq(32,40,1),limits = c(32,40))+
  xlab("Shot Number, st Patricks")+
  ylab("") #removes y axis


grid.arrange(SH,SP,ncol=2, widths= c(5,4)) #arrange plot into a grid (this does not save it only plots to view)

f2.4 <- arrangeGrob(SH, SP,ncol = 2, widths = c(5,4)) # creates plot as an object

                 
ggsave("Figures/shotvariablity.emf",f2.4) # saves the plot to file :)
```

### Gonad Stages
```{r f2.5}
#need to group the frame by sex 
Roughy_4 <- Roughy_1%>%
  filter(year==2024) %>% 
  mutate(stage = as.character(stage))%>% 
 # mutate(Shot = as.character(Shot)) %>% 
  mutate(Date= as.Date(Date,"%d/%m")) %>% 
  group_by(sex)

OR2024_sumFH <- Roughy_2_lengthsum %>% 
  filter(Ground == "St Helens", sex == "F")
OR2024_sumMH <- Roughy_2_lengthsum %>% 
  filter(Ground == "St Helens", sex == "M")
OR2024_sumFP <- Roughy_2_lengthsum %>% 
  filter(Ground == "St Patricks", sex == "F")
OR2024_sumMP <- Roughy_2_lengthsum %>% 
  filter(Ground == "St Patricks", sex == "M")

Fsh <- ggplot(Roughy_4 %>% 
  filter(Ground == "St Helens", sex == "F"),
  aes(x = Date, fill = stage))+
  geom_bar(position = "fill")+
  scale_fill_manual(values = c("yellow","orange","red","grey"))+
  theme(axis.text=element_text(size=9))+
 facet_grid(.~Shot, scales = "free_x", switch ="x")+
  theme(strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"))+ # remove spacing between facets
  theme(legend.position = "none")+
  theme(axis.title.y = element_text(size = 11))+
   theme(axis.title.x = element_text(size = 11),  
        axis.ticks = element_blank()) +
  ylab("Proportion female fish")+
  xlab("")

Msh <- ggplot(Roughy_4 %>% 
  filter(sex == "M",Ground =="St Helens"),
  aes(x = Date, fill = stage))+
  geom_bar(position = "fill")+
  scale_fill_manual(values = c("white","yellow","orange","black","black"))+
  theme(axis.text=element_text(size=9))+
  facet_grid(.~Shot, scales = "free_x", switch ="x")+
  theme(legend.position = "none",
        strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"), # remove spacing between facets
        axis.title.x = element_text(size = 11),  
        axis.ticks = element_blank(),
        axis.title.y = element_text(size = 11))+
  xlab("Shot Number and Date, St Helens")+
  ylab("Proportion male fish")

Fsp <- ggplot(Roughy_4 %>% 
  filter(sex == "F",Ground =="St Patricks"),
  aes(x = Date, fill = stage, ))+
  geom_bar(position = "fill")+
  labs(fill = "stage")+
  scale_fill_manual(values = c("yellow","orange","red","grey"))+
  theme(axis.text=element_text(size=9))+
  facet_grid(.~Shot, scales = "free_x", switch ="x")+
  theme(strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"))+ # remove spacing between facets
   theme(axis.title.x = element_text(size = 11),  
        axis.ticks = element_blank()) +
   ylab("")+
   xlab("")

Msp <- ggplot(Roughy_4 %>% 
  filter(sex == "M",Ground =="St Patricks"),
  aes(x = Date, fill = stage))+
  geom_bar(position = "fill")+
    labs(fill = "stage")+
    scale_fill_manual(values = c("white","yellow","orange","black","black"))+
    theme(axis.text=element_text(size=9))+
  facet_grid(.~Shot, scales = "free_x", switch ="x")+
    theme(strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"), # remove spacing between facets
        axis.title.x = element_text(size = 11),  
        axis.ticks = element_blank())+
  xlab("Shot Number and Date, St Patricks")+
  ylab("")

plot(Fsh)
plot(Msh)
plot(Fsp)
plot(Msp)

grid.arrange(Fsh,Fsp,Msh, Msp, ncol=2, nrow=2, widths= c(12,12), heights= c(10,10)) #arrange plot into a grid (this does not save it only plots to view)

f2.5 <- arrangeGrob(Fsh,Fsp,Msh, Msp, ncol=2, nrow=2, widths= c(5,6), heights= c(7,7)) # creates plot as an object

ggsave("Figures/GonadStage.emf",f2.5) # saves the plot to file :)
```




### Figure 2.5b
```{r my fig2.5b}


Fsh1 <- ggplot(Roughy_4 %>% 
  filter(Ground == "St Helens", sex == "F"),
  aes(x = reorder(Shot, sort(as.numeric(Shot))),fill = stage))+
  geom_bar()+
  scale_fill_manual(values = c("yellow","orange","red","grey"))+
  geom_text(aes(Shot, N +20, label = N, fill = NULL),size =3, data = OR2019_sumFH)+
 ylab("Female fish")+
  xlab("")


Msh1 <- ggplot(Roughy_4 %>% 
  filter(sex == "M",Ground =="St Helens"),
    aes(x = reorder(Shot, sort(as.numeric(Shot))),fill = stage))+
  geom_bar()+
  scale_fill_manual(values = c("white","yellow","orange","black","grey"))+
  geom_text(aes(Shot, N +20, label = N, fill = NULL),size=3, data = OR2019_sumMH)+
  xlab("Shot Number, St Helens")+
  ylab("Male fish")
  

Fsp1 <- ggplot(Roughy_4 %>% 
  filter(sex == "F",Ground =="St Patricks"),
  aes(x = reorder(Shot, sort(as.numeric(Shot))),fill = stage))+
  geom_bar()+
  scale_fill_manual(values = c("yellow","orange","red","grey"))+
  geom_text(aes(Shot, N +20, label = N, fill = NULL), size=3, data = OR2019_sumFP)+
   ylab("")+
   xlab("")

Msp1 <- ggplot(Roughy_4 %>% 
  filter(sex == "M",Ground =="St Patricks"),
    aes(x = reorder(Shot, sort(as.numeric(Shot))),fill = stage))+
  geom_bar()+
  scale_fill_manual(values = c("white","yellow","orange","black","black"))+
  geom_text(aes(Shot, N +20, label = N, fill = NULL), size=3, data = OR2019_sumMP)+
  xlab("Shot Number, St Patricks")+
  ylab("")

#plot(Fsh)
#plot(Msh)
#plot(Fsp)
#plot(Msp)

grid.arrange(Fsh1,Fsp1,Msh1, Msp1, ncol=2, nrow=2, widths= c(5,4.5), heights= c(7,7)) #arrange plot into a grid (this does not save it only plots to view)

f2.5b <- arrangeGrob(Fsh1,Fsp1,Msh1, Msp1, ncol=2, nrow=2, widths= c(5,4.5), heights= c(7,7)) # creates plot as an object

                 
ggsave("Figures/f2.5b.emf",f2.5b) # saves the plot to file :)

```