---
title: 'Orange Roughy Eastern Zone: Biological Summary from the Acoustic Survey 2024'
author: "sutton"
date: "2025-01-28"
output: html_document
---

# Set up

libraries required

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(scales)
library(ggpubr)
library(Rmisc)
library(dplyr)
library(devEMF)
library(tidyr)
library(kableExtra)
library(knitr)
library(cowplot)
library(gridExtra)
library(lubridate)
```

load data
```{r}
# Biological Index
Roughy <- read_csv("Results/OREastern_Tidy_87to2019.csv", col_types = list(ShotWeight_kg = col_double(), mark = col_character()))

Age2019 <- read_csv("Data/RoughyAge2019.csv")
         
# 2024 data
Bio2024 <- read_csv("Data/RoughyBiologicals_EZ_2024.csv")
Shot2024 <- read_csv("Data/RoughyShot_EZ_2024.csv")
discard2024 <- read_csv("Data/Roughy2024_discards.csv")
```


manipulate 2024 data

```{r}
Shot2024_ed <- Shot2024 %>% 
  select(Shot,Location,Day, Month, Year, Date,Lengths,Otoliths,Total_catch_weight_kg)

Bio2024_1 <- Bio2024 %>% 
  left_join(Shot2024_ed, by = "Shot") 
```

## Index data - add new data

explore fields
```{r}
colnames(Roughy)
colnames(Bio2024_1)
```
add relevant columns

```{r}

Bio <- Bio2024_1 %>%
  mutate(zone = "Eastern", year = 2024, Area_MS = Location, Survey = Voyage, stage = Gonad_Stage,
         SLadj_cm = Length_cm)

 Bio_1 <- Bio %>% 
   select(zone, year, Area_MS, Survey, Shot, sex = Sex, stage, SLadj_cm, Otoliths, vessel = Vessel, Date= Date.x, ShotWeight_kg = Total_catch_weight_kg)
```

append to the index data

```{r}
Roughy <- Roughy %>% 
  bind_rows(Bio_1)
```

# Table and Figures

## Summarise bio data 2024

```{r}
BioSummary24 <- Bio2024_1 %>% 
  dplyr::group_by(Shot,Sex) %>% 
  dplyr::summarise(n= n())

BioSummary24 <- spread(BioSummary24, key = Sex,value = n)
```


add other required fields

```{r}
BioSummary24 <- BioSummary24 %>% 
  left_join(Shot2024_ed, by= "Shot") %>% 
  select(Ground = Location, Shot, Date, Total_catch_weight_kg, F, M, Total_Fish_Measured = Lengths, Otoliths_taken = Otoliths)
```

add discards

select relevant fields
```{r}
discard2024 <- discard2024 %>% 
  select(Shot,retained,discarded)
```
 
 join to summary
 
```{r}
BioSummary24 <- BioSummary24 %>% 
  left_join(discard2024, by = "Shot")
```

Write data
```{r}
#write.csv(BioSummary24,"C:/Users/sut109/OneDrive - CSIRO/roughy - working/Figures/t2.4_sumshot_24.csv")
```




---
### summary table with additional 2024 data

```{r mytable, out.width='100%', echo = T}


#remame and select columns
Roughy_1 <- Roughy %>% 
  mutate(Length =round(SLadj_cm)) %>% #round values for length in cm 2019 was to nearest mm
  mutate(Ground = Area_MS) %>% 
  drop_na(Ground) %>% 
  #drop_na(Length) %>% 
  drop_na(sex) %>% 
  select(Ground, year, Shot, Date,sex,stage, SLadj_cm,Length,ShotWeight_kg, weight_kg, Otoliths)


Roughy_DataSummary <- Roughy_1 %>% 
  dplyr::group_by(year, Ground,sex) %>% 
  dplyr::summarise(n= n())

Roughy_DataSummary <- spread(Roughy_DataSummary, key = year,value = n)


knitr::kable(tail(Roughy_DataSummary, n = 5),caption = "Summary data: Number of Fish measured for orange roughy eastern zone spawning surveys by, year, Area, sex") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width = "100%")
```

```{r}
  
Roughyminmax <- Roughy_1 %>% 
  dplyr::group_by(Ground,sex) %>% 
  dplyr::summarise(round(mean(Length),digits = 2))

Roughyminmax
```

### Length - index
Average Standard Length (cm) of orange roughy Eastern Zone spawning population, showing males and females at St Helens and St Patricks for 1987 to 2024. Error bars are 95% confidence limits (updated from Kloser et al. 2016)
```{r myfig2.1}
#average SL by Area and sex

library(Rmisc)
library(ggplot2)

roughy_sum <- summarySE(Roughy_1, measurevar = "Length", groupvars = c("year","Ground","sex")) 


  
pd <- position_dodge(0.6) # move position horizontally b/c of overlap

f2.1 <- ggplot(roughy_sum, aes(x = year,y = Length, shape = Ground, colour = sex)) +
  geom_errorbar(aes(ymin = Length - ci, ymax= Length + ci),
                width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_line(position = pd) +
  xlab("Eastern Zone Spawning Population by year for July")+
  ylab("Average Standard Length (cm)")+
  scale_y_continuous(breaks = seq(27,38,1),limits = c(27,38))+
  scale_x_continuous(limits = c(1986,2024), breaks = seq(1986,2024,4))+
  theme_bw()+
  theme(legend.position = c(.95, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'),
         plot.background = element_rect(linetype =  "solid") )

  
plot(f2.1)
ggsave("Figures/LF_index.emf")
ggsave("Figures/LF_index.jpg")
```


## DATA for 2024
```{r}
Roughy_2 <- Roughy_1 %>% 
filter(year == 2024)
```

### length data table
```{r}

Roughy_2_length_GroundbySex <- summarySE(Roughy_2, measurevar = "Length", 
                                      groupvars = c("Ground","sex"))
Roughy_2_length_GroundbySex

Roughy_2_length_Ground <- summarySE(Roughy_2, measurevar = "Length", 
                                      groupvars = c("Ground"))
Roughy_2_length_Ground
```

### summaries

#### Table 2.4 (needs updating for future surveys to include shot without measures in "OR.R")
```{r table 2.4, echo=FALSE}
#need to update code to include shot weights for shots where fish were not measured **
# fitler data for 2019


Roughy_2_sum <- Roughy_2 %>% 
  dplyr::group_by(Ground,Shot,Date,ShotWeight_kg,sex) %>% 
  dplyr::summarise(num_rows = n()) %>% 
  spread(sex,num_rows) %>% 
  mutate(`Total` = F + M) 

# creating ground/ and survey totals for:

#St Helens
TSH <- Roughy_2_sum %>% 
  filter(Ground == "St Helens") %>% 
  dplyr::group_by(Ground) %>% 
  dplyr::summarise(ShotWeight_kg =sum(ShotWeight_kg), F= sum(F), M = sum(M), Total =sum(Total)) %>% 
  mutate(Ground = "St Helens Total")

#St Patricks
TSP <- Roughy_2_sum %>% 
  filter(Ground == "St Patricks") %>% 
  dplyr::group_by(Ground) %>% 
  dplyr::summarise(ShotWeight_kg =sum(ShotWeight_kg, na.rm = TRUE), F= sum(F), M = sum(M), Total =sum(Total)) %>% 
  mutate(Ground = "St Patricks Total")

# totals for SH and SP
TSHSP <- Roughy_2_sum %>% 
  summarise(ShotWeight_kg = sum(ShotWeight_kg, na.rm = TRUE), F= sum(F), M = sum(M), Total =sum(Total)) %>% 
  mutate(Ground = "Survey Total")

#bind the shot and subtotals and total
t2.4 <- bind_rows(Roughy_2_sum, TSH, TSP,TSHSP) %>%   # put all together for totals
  dplyr::rename("Shot Weight (kg)" = ShotWeight_kg, "Total Nos Fish" = Total)
  

write.csv(t2.4,"Figures/t2.4.csv")

knitr::kable(head(t2.4, n=20), format = "html", caption = "Table 2.4 Summary of shot dates and weights for St Helens and St Patricks, Eastern Zone, 2019") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width  = "100%")
```

### Standard Length by sex and ground
```{r myfig2.2}
OR_LF <- Roughy_2 %>% # This is the ONE
  group_by(Ground, sex) %>% 
  add_tally() %>% 
  group_by(Ground, sex, Length) %>% 
  dplyr::summarise(prop = n() / n[1])

f2.2 <- ggplot(OR_LF, aes(Length, prop, fill = sex)) +
  geom_col(position = "identity", alpha = 0.6) +
  scale_x_continuous(breaks = seq(28,46,2),limits = c(28,46))+
  facet_wrap(~Ground)+
   theme_bw()+
  ylab("proportion")+
  xlab("Length (cm)")+
  theme(legend.position = "bottom")
plot(f2.2)
ggsave(filename = "Figures/LF2024.emf")
ggsave(filename = "Figures/LF2024.jpg")
```


### Cummunlative Frequency
```{r myfig2.3}

#using stat_ecdf to calculate the cummulative frequency

f2.3 <- ggplot(Roughy_2,aes(x=Length, colour= sex)) +
   stat_ecdf(geom = "smooth")+
  scale_x_continuous(breaks = seq(28,44,2),limits = c(28,44))+
  scale_y_continuous(breaks = seq(0,1,0.2))+
  facet_wrap(~Ground)+
  ylab("cummulative frequency")+
  xlab("Length (cm)")+
  theme_bw()+
  theme(legend.position = "bottom")
plot(f2.3)
ggsave(filename = "Figures/cumLF2024.emf")
```

## Mode and Numbers of fish by Year, Ground, Sex 

```{r Table2.5}
#need to group by Ground and sex togehter so need to make a combined variable
Roughy_3 <- Roughy_1 %>% 
  mutate(GbyS= paste(Ground,sex,sep = "_")) %>%  #new variable to combine Ground and sex 
  group_by(GbyS) %>% 
  mutate(sumsum = cumsum(Length))


#create function to find the mode
Mode <- function(x) {
    if (is.numeric(x)) {
        x_table <- table(x)
        return(as.numeric(names(x_table)[which.max(x_table)]))
        }
}

#calcuate the mode
Roughy_mode <- Roughy_3 %>% 
  dplyr::group_by(year,GbyS) %>%
  dplyr::summarise(Mode = Mode(x = Length))%>%
  spread(GbyS, Mode) 
#calcuate the  counts
Roughy_nos <- Roughy_3 %>% 
  dplyr::group_by(year, GbyS) %>%
  dplyr::summarise(Numbers = n())%>%
  spread(GbyS, Numbers)

#bind the mode and counts
R_Mode_Count <-  cbind(Roughy_mode,Roughy_nos)%>%
  select(-6)#remove the year1 colomn

t2.5<-  R_Mode_Count_order <- R_Mode_Count[,c(1,2,6,3,7,4,8,5,9)]  # order the columns
  
  write.csv(t2.5,"Figures/Lmode_index.csv")
  #note that this table is also saved as t2.4.doc in "OR Tables.Rmd"
 
knitr::kable(R_Mode_Count_order, col.names = c("", "Nos. Fish","Mode", "Nos. Fish","Mode", "Nos. Fish","Mode", "Nos. Fish","Mode"), format = "html", caption = " Figure 2.5. Number of Fish measured and mode of standard lengths (cm) for male and femailes measured at St Helens and St Particks from 1987 to 2019") %>% 
  kableExtra::kable_styling("striped") %>%
  add_header_above(c(" ", "Female" = 2, "Male" = 2, "Female" = 2, "Male"=2)) %>%
    add_header_above(c(" ", "St Helens" = 4, "St Patricks" = 4)) %>%
    kableExtra::scroll_box(width = "100%")

```

## Average Length by Shot
### Figure 2.4
```{r myfig2.4, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}

library(gridExtra) #for function grid.arrange

#average SL by Area and sex
Roughy_2_lengthsum <- summarySE(Roughy_2, measurevar = "Length", groupvars = c("Shot","Date","Ground","sex")) %>% 

  mutate(Shot = as.character(Shot)) %>% 
  arrange(Length)

pd <- position_dodge(0) # in case need to experiment with operlap of values - this was not required so set 0

SH <- ggplot(Roughy_2_lengthsum %>% 
               filter(Ground == "St Helens"),
             aes(x = reorder(Shot, sort((Shot))),y = Length, colour = sex)) +
    geom_errorbar(aes(ymin = Length - se, ymax= Length + ci),
                width =.7, position = pd) +
   geom_point( position = pd, size = 2, width = 1)+
# facet_grid(.~Date, 
 #           scales = "free_x")+ # ony dates with values are included
  theme_bw()+
  theme(panel.spacing = unit(0, "mm"), # remove spacing between facets
        
        strip.background = element_rect(size = 0.1), # match default line size of theme_classic)
        strip.text.x = element_text(size =8))  +       # change font size
  theme(legend.position = c(.97, .05), #position legend bottom right
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'))+
   scale_y_continuous(breaks = seq(32,40,1),limits = c(32,40))+ 
  xlab("Shot Number, St Helens")+
  ylab("Average Standard Length (cm)")

SP <- ggplot(Roughy_2_lengthsum %>% 
               filter(Ground =="St Patricks"),
             aes(x = reorder(Shot, sort((Shot))),y = Length, colour = sex)) +
    geom_errorbar(aes(ymin = Length - ci, ymax= Length + ci),
                width =.7, position = pd) +
   geom_point( position = pd, size = 2)+
    #facet_grid(.~Date, scales = "free_x")+
  theme_bw()+
  theme(legend.position = c(.97, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'))+
     theme(panel.spacing = unit(0, "mm"),                      
        strip.background = element_rect(size = 0.1),
        strip.text.x = element_text(size =8))  +
  scale_y_continuous(breaks = seq(32,40,1),limits = c(32,40))+
  xlab("Shot Number, st Patricks")+
  ylab("") #removes y axis


grid.arrange(SH,SP,ncol=2, widths= c(5,4)) #arrange plot into a grid (this does not save it only plots to view)

f2.4 <- arrangeGrob(SH, SP,ncol = 2, widths = c(5,4)) # creates plot as an object

                 
ggsave("Figures/shotvariablity.emf",f2.4) # saves the plot to file :)
```

### Gonad Stages
```{r f2.5}
#need to group the frame by sex 
Roughy_4 <- Roughy_1%>%
  filter(year==2024) %>% 
  mutate(stage = as.character(stage))%>% 
 # mutate(Shot = as.character(Shot)) %>% 
  mutate(Date= as.Date(Date,"%d/%m")) %>% 
  #select(-Date) %>% 
  group_by(sex)

OR2024_sumFH <- Roughy_2_lengthsum %>% 
  filter(Ground == "St Helens", sex == "F")
OR2024_sumMH <- Roughy_2_lengthsum %>% 
  filter(Ground == "St Helens", sex == "M")
OR2024_sumFP <- Roughy_2_lengthsum %>% 
  filter(Ground == "St Patricks", sex == "F")
OR2024_sumMP <- Roughy_2_lengthsum %>% 
  filter(Ground == "St Patricks", sex == "M")

# Fsh <- ggplot(Roughy_4 %>% 
#   filter(Ground == "St Helens", sex == "F"),
#   aes(x = Date, fill = stage))+
#   geom_bar(position = "fill")+
#   scale_fill_manual(values = c("yellow","orange","red","grey"))+
#   theme(axis.text=element_text(size=9))+
#  facet_grid(.~Shot, scales = "free_x", switch ="x")+
#   theme(strip.text.x = element_text(size =8),
#         panel.spacing = unit(0,"mm"))+ # remove spacing between facets
#   theme(legend.position = "none")+
#   theme(axis.title.y = element_text(size = 11))+
#    theme(axis.title.x = element_text(size = 11),  
#         axis.ticks = element_blank()) +
#   ylab("Proportion female fish")+
#   xlab("")

Fsh <- ggplot(Roughy_4 %>% 
  filter(Ground == "St Helens", sex == "F"),
  aes(x = Date, fill = stage))+
  geom_bar(position = "fill")+
  scale_fill_manual(values = c("yellow","orange","red","grey"))+
  theme(axis.text= element_blank())+
 facet_grid(.~Shot, scales = "free_x", switch ="x")+
  theme(strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"))+ # remove spacing between facets
  #theme(legend.position = "none")+
  theme(legend.text=element_text(size=8))+
  theme(legend.title = element_text(size =8))+
  theme(legend.key.size = unit(.3,'cm'))+
  theme(axis.title.y = element_text(size = 8))+
   theme(axis.title.x = element_text(size = 8),  
        axis.ticks = element_blank()) +
  ylab("Proportion female fish")+
  xlab("")

Msh <- ggplot(Roughy_4 %>% 
  filter(sex == "M",Ground =="St Helens"),
  aes(x = Date, fill = stage))+
  geom_bar(position = "fill")+
  scale_fill_manual(values = c("beige","yellow","orange","black","black"))+
  theme(axis.text= element_blank())+
 facet_grid(.~Shot, scales = "free_x", switch ="x")+
    theme(strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"), # remove spacing between facets
        axis.title.x = element_text(size = 8),  
        axis.ticks = element_blank())+
  theme(legend.text=element_text(size=8))+
  theme(legend.title = element_text(size =8))+
  theme(legend.key.size = unit(.3,'cm'))+
    theme(axis.title.y = element_text(size = 8))+
     theme(axis.title.x = element_text(size = 8),  
        axis.ticks = element_blank()) +
  xlab("Shot Number, St Helens")+
  ylab("Proportion male fish")

Fsp <- ggplot(Roughy_4 %>% 
  filter(sex == "F",Ground =="St Patricks"),
  aes(x = Date, fill = stage, ))+
  geom_bar(position = "fill")+
  labs(fill = "stage")+
  scale_fill_manual(values = c("yellow","orange","red","grey"))+
 theme(axis.text= element_blank())+
  facet_grid(.~Shot, scales = "free_x", switch ="x")+
  theme(strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"))+ # remove spacing between facets
   theme(axis.title.x = element_text(size = 8),  
        axis.ticks = element_blank()) +
  theme(legend.text=element_text(size=8))+
  theme(legend.title = element_text(size =8))+
  theme(legend.key.size = unit(.3,'cm'))+
   ylab("")+
   xlab("")

Msp <- ggplot(Roughy_4 %>% 
  filter(sex == "M",Ground =="St Patricks"),
  aes(x = Date, fill = stage))+
  geom_bar(position = "fill")+
    labs(fill = "stage")+
    scale_fill_manual(values = c("yellow","orange","black","black"))+
    theme(axis.text= element_blank())+
  facet_grid(.~Shot, scales = "free_x", switch ="x")+
    theme(strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"), # remove spacing between facets
        axis.title.x = element_text(size = 8),  
        axis.ticks = element_blank())+
  theme(legend.text=element_text(size=8))+
  theme(legend.title = element_text(size =8))+
  theme(legend.key.size = unit(.3,'cm'))+
  xlab("Shot Number, St Patricks")+
  ylab("")

plot(Fsh)
plot(Msh)
plot(Fsp)
plot(Msp)

grid.arrange(Fsh,Fsp,Msh, Msp, ncol=2, nrow=2, widths= c(14,10), heights= c(8,8)) #arrange plot into a grid (this does not save it only plots to view)

f2.5 <- arrangeGrob(Fsh,Fsp,Msh, Msp, ncol=2, nrow=2, widths= c(5,6), heights= c(7,7)) # creates plot as an object

ggsave("Figures/GonadStage.emf",f2.5) # saves the plot to file :)
ggsave("Figures/GonadStage.jpeg",f2.5) # saves the plot to file :)
```

### Table: The proportion of male and female fish at each ground and their associated weighted averages for fish length (cm) for the 2019 Eastern Zone spawning population.

```{r t2.8}
  
# table for SH proportion by sex
  OR_sHg <- Roughy_2 %>%   
    filter(Ground == "St Helens") %>%
    dplyr::group_by(Ground, sex) %>%
    dplyr::summarise(n = n(), mean = mean(Length)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex = c("F","M"))

#table for SP proportion by sex
  OR_sPg <- Roughy_2 %>%                                        
    filter(Ground == "St Patricks") %>%
    dplyr::group_by(Ground, sex) %>%
    dplyr::summarise(n = n(), mean = mean(Length)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex = c("F","M"))
  
# table for Grounds combined sexes  
 OR_SHSPbg <- Roughy_2 %>%                                      
    dplyr::group_by(Ground) %>%
    dplyr::summarise(n = n(), mean = mean(Length)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex = "M & F")

# table for sex combined grounds 
 OR_sex <- Roughy_2 %>%                                      
    dplyr::group_by(sex) %>%
    dplyr::summarise(n = n(), mean = mean(Length)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(Ground = "SH and SP")
 
# table for survey
 OR_survey <- Roughy_2 %>% 
    #ungroup(sex) %>% 
    dplyr::summarise(n = n(), mean = mean(round(Length))) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex ="all fish",
           Ground = "Survey total")
 
# bind all the tables together, create and rename some coloumns, round off new columns
t2.8 <- bind_rows(OR_sHg, OR_sPg, OR_SHSPbg,OR_sex,OR_survey) %>% 
  mutate("Estimated weight (kg) based on Lyle et al (1991)" 
         = (0.037*(mean)^2.942)/1000) %>% 
  dplyr::rename("No of fish" = n, "Avg length (cm)" 
                = mean, "Spawning Ground"= Ground, "Prop"=Proportion) %>% 
  dplyr::mutate_at(4:6, funs(round(.,2)))  

knitr::kable(tail(t2.8, n = 9), format = "html", caption = "Table 2.8 The proportion of male and female fish at each ground and their associated weighted averages for fish length (cm) for the 2019 Eastern Zone spawning population") %>% 

  kableExtra::kable_styling("striped") %>%

  kableExtra::scroll_box(width = "100%")

write.csv(t2.8,"Figures/Prop_sex_Lylewt.csv")
```

### roughy weight - no weight for 2024

```{r myfig2.6}, out.width='100%', fig.align='center', fig.cap= "Length vs Weight, showing the predicted values (Lyle 1991)"}

Lyle_com <- function(x) {(0.037 * (x) ^ 2.942)}
S2019_fun <- function(x) {(0.056 * (x) ^ 2.8502)}
S2019M_fun <-function(x) {(0.073 * (x) ^ 2.7685)}
S2019F_fun <-function(x) {(0.071 * (x) ^ 2.7945)}

  RoughyWeight<- Roughy%>%
    filter(year==2019) %>%
    drop_na(weight_kg) %>%
    filter(weight_kg != 0) %>% 
    mutate(weight_g = weight_kg *1000) %>%
    mutate(year=as.factor(year))


f2.6<- ggplot(data = RoughyWeight,
       mapping = aes(x  = SLadj_cm,
                     y = weight_g, colour = sex)) +
  scale_colour_manual(values = c("red", "blue"))+
  geom_point(size = .3)+
  scale_x_continuous(breaks = seq(20,50,4),limits = c(20,50))+
  ylab("weight (g)")+
  xlab("Standard length (cm)")+ 
  stat_function(fun = S2019F_fun, colour="red")+
  stat_function(fun = S2019M_fun, colour = "blue") +
  stat_function(fun = Lyle_com, colour ="black")+
  stat_function(fun = S2019_fun, colour ="orange")+
  theme_bw()+
  
  annotate('text', x = 22, y = 3800, size =4,hjust =0,
           label = expression(paste("Weight"["Lyle sex combined"]^{}, 
                                    " = 0037 * Length"^{2.942})))+
  
  annotate('text', x = 22, y = 3500, size = 4, hjust = 0,colour = "orange",
           label = expression(paste("Weight"["2019 sex combined"]^{},
                                    " = 0056 * Length"^{2.850})))+
  
  annotate('text', x = 22, y = 3200, size =4,hjust =0,colour = "red",
           label = expression(paste("Weight"["2019 Females"]^{},
                                    "= 0.071 * Length"^{2.769}))) +
  
  annotate('text', x = 22, y = 2900, size =4, hjust =0, colour = "blue",
           label = expression(paste("Weight"["2019 Males"]^{},
                                    " = 0.073 * Length"^{2.794})))


ggsave("Figures/f2.6.emf",f2.6) # saves the plot to file :)
plot(f2.6)
```

# Age 

Read in age data
```{r}
Age <- read_csv("Results/RoughyEZ_AgeIndex_1986to2019.csv")
#Age_new <- read_csv("Data/ORAge_2019.csv")
Age_new <- read_csv("Data/ORAge_2024_otwgt.csv")
Age_2019 <- read_csv("Data/RoughyAge2019.csv")
```

take a look at the data Age index data
```{r}
Age%>% 
  dplyr::group_by(Sex,Year) %>% 
  dplyr::summarise(mean(O,na.rm = TRUE),mean(Age,na.rm = TRUE))
```
need to add age data from 2019
need to add otolith weight data from 2024
ototlith age data not yet available for 2024

## Add 2019 otolith age data 


remove the 2019 data from the AgeIndex and add the 2019 data that inlcudes the "Age" data.

```{r}
Age_filtered <- Age %>% 
  filter(Year != 2019) %>% 
  dplyr::rename(Month=month) %>% 
  mutate(as.character(Month))
```

take a look at the data Age index data

```{r}
Age_filtered%>% 
  dplyr::group_by(Sex,Year) %>% 
  dplyr::summarise(mean(O,na.rm = TRUE),mean(Age,na.rm = TRUE), n())
```

add the 2019 age dataset

what columns are required
```{r}
colnames(Age_filtered)
colnames(Age_2019)
```

select 2019 fields

```{r}
Age_2019ed <- Age_2019 %>% 
  select(Zone, Ground = AreaCapture, Year, Month, Age= `Age1...14`, Sex,Length_1, O= `Otolith_eight (g)`) %>% 
   mutate(as.character(Month)) %>% 
  mutate(length_cm = Length_1/10)
  #filter(Age >0) no age data currently

Age_2019ed %>% 
  dplyr::group_by(Sex,Year) %>% 
  dplyr::summarise(mean(O,na.rm = TRUE), mean(Age,na.rm = TRUE))
```

bind rows to Age Index, rename Age

```{r}
Age_1 <- Age_filtered %>% 
  bind_rows(Age_2019ed)
  # mutate(Sex_no = Sex) %>% 
  # mutate(Sex= as.factor(Sex)) %>% 
  # mutate(Sex = case_when(Sex == "1"~"M", TRUE~as.character(Sex))) %>%
  # mutate(Sex = case_when(Sex == "2"~"F", TRUE~as.character(Sex))) 
```

check data

```{r}
Age_1 %>% 
  dplyr::group_by(Year,Sex,) %>% 
  dplyr::summarise(mean(O,na.rm = TRUE),mean(Age,na.rm = TRUE), n())
```

Add the 2024 data

```{r}
colnames(Age_1)
colnames(Age_otwgt2024)

```

standardise new data to the Age Index

add columns
```{r}
Age_new<- Age_otwgt2024 %>% 
  add_column(Mark = NA, Sex_no = NA) %>% 
  mutate(as.character(Month))
```
bind data
```{r}
Age_2 <- Age_1 %>% 
  dplyr::bind_rows(Age_new) %>% 
   filter(Sex=="M"|Sex=="F")
```

check data

```{r}
Age_2 %>% 
  dplyr::group_by(Year,Sex,) %>% 
  dplyr::summarise(mean(O,na.rm = TRUE),mean(Age,na.rm = TRUE), n())
 
```

## Plot Age data: Age and otolith weight

```{r}
O_Sum <- summarySE(Age_2, measurevar = "O", groupvars = c("Year","Ground","Sex"),na.rm = TRUE) 
Age_Sum <- summarySE(Age_2, measurevar = "Age", groupvars = c("Year","Ground","Sex"),na.rm = TRUE) 

pd <- position_dodge(0.5) # move position horizontally b/c of overlap

f2.7 <- ggplot(O_Sum, aes(x = Year,y = O, shape = Ground, colour = Sex)) +
  geom_errorbar(aes(ymin = O - ci, ymax= O + ci),
                width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_line(position = pd) +
  xlab("Eastern Zone Spawning Population by year for July")+
  ylab("Average otolith weight (g)")+
  scale_y_continuous(breaks = seq(.05,0.3,.05),limits = c(0,0.28))+
  scale_x_continuous(limits = c(1986,2026), breaks = seq(1986,2026,4))+
  theme_bw()+
  theme(legend.position = c(.25, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'),
        plot.background = element_rect(linetype =  "solid") )


f2.8 <- ggplot(Age_Sum, aes(x = Year,y = Age, shape = Ground, colour = Sex)) +
  geom_errorbar(aes(ymin = Age - ci, ymax= Age + ci),
                width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_line(position = pd) +
  xlab("Eastern Zone Spawning Population by year for July")+
  ylab("Average age (years)")+
  scale_y_continuous(breaks = seq(20,70,10),limits = c(20,70))+
  scale_x_continuous(limits = c(1986,2022), breaks = seq(1986,2022,4))+
  theme_bw()+
  theme(legend.position = c(.25, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'),
        plot.background = element_rect(linetype =  "solid") )

plot(f2.7)
ggsave("Figures/Age_owgt.emf")
ggsave("Figures/Age_owgt.jpg")
plot(f2.8)
ggsave("Figures/Age_years.emf")
ggsave("Figures/Age_owgt.jpg")
```
### Figure 2.9 Otolith weight by year
```{r myfig2.9}

O_Sum <- summarySE(Age_2, measurevar = "O", groupvars = c("Year","Ground","Sex"),na.rm = TRUE) 

# Otolith Weight vs time series
f2.9 <- ggplot(O_Sum, aes(x = Year,y = O, shape = Ground, colour = Sex)) +
  geom_errorbar(aes(ymin = O - ci, ymax= O + ci),
                width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_line(position = pd) +
  xlab("Eastern Zone Spawning Population by year for July")+
  ylab("Average otolith weight (g)")+
  scale_y_continuous(breaks = seq(.1,0.28,.02),limits = c(.1,0.28))+
  scale_x_continuous(limits = c(1986,2022), breaks = seq(1986,2022,4))+
  theme_bw()+
  theme(legend.position = c(.2, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'),
        plot.background = element_rect(linetype =  "solid") )
plot(f2.9)
ggsave("Figures/Age_Owgt.emf")
ggsave("Figures/Age_Owgt.jpg")


write.csv(Age_2,"Results/Otolith_wt.csv")
```
