---
title: 'Orange Roughy Eastern Zone: Biological Summary from the Acoustic Survey 2024'
author: "sutton"
date: "2025-01-28"
output: html_document
---

# Set up

libraries required

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(scales)
library(ggpubr)
library(Rmisc)
library(dplyr)
library(devEMF)
library(tidyr)
library(kableExtra)
library(knitr)
library(cowplot)
library(gridExtra)
library(lubridate)
```

load data
```{r}
# Biological Index
Roughy <- read_csv("Results/OREastern_Tidy_87to2019.csv", col_types = list(ShotWeight_kg = col_double(), mark = col_character()))

Age2019 <- read_csv("Data/RoughyAge2019.csv")
         
# 2024 data
Bio2024 <- read_csv("Data/RoughyBiologicals_EZ_2024.csv")
Shot2024 <- read_csv("Data/RoughyShot_EZ_2024.csv")
discard2024 <- read_csv("Data/Roughy2024_discards.csv")
cc <- read_csv("Data/Roughy2024_catchcomp.csv")
```


manipulate 2024 data

```{r}
Shot2024_ed <- Shot2024 %>% 
  select(Shot,Location,Day, Month, Year, Date,Lengths,Otoliths,Total_catch_weight_kg)

Bio2024_1 <- Bio2024 %>% 
  left_join(Shot2024_ed, by = "Shot") 
```

## Index data - add new data

explore fields
```{r}
colnames(Roughy)
colnames(Bio2024_1)
```
add relevant columns

```{r}

Bio <- Bio2024_1 %>%
  mutate(zone = "Eastern", year = 2024, Area_MS = Location, Survey = Voyage, stage = Gonad_Stage,
         SLadj_cm = Length_cm)

 Bio_1 <- Bio %>% 
   select(zone, year, Area_MS, Survey, Shot, sex = Sex, stage, SLadj_cm, Otoliths, vessel = Vessel, Date= Date.x, ShotWeight_kg = Total_catch_weight_kg)
```

append 2024 data to the index data

```{r}
Roughy <- Roughy %>% 
  bind_rows(Bio_1)

write_csv(Roughy,"Results/Roughy_EZIndex_1987to2024.csv")
```

explore data

```{r}
Roughy %>% 
  filter(year==2024) %>% 
  #group_by(sex) %>% 
  distinct(stage,sex)
```

# Table and Figures

## Summarise bio data 2024

```{r}
BioSummary24 <- Bio2024_1 %>% 
  dplyr::group_by(Shot,Sex) %>% 
  dplyr::summarise(n= n())

BioSummary24 <- spread(BioSummary24, key = Sex,value = n)
```


add other required fields

```{r}
BioSummary24 <- BioSummary24 %>% 
  left_join(Shot2024_ed, by= "Shot") %>% 
  select(Ground = Location, Shot, Date, Total_catch_weight_kg, F, M, Total_Fish_Measured = Lengths, Otoliths_taken = Otoliths)
```

add discards

select relevant fields
```{r}
discard2024 <- discard2024 %>% 
  select(Shot,retained,discarded)
```
 
 join to summary
 
```{r}
BioSummary24 <- BioSummary24 %>% 
  left_join(discard2024, by = "Shot")
```

Write data
```{r}
#write.csv(BioSummary24,"C:/Users/sut109/OneDrive - CSIRO/roughy - working/Figures/t2.4_sumshot_24.csv")
```

```{r}
BioSummary24_ground <- BioSummary24 %>% 
  select(Ground,Shot,retained,discarded) %>% 
  dplyr::group_by(Ground) %>% 
  dplyr::summarise(sum(retained),sum(discarded)) 


```


Detailed discard catch complication

```{r}


Discard_cc <-cc %>% 
  filter(Fate == "D") %>% 
  dplyr::group_by(Species,Shot) %>% 
  dplyr::summarise(TW =sum(`Total Weight (kg)`))

Discard_cc_1 <- Discard_cc %>% 
  left_join(Shot2024, by = "Shot") %>% 
  select(Species,Shot, Ground = Location, TW) %>% 
  drop_na(Ground)


     

```

```{r}


ggplot(Discard_cc_1, aes(x = TW, y = Species, fill = Shot))+
  geom_bar(stat = "identity")+
  facet_wrap(~Ground)


ggplot(Discard_cc_1, aes(x = TW, y = Species, fill = Shot))+
  geom_bar(stat = "identity")


ggplot(Discard_cc_1, aes(x = TW, y = Species, fill = as.factor(Shot))) +
  geom_bar(stat = "identity", color = "black", position = "stack") +
  labs(fill = "Shot Number")  # Customize legend title

ggplot(Discard_cc_1, aes(x = TW, y = Species, fill = as.factor(Shot))) +
  geom_bar(stat = "identity", color = "black", position = "stack") +
  labs(fill = "Shot Number", x = "Tow (TW)", y = "Species Value") +
  theme_minimal() +
  theme(legend.position = "right")

ggplot(Discard_cc_1, aes(x = TW, y = Species, fill = factor(Shot))) +
  geom_bar(stat = "identity", color = "black", position = "stack") +
  labs(fill = "Shot Number")


# Generate a gradient palette (e.g., shades of blue) with N levels
n_shots <- length(unique(Discard_cc_1$Shot))
shot_levels <- sort(unique(Discard_cc_1$Shot))
shot_colors <- seq_gradient_pal("lightblue", "darkblue", "Lab")(seq(0, 1, length.out = n_shots))
names(shot_colors) <- as.character(shot_levels)

# Plot
ggplot(Discard_cc_1, aes(x = TW, y = Species, fill = factor(Shot))) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = shot_colors) +
  labs(fill = "Shot Number")

#######################################

# Example: ensure Shot and Ground are factors
# Make sure Shot and Ground are factors
Discard_cc_1 <- Discard_cc_1 %>%
  mutate(
    Shot = factor(Shot),
    Ground = factor(Ground),
    ShotGround = interaction(Ground, Shot, sep = "_", lex.order = TRUE)
  )

# Get levels
shot_levels <- levels(Discard_cc_1$Shot)
ground_levels <- levels(Discard_cc_1$Ground)

# Create color gradients for each ground
color_vector <- c()

for (g in ground_levels) {
  n_shots <- length(shot_levels)
  
  # Assign blue gradient for St Helens, green for St Patricks
  if (g == "St Helens") {
    gradient_colors <- seq_gradient_pal("lightblue", "darkblue", "Lab")(seq(0, 1, length.out = n_shots))
  } else if (g == "St Patricks") {
    gradient_colors <- seq_gradient_pal("lightgreen", "darkgreen", "Lab")(seq(0, 1, length.out = n_shots))
  } else {
    # Default gray if other grounds exist
    gradient_colors <- seq_gradient_pal("grey90", "grey30", "Lab")(seq(0, 1, length.out = n_shots))
  }
  
  combined_levels <- paste(g, shot_levels, sep = "_")
  names(gradient_colors) <- combined_levels
  color_vector <- c(color_vector, gradient_colors)
}

# Reorder color_vector to match factor levels
color_vector <- color_vector[levels(Discard_cc_1$ShotGround)]

# Plot
ggplot(Discard_cc_1, aes(x = TW, y = Species, fill = ShotGround)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = color_vector, guide = guide_legend(title = "Shot by Ground")) +
  labs(x = "Total Weight (TW)", y = "Species / Taxa Group") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right")


###############################################

Discard_cc_1 <- Discard_cc_1 %>%
  mutate(
    Shot = factor(Shot),
    Ground = factor(Ground),
    ShotGround = interaction(Ground, Shot, sep = "_", lex.order = TRUE)
  )

shot_levels <- levels(Discard_cc_1$Shot)
ground_levels <- levels(Discard_cc_1$Ground)

color_vector <- c()

for (g in ground_levels) {
  n_shots <- length(shot_levels)
  
  if (g == "St Helens") {
    # Stronger blue gradient: from very light cyan to deep navy blue
    gradient_colors <- seq_gradient_pal("#E0FFFF", "#000080", "Lab")(seq(0, 1, length.out = n_shots))
  } else if (g == "St Patricks") {
    # Stronger green gradient: from pale green to deep forest green
    gradient_colors <- seq_gradient_pal("#CCFFCC", "#004d00", "Lab")(seq(0, 1, length.out = n_shots))
  } else {
    gradient_colors <- seq_gradient_pal("grey90", "grey30", "Lab")(seq(0, 1, length.out = n_shots))
  }
  
  combined_levels <- paste(g, shot_levels, sep = "_")
  names(gradient_colors) <- combined_levels
  color_vector <- c(color_vector, gradient_colors)
}

color_vector <- color_vector[levels(Discard_cc_1$ShotGround)]

ggplot(Discard_cc_1, aes(x = TW, y = Species, fill = ShotGround)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = color_vector, guide = guide_legend(title = "Shot by Ground")) +
  labs(x = "Tow (TW)", y = "Species Value") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right")

##################################################

Discard_cc_1 <- Discard_cc_1 %>%
  mutate(
    Shot = factor(Shot),
    Ground = factor(Ground),
    ShotGround = interaction(Ground, Shot, sep = "_", lex.order = TRUE)
  )

shot_levels <- levels(Discard_cc_1$Shot)
ground_levels <- levels(Discard_cc_1$Ground)

# Function to generate clearly spaced colors by manual interpolation
generate_contrast_colors <- function(start_color, end_color, n) {
  # Generate colors evenly spaced in LAB color space for better perceptual difference
  colors <- grDevices::colorRampPalette(c(start_color, end_color))(n)
  return(colors)
}

color_vector <- c()

for (g in ground_levels) {
  n_shots <- length(shot_levels)
  
  if (g == "St Helens") {
    # Strong distinct blues - bright cyan to dark royal blue with bigger perceptual jumps
    gradient_colors <- generate_contrast_colors("#00FFFF", "#00008B", n_shots)
  } else if (g == "St Patricks") {
    # Strong distinct greens - bright lime to dark green
    gradient_colors <- generate_contrast_colors("#32CD32", "#004400", n_shots)
  } else {
    # Neutral grays
    gradient_colors <- generate_contrast_colors("#CCCCCC", "#333333", n_shots)
  }
  
  combined_levels <- paste(g, shot_levels, sep = "_")
  names(gradient_colors) <- combined_levels
  color_vector <- c(color_vector, gradient_colors)
}

color_vector <- color_vector[levels(Discard_cc_1$ShotGround)]

ggplot(Discard_cc_1, aes(x = TW, y = Species, fill = ShotGround)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = color_vector, guide = guide_legend(title = "Shot by Ground")) +
  labs(x = "Tow (TW)", y = "Species Value") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right")

#######################################################

Discard_cc_1 <- Discard_cc_1 %>%
  mutate(
    Shot = factor(Shot),
    Ground = factor(Ground),
    ShotGround = interaction(Ground, Shot, sep = "_", lex.order = TRUE)
  )

shot_levels <- levels(Discard_cc_1$Shot)
ground_levels <- levels(Discard_cc_1$Ground)

# Function to generate colors with bigger perceptual jumps by limiting number of shades and picking evenly spaced ones
generate_stepped_colors <- function(start_color, end_color, n_steps, n_needed) {
  # Generate a big palette first (n_steps), then pick n_needed colors evenly spaced out
  big_palette <- grDevices::colorRampPalette(c(start_color, end_color))(n_steps)
  # Pick evenly spaced colors from big palette
  indices <- round(seq(1, n_steps, length.out = n_needed))
  stepped_colors <- big_palette[indices]
  return(stepped_colors)
}

color_vector <- c()

for (g in ground_levels) {
  n_shots <- length(shot_levels)
  
  if (g == "St Helens") {
    # Generate a palette with bigger jumps (e.g. 7 steps, pick 5 for 5 shots)
    gradient_colors <- generate_stepped_colors("#00BFFF", "#00008B", n_steps = 7, n_needed = n_shots)
  } else if (g == "St Patricks") {
    gradient_colors <- generate_stepped_colors("#7CFC00", "#004400", n_steps = 7, n_needed = n_shots)
  } else {
    gradient_colors <- generate_stepped_colors("#CCCCCC", "#333333", n_steps = 7, n_needed = n_shots)
  }
  
  combined_levels <- paste(g, shot_levels, sep = "_")
  names(gradient_colors) <- combined_levels
  color_vector <- c(color_vector, gradient_colors)
}

color_vector <- color_vector[levels(Discard_cc_1$ShotGround)]

ggplot(Discard_cc_1, aes(x = TW, y = Species, fill = ShotGround)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = color_vector, guide = guide_legend(title = "Shot by Ground")) +
  labs(x = "Tow (TW)", y = "Species Value") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right")


##############################

Discard_cc_1 <- Discard_cc_1 %>%
  mutate(
    Shot = factor(Shot),
    Ground = factor(Ground),
    ShotGround = interaction(Ground, Shot, sep = "_", lex.order = TRUE)
  )

shot_levels <- levels(Discard_cc_1$Shot)
ground_levels <- levels(Discard_cc_1$Ground)

# Manually define very distinct blue shades for St Helens
st_helens_colors <- c(
  "#003f5c", # very dark blue
  "#2f4b7c",
  "#665191",
  "#a05195",
  "#d45087",
  "#f95d6a",
  "#ff7c43"
)

# Manually define very distinct green shades for St Patricks
st_patricks_colors <- c(
  "#004d00", # very dark green
  "#237a23",
  "#47a347",
  "#6fc46f",
  "#98d398",
  "#c0e4c0",
  "#e6f1e6"
)

# Adjust length to number of shots
st_helens_colors <- st_helens_colors[seq_len(length(shot_levels))]
st_patricks_colors <- st_patricks_colors[seq_len(length(shot_levels))]

color_vector <- c()

for (g in ground_levels) {
  combined_levels <- paste(g, shot_levels, sep = "_")
  
  if (g == "St Helens") {
    names(st_helens_colors) <- combined_levels
    color_vector <- c(color_vector, st_helens_colors)
  } else if (g == "St Patricks") {
    names(st_patricks_colors) <- combined_levels
    color_vector <- c(color_vector, st_patricks_colors)
  } else {
    # fallback gray scale
    fallback_colors <- grDevices::colorRampPalette(c("grey90", "grey30"))(length(shot_levels))
    names(fallback_colors) <- combined_levels
    color_vector <- c(color_vector, fallback_colors)
  }
}

color_vector <- color_vector[levels(Discard_cc_1$ShotGround)]

ggplot(Discard_cc_1, aes(x = TW, y = Species, fill = ShotGround)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = color_vector, guide = guide_legend(title = "Shot by Ground")) +
  labs(x = "Tow (TW)", y = "Species Value") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right")

###########


Discard_cc_1 <- Discard_cc_1 %>%
  mutate(
    Shot = factor(Shot),
    Ground = factor(Ground),
    ShotGround = interaction(Ground, Shot, sep = "_", lex.order = TRUE)
  )

shot_levels <- levels(Discard_cc_1$Shot)
ground_levels <- levels(Discard_cc_1$Ground)

stopifnot(all(ground_levels %in% c("St Helens", "St Patricks")))

# Manually define strong distinct colors for St Helens (blue hues)
st_helens_colors <- c(
  "#003f5c", "#2f4b7c", "#665191", "#a05195", "#d45087", "#f95d6a", "#ff7c43"
)[seq_along(shot_levels)]

# Manually define strong distinct colors for St Patricks (green hues)
st_patricks_colors <- c(
  "#004d00", "#237a23", "#47a347", "#6fc46f", "#98d398", "#c0e4c0", "#e6f1e6"
)[seq_along(shot_levels)]

color_vector <- c()

for (g in ground_levels) {
  combined_levels <- paste(g, shot_levels, sep = "_")
  if (g == "St Helens") {
    names(st_helens_colors) <- combined_levels
    color_vector <- c(color_vector, st_helens_colors)
  } else if (g == "St Patricks") {
    names(st_patricks_colors) <- combined_levels
    color_vector <- c(color_vector, st_patricks_colors)
  }
}

# Reorder colors to match ShotGround factor levels exactly
color_vector <- color_vector[levels(Discard_cc_1$ShotGround)]

ggplot(Discard_cc_1, aes(x = TW, y = Species, fill = ShotGround)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = color_vector, guide = guide_legend(title = "Shot by Ground")) +
  labs(x = "Tow (TW)", y = "Species Value") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right")

###########################

Discard_cc_1 <- Discard_cc_1 %>%
  mutate(
    Shot = factor(Shot),
    Ground = factor(Ground),
    ShotGround = interaction(Ground, Shot, sep = "_", lex.order = TRUE)
  )

# Get levels of ShotGround (all combinations in your data)
all_levels <- levels(Discard_cc_1$ShotGround)

# Extract unique Grounds and Shots from those levels
# (splitting by underscore to get Ground and Shot parts)
library(stringr)
ground_shot_df <- data.frame(
  ShotGround = all_levels,
  Ground = str_extract(all_levels, "^[^_]+"),
  Shot = str_extract(all_levels, "(?<=_).*")
)

# Define strong distinct blue colors for St Helens
st_helens_colors <- c(
  "darkblue","#003f5c", "#2f4b7c", "lightblue","#665191", "#a05195", "#d45087", "#f95d6a", "#ff7c43","orange","yellow","beige"
)

# Define strong distinct green colors for St Patricks
st_patricks_colors <- c(
  "#004d00", "#237a23", "#47a347", "#6fc46f", "#98d398", "#c0e4c0", "#e6f1e6", "white"
)

# Get unique Shots (to know how many colors needed)
unique_shots <- unique(ground_shot_df$Shot)

# Assign colors per Ground and Shot
color_list <- sapply(all_levels, function(level) {
  parts <- strsplit(level, "_")[[1]]
  ground <- parts[1]
  shot <- parts[2]
  shot_index <- which(unique_shots == shot)
  
  if (ground == "St Helens") {
    # Use color based on shot index; recycle if shots > color length
    st_helens_colors[(shot_index - 1) %% length(st_helens_colors) + 1]
  } else if (ground == "St Patricks") {
    st_patricks_colors[(shot_index - 1) %% length(st_patricks_colors) + 1]
  } else {
    "#999999"  # fallback grey for unexpected Grounds
  }
})

names(color_list) <- all_levels

# Plot with manual fill scale
ggplot(Discard_cc_1, aes(x = TW, y = Species, fill = ShotGround)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = color_list, guide = guide_legend(title = "Shot by Ground")) +
  labs(x = "Tow (TW)", y = "Species Value") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right")


###############################

# Assuming Discard_cc_1 has columns: TW, Species, Shot, Ground

Discard_cc_1 <- Discard_cc_1 %>%
  mutate(
    ShotGround_char = paste(Ground, Shot_num, sep = "_")
  )

# Create ordered levels
levels_ordered <- Discard_cc_1 %>%
  distinct(Ground, Shot_num) %>%
  arrange(Ground, Shot_num) %>%
  mutate(ShotGround = paste(Ground, Shot_num, sep = "_")) %>%
  pull(ShotGround) %>%
  unique()

# Convert ShotGround_char to factor with ordered levels
Discard_cc_1$ShotGround <- factor(Discard_cc_1$ShotGround_char, levels = levels_ordered)

# 4. Define manual colors (make sure length matches number of Shots)
st_helens_colors <- c(
  "darkblue", "#003f5c", "#2f4b7c", "lightblue", "#665191", "#a05195",
  "#d45087", "#f95d6a", "#ff7c43", "orange", "yellow", "beige"
)

st_patricks_colors <- c(
  "#004d00", "#237a23", "#47a347", "#6fc46f", "#98d398", "#c0e4c0",
  "#e6f1e6", "white"
)

num_shots <- length(levels(Discard_cc_1$Shot))
st_helens_colors <- st_helens_colors[1:num_shots]
st_patricks_colors <- st_patricks_colors[1:num_shots]

# 5. Assign colors to each ShotGround level
all_levels <- levels(Discard_cc_1$ShotGround)

color_vector <- sapply(all_levels, function(level) {
  parts <- strsplit(level, "_")[[1]]
  ground <- parts[1]
  shot_num <- as.numeric(parts[2])
  shot_index <- which(levels(Discard_cc_1$Shot) == as.character(shot_num))
  
  if (ground == "St Helens") {
    st_helens_colors[shot_index]
  } else if (ground == "St Patricks") {
    st_patricks_colors[shot_index]
  } else {
    "#999999" # fallback grey for unexpected Grounds
  }
})

names(color_vector) <- all_levels

# 6. Plot
ggplot(Discard_cc_1, aes(x = TW, y = Species, fill = ShotGround)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = color_vector, guide = guide_legend(title = "Shot by Ground")) +
  labs(x = "Tow (TW)", y = "Species Value") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right")

################
# Your original data: Discard_cc_1

# Step 1: Get unique ShotGround levels and order them as you want
levels_ordered <- sort(unique(Discard_cc_1$ShotGround))

# Step 2: Split levels by ground
st_helens_shots <- levels_ordered[grepl("St Helens", levels_ordered)]
st_patricks_shots <- levels_ordered[grepl("St Patricks", levels_ordered)]

# Step 3: Your defined colors (make sure lengths match counts exactly)
# st_helens_colors <- c(
#   "darkblue", "#003f5c", "#2f4b7c", "lightblue", "#665191", 
#   "#a05195", "#d45087", "#f95d6a", "#ff7c43", "orange", "yellow", "beige"
# )

st_helens_colors <- c(
  "#800000",  # Maroon
  "#B22000",  # Firebrick
  "#FF0000",  # Red
    "pink",
  "#FF4500",  # OrangeRed
  "#FF6347",  # Tomato
  "#FF7F50",  # Coral

  "#FFA500",  # Orange
  "yellow",
  "#FFD700",  # Gold
  "beige",
  "white"
)

# st_patricks_colors <- c(
#   "#004d00", "#237a23", "#47a347", "#6fc46f", "#98d398", 
#   "#c0e4c0", "#e6f1e6", "white"
# )

st_patricks_colors <- c(
  "#001f3f",  # Navy
  "#003f5c",  # Dark Blue
  "#2f4b7c",  # Indigo Blue
  "#3f6caa",  # Medium Blue
  "#4682B4",  # Steel Blue
  "#5dade2",  # Sky Blue
  "#87cefa",  # Light Sky Blue
  "#d0e7f9"   # Very Light Blue
)

# Step 4: Create named vector of colors matching factor levels
colors_named <- c(
  setNames(st_helens_colors, st_helens_shots),
  setNames(st_patricks_colors, st_patricks_shots)
)

# Step 5: Set ShotGround factor levels in your data to ensure order
Discard_cc_1$ShotGround <- factor(Discard_cc_1$ShotGround, levels = levels_ordered)

# Step 6: Plot
ggplot(Discard_cc_1, aes(x = TW, y = Species, fill = ShotGround)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = colors_named) +
  theme_minimal() +
  labs(fill = "Shot and Ground") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
unique_shotground_values <- unique(Discard_cc_1$ShotGround_char)
print("Unique ShotGround values in data:")
print(unique_shotground_values)

print("Levels ordered:")
print(levels_ordered)

# Which ShotGround values are NOT in levels_ordered?
mismatched_levels <- setdiff(unique_shotground_values, levels_ordered)
print("Values in ShotGround NOT in levels_ordered:")
print(mismatched_levels)

# Which levels_ordered are NOT in ShotGround?
missing_in_data <- setdiff(levels_ordered, unique_shotground_values)
print("Levels in levels_ordered NOT in ShotGround:")
print(missing_in_data)
```

---
### summary table with additional 2024 data

```{r mytable, out.width='100%', echo = T}


#remame and select columns
Roughy_1 <- Roughy %>% 
  mutate(Length =round(SLadj_cm)) %>% #round values for length in cm 2019 was to nearest mm
  mutate(Ground = Area_MS) %>% 
  drop_na(Ground) %>% 
  #drop_na(Length) %>% 
  drop_na(sex) %>% 
  select(Ground, year, Shot, Date,sex,stage, SLadj_cm,Length,ShotWeight_kg, weight_kg, Otoliths)


Roughy_DataSummary <- Roughy_1 %>% 
  dplyr::group_by(year, Ground,sex) %>% 
  dplyr::summarise(n= n())

Roughy_DataSummary <- spread(Roughy_DataSummary, key = year,value = n)


knitr::kable(tail(Roughy_DataSummary, n = 5),caption = "Summary data: Number of Fish measured for orange roughy eastern zone spawning surveys by, year, Area, sex") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width = "100%")
```

```{r}
  
Roughyminmax <- Roughy_1 %>% 
  dplyr::group_by(year,Ground,sex) %>% 
  dplyr::summarise(round(mean(Length),digits = 2))

Roughyminmax
```

### Length - index
Average Standard Length (cm) of orange roughy Eastern Zone spawning population, showing males and females at St Helens and St Patricks for 1987 to 2024. Error bars are 95% confidence limits (updated from Kloser et al. 2016)
```{r myfig2.1}
#average SL by Area and sex

library(Rmisc)
library(ggplot2)

roughy_sum <- summarySE(Roughy_1, measurevar = "Length", groupvars = c("year","Ground","sex")) 


  
pd <- position_dodge(0.6) # move position horizontally b/c of overlap

f2.1 <- ggplot(roughy_sum, aes(x = year,y = Length, shape = Ground, colour = sex)) +
  geom_errorbar(aes(ymin = Length - ci, ymax= Length + ci),
                width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_line(position = pd) +
  xlab("Eastern Zone Spawning Population by year for July")+
  ylab("Average Standard Length (cm)")+
  scale_y_continuous(breaks = seq(27,38,1),limits = c(27,38))+
  scale_x_continuous(limits = c(1986,2026), breaks = seq(1986,2026,4))+
  theme_bw()+
  theme(legend.position = c(.95, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'),
         plot.background = element_rect(linetype =  "solid") )

  
plot(f2.1)
ggsave("Figures/LF_index.emf")
ggsave("Figures/LF_index.jpg")
```
## Roughy series - adding weigth estimates to the data

* Lyle et al. 1991 Pg 63. 

Combined length – weight regression. W = 0.037 * SL ^ 2.952.

Female length – weight regression. W = 0.0351 * SL ^ 2.970

Male length – weight regression. W = 0.0383 * SL ^ 2.942  


```{r}

Roughy_1 <-Roughy_1 %>% 
  #filter(year==2019) %>%
   #drop_na(weight_kg) %>%
   #filter(weight_kg != 0) %>% 
   mutate(year=as.factor(year)) %>% 
   mutate(Wgt_derived = (0.037 * SLadj_cm ^ 2.952)/1000) %>%
   mutate(Wgt_derived_MF = if_else(sex == "M",(0.0383 * SLadj_cm ^ 2.942)/1000, 
                                   (0.0351 * SLadj_cm ^ 2.970)/1000, NA_real_)) %>% 
  mutate(MvsD = weight_kg / Wgt_derived_MF) %>%      
  mutate(percentchange = ((weight_kg - Wgt_derived_MF)/ weight_kg)) %>% 
  mutate(wgt_diff = Wgt_derived_MF - weight_kg) %>% 
  arrange(sex)
```

## DATA for 2024
```{r}
Roughy_2 <- Roughy_1 %>% 
filter(year == 2024)
```

### length data table
exploring the length data
```{r}

Roughy_2_length_GroundbySex <- summarySE(Roughy_2, measurevar = "Length", 
                                      groupvars = c("Ground","sex"))
Roughy_2_length_GroundbySex

Roughy_2_length_Ground <- summarySE(Roughy_2, measurevar = "Length", 
                                      groupvars = c("Ground"))
Roughy_2_length_Ground

Roughy_3 <-Roughy_2 %>% 
  filter(SLadj_cm >= 42)

summarySE(Roughy_3, measurevar = "Length", 
                                      groupvars = c("Ground","sex"))
```

### summaries

#### Table 2.4 (needs updating for future surveys to include shot without measures in "OR.R")
```{r table 2.4, echo=FALSE}
#need to update code to include shot weights for shots where fish were not measured **
# fitler data for 2019


Roughy_2_sum <- Roughy_2 %>% 
  dplyr::group_by(Ground,Shot,Date,ShotWeight_kg,sex) %>% 
  dplyr::summarise(num_rows = n()) %>% 
  spread(sex,num_rows) %>% 
  mutate(`Total` = F + M) %>% 
  mutate(Ratio_F = F/`Total`) 

# creating ground/ and survey totals for:

#St Helens
TSH <- Roughy_2_sum %>% 
  filter(Ground == "St Helens") %>% 
  dplyr::group_by(Ground) %>% 
  dplyr::summarise(ShotWeight_kg =sum(ShotWeight_kg), F= sum(F), M = sum(M), Total =sum(Total)) %>% 
  mutate(Ground = "St Helens Total")

#St Patricks
TSP <- Roughy_2_sum %>% 
  filter(Ground == "St Patricks") %>% 
  dplyr::group_by(Ground) %>% 
  dplyr::summarise(ShotWeight_kg =sum(ShotWeight_kg, na.rm = TRUE), F= sum(F), M = sum(M), Total =sum(Total)) %>% 
  mutate(Ground = "St Patricks Total")

# totals for SH and SP
TSHSP <- Roughy_2_sum %>% 
  summarise(ShotWeight_kg = sum(ShotWeight_kg, na.rm = TRUE), F= sum(F), M = sum(M), Total =sum(Total)) %>% 
  mutate(Ground = "Survey Total")

#bind the shot and subtotals and total
t2.4 <- bind_rows(Roughy_2_sum, TSH, TSP,TSHSP) %>%   # put all together for totals
  dplyr::rename("Shot Weight (kg)" = ShotWeight_kg, "Total Nos Fish" = Total)
  

write.csv(t2.4,"Figures/t2.4.csv")

knitr::kable(head(t2.4, n=20), format = "html", caption = "Table 2.4 Summary of shot dates and weights for St Helens and St Patricks, Eastern Zone, 2019") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width  = "100%")
```

## Data by shot: Length and derived

```{r}
Roughy_2_sum_L <- Roughy_2 %>% 
  dplyr::group_by(Ground,Shot,sex) %>% 
  dplyr::summarise(avgL = mean(Length)) %>% 
  spread(sex,avgL) %>% 
  mutate(F_Lgt =round(F,2), M_Lgt =round(M,2)) %>% 
  select(-F,-M)
  
                     
Roughy_2_sum_W <- Roughy_2 %>% 
  dplyr::group_by(Ground,Shot,sex) %>% 
  dplyr::summarise(avgW = mean(Wgt_derived)) %>%
  spread(sex,avgW) %>% 
  mutate(F_Dw =round(F,2), M_DW =round(M,2)) %>% 
  ungroup() %>% 
  select(-F,-M,-Ground,-Shot)
                  
Roughy_2_sum_WFM <- Roughy_2 %>% 
  dplyr::group_by(Ground,Shot,sex) %>% 
  dplyr::summarise(avgWFM = mean(Wgt_derived_MF)) %>%
  spread(sex,avgWFM) %>% 
  mutate(F_wFM =round(F,2), M_wFM =round(M,2)) %>% 
  ungroup() %>% 
  select(-F,-M,-Ground,-Shot)
                   
                   #avgwgt = mean(Wgt_derived),avgwgtMF = mean(Wgt_derived_MF)) 
shotdata_LWW <- bind_cols(Roughy_2_sum_L,Roughy_2_sum_W,Roughy_2_sum_WFM)
  
  # t2.4 <- bind_rows(Roughy_2_sum, TSH, TSP,TSHSP) %>%   # put all together for totals
  # dplyr::rename("Shot Weight (kg)" = ShotWeight_kg, "Total Nos Fish" = Total)



```

### Standard Length by sex and ground
```{r myfig2.2}
OR_LF <- Roughy_2 %>% # This is the ONE
  group_by(Ground, sex) %>% 
  add_tally() %>% 
  group_by(Ground, sex, Length) %>% 
  dplyr::summarise(prop = n() / n[1])

f2.2 <- ggplot(OR_LF, aes(Length, prop, fill = sex)) +
  geom_col(position = "identity", alpha = 0.6) +
  scale_x_continuous(breaks = seq(28,46,2),limits = c(28,46))+
  facet_wrap(~Ground)+
   theme_bw()+
  ylab("proportion")+
  xlab("Length (cm)")+
  theme(legend.position = "bottom")
plot(f2.2)
ggsave(filename = "Figures/LF2024.emf")
ggsave(filename = "Figures/LF2024.jpg")
```


### Cummunlative Frequency
```{r myfig2.3}

#using stat_ecdf to calculate the cummulative frequency

f2.3 <- ggplot(Roughy_2,aes(x=Length, colour= sex)) +
   stat_ecdf(geom = "smooth")+
  scale_x_continuous(breaks = seq(28,44,2),limits = c(28,44))+
  scale_y_continuous(breaks = seq(0,1,0.2))+
  facet_wrap(~Ground)+
  ylab("cummulative frequency")+
  xlab("Length (cm)")+
  theme_bw()+
  theme(legend.position = "bottom")
plot(f2.3)
ggsave(filename = "Figures/cumLF2024.emf")
```

## Mode and Numbers of fish by Year, Ground, Sex 

```{r Table2.5}
#need to group by Ground and sex togehter so need to make a combined variable
Roughy_3 <- Roughy_1 %>% 
  mutate(GbyS= paste(Ground,sex,sep = "_")) %>%  #new variable to combine Ground and sex 
  group_by(GbyS) %>% 
  mutate(sumsum = cumsum(Length))


#create function to find the mode
Mode <- function(x) {
    if (is.numeric(x)) {
        x_table <- table(x)
        return(as.numeric(names(x_table)[which.max(x_table)]))
        }
}

#calcuate the mode
Roughy_mode <- Roughy_3 %>% 
  dplyr::group_by(year,GbyS) %>%
  dplyr::summarise(Mode = Mode(x = Length))%>%
  spread(GbyS, Mode) 
#calcuate the  counts
Roughy_nos <- Roughy_3 %>% 
  dplyr::group_by(year, GbyS) %>%
  dplyr::summarise(Numbers = n())%>%
  spread(GbyS, Numbers)

#bind the mode and counts
R_Mode_Count <-  cbind(Roughy_mode,Roughy_nos)%>%
  select(-6)#remove the year1 colomn

t2.5<-  R_Mode_Count_order <- R_Mode_Count[,c(1,2,6,3,7,4,8,5,9)]  # order the columns
  
  write.csv(t2.5,"Figures/Lmode_index.csv")
  #note that this table is also saved as t2.4.doc in "OR Tables.Rmd"
 
knitr::kable(R_Mode_Count_order, col.names = c("", "Nos. Fish","Mode", "Nos. Fish","Mode", "Nos. Fish","Mode", "Nos. Fish","Mode"), format = "html", caption = " Figure 2.5. Number of Fish measured and mode of standard lengths (cm) for male and femailes measured at St Helens and St Particks from 1987 to 2019") %>% 
  kableExtra::kable_styling("striped") %>%
  add_header_above(c(" ", "Female" = 2, "Male" = 2, "Female" = 2, "Male"=2)) %>%
    add_header_above(c(" ", "St Helens" = 4, "St Patricks" = 4)) %>%
    kableExtra::scroll_box(width = "100%")

```

## Average Length by Shot
### Figure 2.4
```{r myfig2.4, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}

library(gridExtra) #for function grid.arrange

#average SL by Area and sex
Roughy_2_lengthsum <- summarySE(Roughy_2, measurevar = "Length", groupvars = c("Shot","Date","Ground","sex")) %>% 

  mutate(Shot = as.character(Shot)) %>% 
  arrange(Length)

pd <- position_dodge(0) # in case need to experiment with operlap of values - this was not required so set 0

SH <- ggplot(Roughy_2_lengthsum %>% 
               filter(Ground == "St Helens"),
             aes(x = reorder(Shot, sort((Shot))),y = Length, colour = sex)) +
    geom_errorbar(aes(ymin = Length - se, ymax= Length + ci),
                width =.7, position = pd) +
   geom_point( position = pd, size = 2, width = 1)+
# facet_grid(.~Date, 
 #           scales = "free_x")+ # ony dates with values are included
  theme_bw()+
  theme(panel.spacing = unit(0, "mm"), # remove spacing between facets
        
        strip.background = element_rect(size = 0.1), # match default line size of theme_classic)
        strip.text.x = element_text(size =8))  +       # change font size
  theme(legend.position = c(.97, .05), #position legend bottom right
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'))+
   scale_y_continuous(breaks = seq(32,40,1),limits = c(32,40))+ 
  xlab("Shot Number, St Helens")+
  ylab("Average Standard Length (cm)")

SP <- ggplot(Roughy_2_lengthsum %>% 
               filter(Ground =="St Patricks"),
             aes(x = reorder(Shot, sort((Shot))),y = Length, colour = sex)) +
    geom_errorbar(aes(ymin = Length - ci, ymax= Length + ci),
                width =.7, position = pd) +
   geom_point( position = pd, size = 2)+
    #facet_grid(.~Date, scales = "free_x")+
  theme_bw()+
  theme(legend.position = c(.97, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'))+
     theme(panel.spacing = unit(0, "mm"),                      
        strip.background = element_rect(size = 0.1),
        strip.text.x = element_text(size =8))  +
  scale_y_continuous(breaks = seq(32,40,1),limits = c(32,40))+
  xlab("Shot Number, st Patricks")+
  ylab("") #removes y axis


grid.arrange(SH,SP,ncol=2, widths= c(5,4)) #arrange plot into a grid (this does not save it only plots to view)

f2.4 <- arrangeGrob(SH, SP,ncol = 2, widths = c(5,4)) # creates plot as an object

                 
ggsave("Figures/shotvariablity.emf",f2.4) # saves the plot to file :)
```

### Gonad Stages
```{r f2.5}
#need to group the frame by sex 
Roughy_4 <- Roughy_1%>%
  filter(year==2024) %>% 
  mutate(stage = as.character(stage))%>% 
 # mutate(Shot = as.character(Shot)) %>% 
  mutate(Date= as.Date(Date,"%d/%m")) %>% 
  #select(-Date) %>% 
  group_by(sex)

OR2024_sumFH <- Roughy_2_lengthsum %>% 
  filter(Ground == "St Helens", sex == "F")
OR2024_sumMH <- Roughy_2_lengthsum %>% 
  filter(Ground == "St Helens", sex == "M")
OR2024_sumFP <- Roughy_2_lengthsum %>% 
  filter(Ground == "St Patricks", sex == "F")
OR2024_sumMP <- Roughy_2_lengthsum %>% 
  filter(Ground == "St Patricks", sex == "M")

# Fsh <- ggplot(Roughy_4 %>% 
#   filter(Ground == "St Helens", sex == "F"),
#   aes(x = Date, fill = stage))+
#   geom_bar(position = "fill")+
#   scale_fill_manual(values = c("yellow","orange","red","grey"))+
#   theme(axis.text=element_text(size=9))+
#  facet_grid(.~Shot, scales = "free_x", switch ="x")+
#   theme(strip.text.x = element_text(size =8),
#         panel.spacing = unit(0,"mm"))+ # remove spacing between facets
#   theme(legend.position = "none")+
#   theme(axis.title.y = element_text(size = 11))+
#    theme(axis.title.x = element_text(size = 11),  
#         axis.ticks = element_blank()) +
#   ylab("Proportion female fish")+
#   xlab("")

Fsh <- ggplot(Roughy_4 %>% 
  filter(Ground == "St Helens", sex == "F"),
  aes(x = Date, fill = stage))+
  geom_bar(position = "fill")+
  scale_fill_manual(values = c("yellow","orange","red","grey"))+
  theme(axis.text= element_blank())+
 facet_grid(.~Shot, scales = "free_x", switch ="x")+
  theme(strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"))+ # remove spacing between facets
  #theme(legend.position = "none")+
  theme(legend.text=element_text(size=8))+
  theme(legend.title = element_text(size =8))+
  theme(legend.key.size = unit(.3,'cm'))+
  theme(axis.title.y = element_text(size = 8))+
   theme(axis.title.x = element_text(size = 8),  
        axis.ticks = element_blank()) +
  ylab("Proportion female fish")+
  xlab("")

Msh <- ggplot(Roughy_4 %>% 
  filter(sex == "M",Ground =="St Helens"),
  aes(x = Date, fill = stage))+
  geom_bar(position = "fill")+
  scale_fill_manual(values = c("yellow","orange","red","grey","black"))+
  theme(axis.text= element_blank())+
 facet_grid(.~Shot, scales = "free_x", switch ="x")+
    theme(strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"), # remove spacing between facets
        axis.title.x = element_text(size = 8),  
        axis.ticks = element_blank())+
  theme(legend.text=element_text(size=8))+
  theme(legend.title = element_text(size =8))+
  theme(legend.key.size = unit(.3,'cm'))+
    theme(axis.title.y = element_text(size = 8))+
     theme(axis.title.x = element_text(size = 8),  
        axis.ticks = element_blank()) +
  xlab("Shot Number, St Helens")+
  ylab("Proportion male fish")

Fsp <- ggplot(Roughy_4 %>% 
  filter(sex == "F",Ground =="St Patricks"),
  aes(x = Date, fill = stage, ))+
  geom_bar(position = "fill")+
  labs(fill = "stage")+
  scale_fill_manual(values = c("yellow","orange","red","grey"))+
 theme(axis.text= element_blank())+
  facet_grid(.~Shot, scales = "free_x", switch ="x")+
  theme(strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"))+ # remove spacing between facets
   theme(axis.title.x = element_text(size = 8),  
        axis.ticks = element_blank()) +
  theme(legend.text=element_text(size=8))+
  theme(legend.title = element_text(size =8))+
  theme(legend.key.size = unit(.3,'cm'))+
   ylab("")+
   xlab("")

Msp <- ggplot(Roughy_4 %>% 
  filter(sex == "M",Ground =="St Patricks"),
  aes(x = Date, fill = stage))+
  geom_bar(position = "fill")+
    labs(fill = "stage")+
    scale_fill_manual(values = c("orange","red","grey","black"))+
    theme(axis.text= element_blank())+
  facet_grid(.~Shot, scales = "free_x", switch ="x")+
    theme(strip.text.x = element_text(size =8),
        panel.spacing = unit(0,"mm"), # remove spacing between facets
        axis.title.x = element_text(size = 8),  
        axis.ticks = element_blank())+
  theme(legend.text=element_text(size=8))+
  theme(legend.title = element_text(size =8))+
  theme(legend.key.size = unit(.3,'cm'))+
  xlab("Shot Number, St Patricks")+
  ylab("")

plot(Fsh)
plot(Msh)
plot(Fsp)
plot(Msp)

grid.arrange(Fsh,Fsp,Msh, Msp, ncol=2, nrow=2, widths= c(14,10), heights= c(8,8)) #arrange plot into a grid (this does not save it only plots to view)

f2.5 <- arrangeGrob(Fsh,Fsp,Msh, Msp, ncol=2, nrow=2, widths= c(11,8.5), heights= c(5,5)) # creates plot as an object


ggsave("Figures/GonadStage.emf",f2.5) # saves the plot to file :)
ggsave("Figures/GonadStage.jpeg",f2.5) # saves the plot to file :)
```
### gonad stage summary
exploring gonad data
```{r}
Roughy_4 %>% 
  filter(Ground =="St Patricks") %>% 
  add_tally() %>% 
  dplyr::group_by(sex,stage) %>% 
 dplyr::summarise(prop = n()/ n[1]) 


Roughy_4 %>% 
  add_tally() %>% 
  dplyr::group_by(sex,stage) %>% 
 dplyr::summarise(n())
  # group_by(Ground, sex) %>% 
  # add_tally() %>% 
  # group_by(Ground, sex, Length) %>% 
  # dplyr::summarise(prop = n() / n[1])
```


### Table: The proportion of male and female fish at each ground and their associated weighted averages for fish length (cm) for the 2024 Eastern Zone spawning population.

```{r t2.8}
  
# table for SH proportion by sex
  OR_sHg <- Roughy_2 %>%   
    filter(Ground == "St Helens") %>%
    dplyr::group_by(Ground, sex) %>%
    dplyr::summarise(n = n(), mean = mean(Length), estimated_wgt = mean(Wgt_derived), est_wgt_FM = mean(Wgt_derived_MF)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex = c("F","M"))
   

#table for SP proportion by sex
  OR_sPg <- Roughy_2 %>%                                        
    filter(Ground == "St Patricks") %>%
    dplyr::group_by(Ground, sex) %>%
    dplyr::summarise(n = n(), mean = mean(Length),estimated_wgt = mean(Wgt_derived), est_wgt_FM = mean(Wgt_derived_MF)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex = c("F","M"))
  
# table for Grounds combined sexes  
 OR_SHSPbg <- Roughy_2 %>%                                      
    dplyr::group_by(Ground) %>%
    dplyr::summarise(n = n(), mean = mean(Length),estimated_wgt = mean(Wgt_derived), est_wgt_FM = mean(Wgt_derived_MF)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex = "M & F")

# table for sex combined grounds 
 OR_sex <- Roughy_2 %>%                                      
    dplyr::group_by(sex) %>%
    dplyr::summarise(n = n(), mean = mean(Length),estimated_wgt = mean(Wgt_derived), est_wgt_FM = mean(Wgt_derived_MF)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(Ground = "SH and SP")
 
# table for survey
 OR_survey <- Roughy_2 %>% 
    #ungroup(sex) %>% 
    dplyr::summarise(n = n(), mean = mean(Length),estimated_wgt = mean(Wgt_derived), est_wgt_FM = mean(Wgt_derived_MF)) %>%
    mutate(Proportion = n / sum(n)) %>% 
    mutate(sex ="all fish",
           Ground = "Survey total")
 
v# bind all the tables together, create and rename some coloumns, round off new columns
t2.8 <- bind_rows(OR_sHg, OR_sPg, OR_SHSPbg,OR_sex,OR_survey) %>% 
  mutate("Estimated weight (kg) based on combined Lyle et al (1991)"
         = (0.037*(mean)^2.952)/1000) %>% 
  mutate("Estimated weight (kg) based on sex specific Lyle et al (1991)" =
           if_else(sex == "F",(0.0351 *(mean) ^ 2.970)/1000,
          if_else(sex == "M",(0.0383 * (mean) ^ 2.942)/1000,
                   (0.037*(mean)^2.952)/1000))) %>%
  dplyr::rename("No of fish" = n, "Avg length (cm)" 
                = mean, "Spawning Ground"= Ground, "Prop"=Proportion) %>% 
  dplyr::mutate_at(4:7, funs(round(.,4)))

knitr::kable(tail(t2.8, n = 9), format = "html", caption = "Table 2.8 The proportion of male and female fish at each ground and their associated weighted averages for fish length (cm) for the 2019 Eastern Zone spawning population") %>% 

  kableExtra::kable_styling("striped") %>%

  kableExtra::scroll_box(width = "100%")

write.csv(t2.8,"Figures/Prop_sex_Lylewt.csv")
```

### roughy weight - no weight for 2024

note that this graph has been corrected. The 2019 report the combined Lyle formula was incorrect. The  Lyle formula for Males (Lyle_com_Males) was used instead of the combined sex formula (Lyle_com)

```{r myfig2.6}, out.width='100%', fig.align='center', fig.cap= "Length vs Weight, showing the predicted values (Lyle 1991)"}

Lyle_com <- function(x) {(0.037 * (x) ^ 2.952)}
#Lyle_com_Males <- function(x) {(0.037 * (x) ^ 2.942)}
S2019_fun <- function(x) {(0.056 * (x) ^ 2.8502)}
S2019M_fun <-function(x) {(0.073 * (x) ^ 2.7685)}
S2019F_fun <-function(x) {(0.071 * (x) ^ 2.7945)}

  RoughyWeight<- Roughy%>%
    filter(year==2019) %>%
    drop_na(weight_kg) %>%
    filter(weight_kg != 0) %>% 
    mutate(weight_g = weight_kg *1000) %>%
    mutate(year=as.factor(year))


f2.6<- ggplot(data = RoughyWeight,
       mapping = aes(x  = SLadj_cm,
                     y = weight_g, colour = sex)) +
  scale_colour_manual(values = c("red", "blue"))+
  geom_point(size = .3)+
  scale_x_continuous(breaks = seq(20,50,4),limits = c(20,50))+
  ylab("weight (g)")+
  xlab("Standard length (cm)")+ 
  stat_function(fun = S2019F_fun, colour="red")+
  stat_function(fun = S2019M_fun, colour = "blue") +
  stat_function(fun = Lyle_com, colour ="black")+
  #stat_function(fun = Lyle_com_1, colour ="pink")+
  stat_function(fun = S2019_fun, colour ="orange")+
  theme_bw()+
  
  annotate('text', x = 22, y = 3800, size =4,hjust =0,
           label = expression(paste("Weight"["Lyle sex combined"]^{}, 
                                    " = 0.037 * Length"^{2.952})))+
  
  annotate('text', x = 22, y = 3500, size = 4, hjust = 0,colour = "orange",
           label = expression(paste("Weight"["2019 sex combined"]^{},
                                    " = 0.056 * Length"^{2.850})))+
  
  annotate('text', x = 22, y = 3200, size =4,hjust =0,colour = "red",
           label = expression(paste("Weight"["2019 Females"]^{},
                                    "= 0.071 * Length"^{2.769}))) +
  
  annotate('text', x = 22, y = 2900, size =4, hjust =0, colour = "blue",
           label = expression(paste("Weight"["2019 Males"]^{},
                                    " = 0.073 * Length"^{2.794})))

plot(f2.6)
ggsave("Figures/LvsW.emf",f2.6) # saves the plot to file :)
ggsave("Figures/LvsW.jpg",f2.6) # saves the plot to file :)
ggsave("Figures/LvsW.tiff",f2.6) # saves the plot to file :)


```

### Figure 2.7a
Standard Length (cm) vs weight (kg) for the Eastern Zone spawning population from 1987 to 2019 for both grounds combined, showing the predicted values (plotted black line) from Lyle (1991).

NOTE - this has been corrected as per the above - the correct combined Lyle formula has replaced the Lyle Male formula!!


* Lyle et al. Combined length – weight regression. 
      W = 3.7*10-2 * ^2.952
      
Lyle et al. Female length – weight regression. 
    W = 3.51*10-2 * ^2.970
    
Lyle et al. Male length – weight regression. 
      W = 3.83*10-2 * ^2.942 and 


```{r myfig2.7}, out.width='100%', fig.align='center', fig.cap= "Length vs Weight, showing the predicted values (Lyle 1991)"}

OR_Weight_series_allyears <-Roughy_1 %>% 
  #filter(year==2019) %>%
   #drop_na(weight_kg) %>%
   #filter(weight_kg != 0) %>% 
   mutate(year=as.factor(year)) %>% 
   mutate(Wgt_derived = (0.037 * SLadj_cm ^ 2.952)/1000) %>%
   mutate(Wgt_derived_MF = if_else(sex == "M",(0.0383 * SLadj_cm ^ 2.942)/1000, 
                                   (0.0351 * SLadj_cm ^ 2.970)/1000, NA_real_)) %>% 
  #mutate(Wgt_derived_F = (0.0351 * SLadj_cm ^ 2.970)/1000) %>% 
  #mutate(Wgt_derived_M = (0.0383 * SLadj_cm ^ 2.942)/1000) %>% 
  mutate(MvsD = weight_kg / Wgt_derived_MF) %>%      
  mutate(percentchange = ((weight_kg - Wgt_derived_MF)/ weight_kg)) %>% 
  mutate(wgt_diff = Wgt_derived_MF - weight_kg) %>% 
  arrange(sex)
```

OR Weight vs Length series by year

```{r}
OR_Weight_series <- OR_Weight_series_allyears %>%
  drop_na(weight_kg) %>%
  filter(weight_kg != 0)
    

Lyle_com_kg <- function(x) {(0.037 * (x) ^ 2.952)/1000}
    
f2.7<- ggplot(data = OR_Weight_series,
       mapping = aes(x  = Length,
                     y = weight_kg, colour = year)) +
  geom_point()+
  #geom_smooth(method="glm", method.args=list(family=gaussian(link="log")),
  #colour = "blue", linetype = 2, size = .5)+
  ylab("weight (kg)")+
  xlab("Standard length (cm)")+            
  stat_function(fun = Lyle_com_kg, colour ="black")+
  facet_wrap(~year)

plot(f2.7)
ggsave("Figures/LWindex.emf",f2.7)
ggsave("Figures/LWindex.jpg",f2.7)# saves the plot to file :)
plot(f2.7)

```

## calculate change in LW relationship

```{r}

  OR_Weight_series %>% 
  group_by(Ground,year) %>% 
  dplyr::summarise(n(),MvsD = round(mean(MvsD),2),
                   perchange = round(mean(percentchange),2))
                  


```

Plot measured vs predicted weights

```{r}

OR_Weight_series_sum <- summarySE(OR_Weight_series, measurevar = "MvsD", groupvars = c("year","Ground")) 

knitr::kable(tail(OR_Weight_series_sum, n = 9), format = "html", caption = "Table 2.8 The proportion of male and female fish at each ground and their associated weighted averages for fish length (cm) for the 2019 Eastern Zone spawning population") %>% 

  kableExtra::kable_styling("striped") %>%

  kableExtra::scroll_box(width = "100%")
```


Plot the weight data series
```{r}
pd <- position_dodge(.2)  

f_Wgt_ratio <- ggplot(OR_Weight_series_sum, aes(x=year, y = MvsD, 
                                                shape = Ground))+
  geom_errorbar(aes(ymin = MvsD - ci, ymax = MvsD + ci),
                width = 0.3, position = pd) +
  geom_point(position = pd, size = 2)+
  geom_line()+
  ylab("Measured / Derived (Lyle 1991) Weight (kg)")+
  xlab("Eastern Zone Spawning Population by Year for July")+
  theme_bw()+
   theme(legend.position = c(.95, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'),
        plot.background = element_rect(linetype =  "solid") )
 

plot(f_Wgt_ratio)
ggsave("Figures/f_Wgt_ratio.jpg",f_Wgt_ratio)
ggsave("Figures/f_Wgt_ratio.emf",f_Wgt_ratio)
```

weight data for 2019

```{r}
Wgt_2019_GroundFM <- OR_Weight_series %>% 
  filter(year == 2019) %>% 
  dplyr::group_by(Ground,sex) %>% 
  dplyr::summarise(n= n(), measured = mean(weight_kg), derived = mean(Wgt_derived_MF), mean(MvsD)) %>% 
  mutate(Proportion = n / sum(n)) 

Wgt_ground <- OR_Weight_series %>% 
    filter(year == 2019) %>% 
  dplyr::group_by(Ground) %>% 
  dplyr::summarise(n(), mean(weight_kg), mean(Wgt_derived_MF), mean(Wgt_derived))

Wgt_sex <- OR_Weight_series %>% 
    filter(year == 2019) %>% 
  dplyr::group_by(sex) %>% 
  dplyr::summarise(n(), mean(weight_kg), mean(Wgt_derived_MF))

Wgt_sex <- OR_Weight_series %>% 
    filter(year == 2019) %>% 
  dplyr::group_by(sex) %>% 
  dplyr::summarise(n(), mean(weight_kg), mean(Wgt_derived_MF))

wgt_survey <- OR_Weight_series %>% 
  filter(year == 2019) %>% 
  dplyr::summarise(n(), mean(weight_kg), mean(Wgt_derived))
```

weight data for 2024

```{r}
Wgt_2024_GroundFM <- OR_Weight_series_allyears %>% 
  filter(year == 2024) %>% 
  dplyr::group_by(Ground,sex) %>% 
  dplyr::summarise(n= n(), derived = mean(Wgt_derived_MF)) %>% 
  mutate(Proportion = n / sum(n)) 

Wgt_ground_2024 <- OR_Weight_series_allyears %>% 
    filter(year == 2024) %>% 
  dplyr::group_by(Ground) %>% 
  dplyr::summarise(n(),mean(Wgt_derived_MF))

Wgt_sex_2024 <- OR_Weight_series_allyears %>% 
    filter(year == 2024) %>% 
  dplyr::group_by(sex) %>% 
  dplyr::summarise(n(), mean(Wgt_derived_MF))

Wgt_sex <- OR_Weight_series %>% 
    filter(year == 2019) %>% 
  dplyr::group_by(sex) %>% 
  dplyr::summarise(n(), mean(weight_kg), mean(Wgt_derived_MF))

wgt_survey_2024 <- OR_Weight_series_allyears %>% 
  filter(year == 2024) %>% 
  dplyr::summarise(n(), mean(Wgt_derived))
```


# Age 

Read in age data
```{r}
Age <- read_csv("Results/RoughyEZ_AgeIndex_1986to2019.csv")
Age2024_Owgt <- read_csv("Data/Roughy_Age2024_owgt.csv")
Age_2019 <- read_csv("Data/RoughyAge2019.csv")
```

take a look at the data Age index data
```{r}
Age%>% 
  dplyr::group_by(Sex,Year) %>% 
  dplyr::summarise(mean(O,na.rm = TRUE),mean(Age,na.rm = TRUE))
```
need to add age data from 2019
need to add otolith weight data from 2024
ototlith age data not yet available for 2024

## Add 2019 otolith age data 


remove the 2019 data from the AgeIndex and add the 2019 data that inlcudes the "Age" data.

```{r}
Age_filtered <- Age %>% 
  filter(Year != 2019) %>% 
  dplyr::rename(Month=month) %>% 
  mutate(as.character(Month))
```

take a look at the data Age index data

```{r}
Age_filtered%>% 
  dplyr::group_by(Sex,Year) %>% 
  dplyr::summarise(mean(O,na.rm = TRUE),mean(Age,na.rm = TRUE), n())
```

add the 2019 age dataset

what columns are required
```{r}
colnames(Age_filtered)
colnames(Age_2019)
```

select 2019 fields

```{r}
Age_2019ed <- Age_2019 %>% 
  select(Zone, Ground = AreaCapture, Year, Month, Age= `Age1...14`, Sex,Length_1, O= `Otolith_eight (g)`) %>% 
   mutate(as.character(Month)) %>% 
  mutate(length_cm = Length_1/10)
  #filter(Age >0) no age data currently

Age_2019ed %>% 
  dplyr::group_by(Sex,Year) %>% 
  dplyr::summarise(mean(O,na.rm = TRUE), mean(Age,na.rm = TRUE))
```

bind 2019 rows to Age Index, rename Age

```{r}
Age_1 <- Age_filtered %>% 
  bind_rows(Age_2019ed)
  # mutate(Sex_no = Sex) %>% 
  # mutate(Sex= as.factor(Sex)) %>% 
  # mutate(Sex = case_when(Sex == "1"~"M", TRUE~as.character(Sex))) %>%
  # mutate(Sex = case_when(Sex == "2"~"F", TRUE~as.character(Sex))) 
```

check data

```{r}
Age_1 %>% 
  dplyr::group_by(Year,Sex,) %>% 
  dplyr::summarise(mean(O,na.rm = TRUE),mean(Age,na.rm = TRUE), n())
```

Add the 2024 Age data - otolith weight as proxy

```{r}
colnames(Age_1)
colnames(Age2024_Owgt)

```

standardise new data to the Age Index

add columns
```{r}
Age2024<- Age2024_Owgt %>% 
  select(Zone,Shot,Sex, Length_cm = Length_1, O = Oto_Wght) %>% 
  add_column(Mark = NA, Sex_no = NA) 
  # mutate(as.character(Month))
```

bind to shot data to get Location!
```{r}
Age2024 <- Age2024 %>% 
  left_join(Shot2024_ed, by = "Shot")
```

select matching columns

```{r}
colnames(Age_1)

Age2024 <- Age2024 %>% 
  select(Zone, Ground = Location, Year, Month, Sex, Length_cm, O, Mark, Sex_no)

```

bind data
```{r}
Age_2 <- Age_1 %>% 
  dplyr::bind_rows(Age2024) %>% 
   filter(Sex=="M"|Sex=="F")
```

check data

```{r}
Age_2 %>% 
  dplyr::group_by(Year,Sex,) %>% 
  dplyr::summarise(mean(O,na.rm = TRUE),mean(Age,na.rm = TRUE), n())
 
```

## Plot Age data: Age and otolith weight

```{r}
O_Sum <- summarySE(Age_2, measurevar = "O", groupvars = c("Year","Ground","Sex"),na.rm = TRUE) 
Age_Sum <- summarySE(Age_2, measurevar = "Age", groupvars = c("Year","Ground","Sex"),na.rm = TRUE) 

pd <- position_dodge(0.5) # move position horizontally b/c of overlap

f2.7 <- ggplot(O_Sum, aes(x = Year,y = O, shape = Ground, colour = Sex)) +
  geom_errorbar(aes(ymin = O - ci, ymax= O + ci),
                width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_line(position = pd) +
  xlab("Eastern Zone Spawning Population by year for July")+
  ylab("Average otolith weight (g)")+
  scale_y_continuous(breaks = seq(.05,0.3,.05),limits = c(0,0.28))+
  scale_x_continuous(limits = c(1986,2026), breaks = seq(1986,2026,4))+
  theme_bw()+
  theme(legend.position = c(.25, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'),
        plot.background = element_rect(linetype =  "solid") )


f2.8 <- ggplot(Age_Sum, aes(x = Year,y = Age, shape = Ground, colour = Sex)) +
  geom_errorbar(aes(ymin = Age - ci, ymax= Age + ci),
                width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_line(position = pd) +
  xlab("Eastern Zone Spawning Population by year for July")+
  ylab("Average age (years)")+
  scale_y_continuous(breaks = seq(20,70,10),limits = c(20,70))+
  scale_x_continuous(limits = c(1986,2022), breaks = seq(1986,2022,4))+
  theme_bw()+
  theme(legend.position = c(.25, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'),
        plot.background = element_rect(linetype =  "solid") )

plot(f2.7)
ggsave("Figures/Age_owgt.emf")
ggsave("Figures/Age_owgt.jpg")
plot(f2.8)
ggsave("Figures/Age_years.emf")
ggsave("Figures/Age_owgt.jpg")
```
### Figure 2.9 Otolith weight by year
```{r myfig2.9}

O_Sum <- summarySE(Age_2, measurevar = "O", groupvars = c("Year","Ground","Sex"),na.rm = TRUE) 

# Otolith Weight vs time series
f2.9 <- ggplot(O_Sum, aes(x = Year,y = O, shape = Ground, colour = Sex)) +
  geom_errorbar(aes(ymin = O - ci, ymax= O + ci),
                width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_line(position = pd) +
  xlab("Eastern Zone Spawning Population by year for July")+
  ylab("Average otolith weight (g)")+
  scale_y_continuous(breaks = seq(.1,0.28,.02),limits = c(.1,0.28))+
  scale_x_continuous(limits = c(1986,2022), breaks = seq(1986,2022,4))+
  theme_bw()+
  theme(legend.position = c(.2, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'),
        plot.background = element_rect(linetype =  "solid") )
plot(f2.9)
ggsave("Figures/Age_Owgt.emf")
ggsave("Figures/Age_Owgt.jpg")


write.csv(Age_2,"Results/Otolith_wt.csv")
```

### Age regression

```{r}
AgeRegression <- summarySE(Age_2, measurevar = "Age", groupvars = c("Year","Sex", "Ground"),na.rm = TRUE) 

ggplot(AgeRegression, aes(x = Year,y = Age, colour = Sex, shape = Ground)) +
#ggplot(AgeRegression, aes(x = Year,y = Age)) +
  # geom_errorbar(aes(ymin = O - ci, ymax= O + ci),
  #               width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_smooth(method='lm')
```
Regression for data up to 2019 by sex and ground
```{r}
A_SHF <- AgeRegression %>% 
  filter(Sex == "F") %>% 
  filter(Ground == "St Helens")

A_SHM <- AgeRegression %>% 
  filter(Sex == "M") %>% 
  filter(Ground == "St Helens")

A_SPF <- AgeRegression %>% 
  filter(Sex == "F") %>% 
  filter(Ground == "St Patricks")

A_SPM <- AgeRegression %>% 
  filter(Sex == "M") %>% 
  filter(Ground == "St Patricks")

regression_shf <- lm(Age ~ Year , data = A_SHF)
regression_shm <- lm(Age ~ Year , data = A_SHM)
regression_spf <- lm(Age ~ Year , data = A_SPF)
regression_spm <- lm(Age ~ Year , data = A_SPM)

#Short summary of the model


#Longer summary of the model
summary(regression_shf)
summary(regression_shm)
summary(regression_spf)
summary(regression_spm)
```
Regression for recent data 
```{r}
Age_recent <- AgeRegression %>% 
  filter(Sex == "F" & Ground == "St Helens") %>% 
  filter(Year >= 2010)

  distinct(Age_recent, Year)

ggplot(Age_recent, aes(x = Year,y = Age, colour = Sex, shape = Ground)) +
  # geom_errorbar(aes(ymin = O - ci, ymax= O + ci),
  #               width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_smooth(method='lm')



regression_recent <- lm(Age ~ Year, data = Age_recent)


#Longer summary of the model
summary(regression_recent)
```
results
SH F, Year          0.0040  
SH M, Year           0.3357
SP F, Year           0.1016 NS
SP M, Year          0.02392 NS

checking Rudy's statistics.. 
all good :)
```{r}
Age_2016 <- AgeRegression %>% 
  filter(Year <= 2016)%>% 
  #filter(Year >= 2010) %>% 
  filter(Sex == "F" & Ground == "St Patricks")

ggplot(Age_2016, aes(x = Year,y = Age, colour = Sex, shape = Ground)) +
  
  # geom_errorbar(aes(ymin = O - ci, ymax= O + ci),
  #               width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_smooth(method='lm')



#regression_2016 <- lm(Age ~ Year + Sex + Ground, data = Age_2016)
regression_2016 <- lm(Age ~ Year, data = Age_2016)

#Longer summary of the model
summary(regression_2016)
```

### compare ototlith weight between 2019 and 2024

```{r}
Owgt_2024 <- Age_2 %>% 
  filter(Year == 2019 | Year == 2024)
  #filter(Year == 2019 | Year == 2016)
  #filter(Year >= 2010) %>% 

Owgt_2024 %>%

  dplyr::group_by(Year,Sex) %>%

  dplyr::group_by(Sex) %>%

  dplyr::summarise(mean(O,na.rm = TRUE),mean(Age,na.rm = TRUE), n())

#regression_2016 <- lm(Age ~ Year + Sex + Ground, data = Age_2016)
regression_Owgt_2024 <- lm(O ~ Year + Sex + Ground, data = Owgt_2024)

#Longer summary of the model
summary(regression_Owgt_2024)
```

### compare age between 2016 and 2019

```{r}
Owgt_2024 <- Age_2 %>% 
  filter(Year == 2019 | Year == 2016)
  #filter(Year == 2019 | Year == 2016)
  #filter(Year >= 2010) %>% 

Owgt_2024 %>%
  dplyr::group_by(Ground, Sex, Year) %>%
  dplyr::summarise(mean(O,na.rm = TRUE),mean(Age,na.rm = TRUE), n())

#regression_2016 <- lm(Age ~ Year + Sex + Ground, data = Age_2016)
regression_Owgt_2024 <- lm(Age ~ Year + Sex + Ground, data = Owgt_2024)

#Longer summary of the model
summary(regression_Owgt_2024)
```