---
title: "RoughyData_exploration"
author: "sutton"
date: "2025-07-14"
output: html_document
---

NOTES

Length weight data from Simon Boags (SETFIA) for the 2025 Eastern Zone

Simon Boags data fits well into the Roughy index data for the Eastern Zone
- residuals vs fitted look good for both F and M
- consistent with both Lyle and 2019 data


## Set up
libraries required

```{r setup, include=FALSE}
library(tidyverse)    # includes dplyr, ggplot2, tibble, tidyr,
library(broom)
library(kableExtra)
library(knitr)
library(cowplot)
library(gridExtra)
library(lubridate)
library(devEMF)

#mapping
# library(maps)
# library(mapdata)
# library(mapproj)
# library(sf)


```



# Eastern Zone Roughy LW data - Simon Boags

load data
```{r}
Boags <- read_csv("Data/Roughy_SESSF/SimonBoags_EasternLWdata.csv")
```

## wrangle data

name fix
```{r}
colnames(Boags)

Boags <- Boags %>% 
  dplyr::rename(sex = Sex) %>% 
  dplyr::rename(length_cm = Length_cm) %>% 

colnames(Boags)
```

add predicted weights to the Boags data based on Lyle and 2019 models
```{r}
Boags <- Boags %>% 
  mutate(Wgt_Lyle = if_else(sex == "Male",(0.0383 * length_cm ^ 2.942), 
                                   (0.0351 * length_cm ^ 2.970), NA_real_)) %>% 
  mutate(Wgt_2019 = if_else(sex == "Male",(0.2006 * length_cm ^ 2.479), 
                                   (0.1073 * length_cm ^ 2.673), NA_real_)) %>% 
  mutate()
```


## Boags data - length vs weight

boags data fit model
use nls() instead to negate the need to back-transform
plot place the plot into context of the index series

do everything by sex
plot residuals
compare value from model with real value
predicted (boags and lyle) vs real for each year



Boags LW
```{r}
ggplot(data = Boags,
       aes(x = length_cm, y = weight_g, colour = sex)) +
  geom_point(size = 0.5)
  
```

## LW models

create data frames for each sex
```{r}
# Filter by sex
boags_female <- Boags %>% filter(sex == "Female")
boags_male   <- Boags %>% filter(sex == "Male")
```

fit a model - lm
```{r}
# Fit non-linear model: weight = a * length^b
model_female <- lm(log(weight_g) ~ log(length_cm), data = boags_female)
model_male   <- lm(log(weight_g) ~ log(length_cm), data = boags_male)
```

```{r}
summary(model_female)
```
```{r}
coef(model_female)
```


```{r}
residuals(model_female)
```

```{r}
fitted(model_female)
```

residuals 
looks ok
```{r}
plot(model_female)
```

looks ok
```{r}
plot(model_male)
```
female: intercept = -1.89, slope = 2.57, 
male:   intercept = -1.91, slope = 2.55, 

F : a = 0.1503, b = 2.5721
M : a = 0.1480  b = 2.5526

intercept needs to be back-transformed 
```{r}
#exp(coef(model_female)[1])  # a_female
exp(-1.89)
coef(model_female)[2]       # b_female

exp(coef(model_male)[1])    # a_male
coef(model_male)[2]         # b_male
```


LW functions
```{r myfigBoags}
Lyle_com <- function(x) {0.037 * x^2.952}
Boags_f <- function (x){0.151 * x ^2.572}
Boags_m <- function (x){0.148 * x ^2.552}
Lyle_m <- function(x) {0.0383 * x^2.942}
Lyle_f <- function(x) {0.0351 * x^2.970}

# checked in 2025
S2019F <- function(x) {(0.1073 * (x) ^ 2.673)} 
S2019M <- function(x) {(0.2006 * (x) ^ 2.479)} 
S2019 <- function(x) {(0.0555 * (x) ^ 2.850)} 

```

## Plots
plot LW data
```{r}
ggplot(data = Boags,
       aes(x = length_cm, y = weight_g, colour = sex)) +
  geom_point(size = 0.3) +
  # geom_smooth(method = "nls",
  #             formula = y ~ a * x^b,
  #             method.args = list(start = list(a = 0.03, b = 3)),
  #             se = FALSE, linewidth = 1) +
  scale_colour_manual(values = c("Female" = "red", "Male" = "blue")) +
  #stat_function(fun = Lyle_com, colour = "black", linewidth = 0.8) +
  stat_function(fun = Lyle_f, colour = "black", linewidth = 0.8, linetype = 2) +
  stat_function(fun = Lyle_m, colour = "black", linewidth = 0.8) +
  stat_function(fun = Boags_f, colour = "red", linewidth = 0.8) +
    stat_function(fun = Boags_m, colour = "blue", linewidth = 0.8) 


```
Plot by sex
```{r}
# Annotation dataframe
annotation_df <- tribble(
  ~sex,      ~x,  ~y,   ~label,                         ~text_colour,
  "Female",  15, 2800, "Boags: W = 0.151 × L^2.572",    "red",
  "Female",  15, 2700, "Lyle:  W = 0.0351 × L^2.970 -black", "black",
  "Female",  15, 2600, "2019:  W = 0.1073 × L^2.673 -brown", "brown",
  "Male",    15, 2800, "Boags: W = 0.148 × L^2.552",    "blue",
  "Male",    15, 2700, "Lyle:  W = 0.0383 × L^2.942 -black", "black",
  "Male",    15, 2600, "2019:  W = 0.2006 × L^2.479- brown", "brown"
)

# Plot
ggplot(data = Boags, aes(x = length_cm, y = weight_g)) +
  geom_point(aes(colour = sex), size = 0.3) +

  # Boags fits
  stat_function(data = Boags %>% filter(sex == "Female"),
                fun = Boags_f, colour = "red", linewidth = 0.8) +
  stat_function(data = Boags %>% filter(sex == "Male"),
                fun = Boags_m, colour = "blue", linewidth = 0.8) +

  # Lyle fits
  stat_function(data = Boags %>% filter(sex == "Female"),
                fun = Lyle_f, colour = "black", linewidth = 0.8, linetype = "dashed") +
  stat_function(data = Boags %>% filter(sex == "Male"),
                fun = Lyle_m, colour = "black", linewidth = 0.8, linetype = "dashed") +

  # 2019 fits
  stat_function(data = Boags %>% filter(sex == "Female"),
                fun = S2019F, colour = "brown", linewidth = 0.8, linetype = "dashed") +
  stat_function(data = Boags %>% filter(sex == "Male"),
                fun = S2019M, colour = "brown", linewidth = 0.8, linetype = "dashed") +

  # Add annotations
  geom_text(
    data = annotation_df,
    aes(x = x, y = y, label = label, colour = text_colour),
    inherit.aes = FALSE,
    hjust = 0, size = 3
  ) +
scale_colour_identity(guide = "none")+
  # Facet by sex
  facet_wrap(~sex) +

  # Set point colours only
  scale_colour_manual(values = c("Female" = "red", "Male" = "blue")) +

  labs(x = "Length (cm)", y = "Weight (g)", colour = "Sex") +
  theme_minimal()
```


### compare value from model with real value
add predicted weights to the Boags data based on Lyle and 2019 models
```{r}
Boags <- Boags %>% 
  mutate(Wgt_Lyle = if_else(sex == "Male",(0.0383 * length_cm ^ 2.942), 
                                   (0.0351 * length_cm ^ 2.970), NA_real_)) %>% 
  mutate(Wgt_2019 = if_else(sex == "Male",(0.2006 * length_cm ^ 2.479), 
                                   (0.1073 * length_cm ^ 2.673), NA_real_)) %>% 
 mutate(Wgt_boags = if_else(sex == "Male",(0.148 * length_cm ^ 2.552), 
                                   (0.151 * length_cm ^ 2.572), NA_real_))


```

look at difference between model and real values for weight

Predicted vs Observed for Boags data
```{r}
ggplot(Boags, aes(x = Wgt_boags, y = weight_g, colour = sex)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(x = "Wgt_boags_model", y = "weight_g") +
  theme_minimal()
```

Predicted_Lyle vs Observed 
```{r}
ggplot(Boags, aes(x = Wgt_Lyle, y = weight_g, colour = sex)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  labs(x = "Predicted", y = "Observed") +
  theme_minimal()
```

Predicted_2019 vs observed
```{r}
ggplot(Boags, aes(x = Wgt_2019, y = weight_g, colour = sex)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "brown") +
  labs(x = "Predicted", y = "Observed") +
  theme_minimal()
```

--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
#WORDAC                                    --------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

Franzis requested WORDAC data from AFMA - July 2025

load data

```{r}
### Biological Index
WORDAC <- read_csv("Data/Roughy_SESSF/WORDAC_2024in2025_extract_biological.csv",
                   na = c("", "(null)"),
                  col_types = list(LENGTH_CM = col_double()))
```

separate date into D M Y
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
- had trouble with dates 
- lubridate can't cope with two date formats - the last 52 lines had hh:mm in the date. This was problematic. I went back to excel to make all the formats the same b/c it couldn't make it work in R.

so manipulated dates in excel instead - attempted code below
```{r}
# WORDAC <- WORDAC %>% 
#   # lubridate - turns character date to date-date
#   mutate(x = dmy(SHOT_DATE)) %>% 
#   # splits the date into year month and day
#   mutate_at(vars(x), funs(year, month, day))

#gsub(x=v,pattern=" 0:00:00",replacement="",fixed=T)
```

explore the data

```{r}
WORDAC %>% 
  # dplyr::group_by(ZONE,year,month) %>% 
  # dplyr::summarise(n = n())
    dplyr::group_by(ZONE,year,month) %>% 
  dplyr::summarise(n = n())

unique(WORDAC$SEX)
unique(WORDAC$ZONE)
unique(WORDAC$SAMPLING_LOCATION)
unique(WORDAC$SEASON)
unique(WORDAC$STRATUM)

```

## Manipulate data


Drop NA in length column
```{r}
WORDAC_1 <-WORDAC %>% 
  drop_na(LENGTH_CM)

unique(WORDAC_1$LENGTH_CM)
```
Note that dropping (null) for length has also dropped the (null) for Sex


ZONE colunm - combine the null and 0 to be NA
- makes graphing easier
```{r}

 
WORDAC_1 <- WORDAC_1 %>%  
  mutate(ZONE = ifelse(ZONE == 0, NA, ZONE)) %>% 
  mutate(ZONE = ifelse(ZONE == "(null)", NA, ZONE)) 

# check
unique(WORDAC_1$ZONE)
```


change months to character - in order to read the months 
```{r}
WORDAC_1 <- WORDAC_1 %>%
  mutate(MonthName = factor(month.abb[as.numeric(month)], levels = month.abb))

#check
WORDAC_1 %>% 
  dplyr::group_by(SEX,month,MonthName) %>% 
  dplyr::summarise(n = n())
```
## Plot data on a map

code from chatgpt
load appropriate libraries for mapping

```{r}
library(maps)
library(mapdata)
library(mapproj)
library(sf)
``` 

```{r}
#Example data format

colnames(WORDAC)

aus_map <- map_data("world", region = "Australia")
padding <- 0.5  # degrees of lat/lon padding

lon_min <- min(WORDAC_1$LONGITUDE, na.rm = TRUE) - padding
lon_max <- max(WORDAC_1$LONGITUDE, na.rm = TRUE) + padding
lat_min <- min(WORDAC_1$LATITUDE, na.rm = TRUE) - padding
lat_max <- max(WORDAC_1$LATITUDE, na.rm = TRUE) + padding

# the map
ggplot() +
  geom_polygon(data = aus_map, aes(x = long, y = lat, group = group),
               fill = "gray90", color = "black") +
  geom_point(data = WORDAC_1, aes(x = LONGITUDE, y = LATITUDE, color = ZONE),
             size = 2, alpha = 0.8) +
  coord_fixed(1.3, 
              xlim = c(lon_min, lon_max), 
              ylim = c(lat_min, lat_max)) +
  theme_minimal() +
  labs(title = "Sample Locations by Zone",
       x = "Longitude", y = "Latitude",
       color = "Zone") +
  theme(legend.position = "bottom")
```

### checking zones from Lookups
note that all the samples from the WORDAC data fit into ORO - Northern Zone
but some of the data don't fit correctly into the data zones.. need to rezone some of the data

load lookup library
```{r}
library(SESSFlookups)
```

```{r}
Zone_ORO <- zoneLU_2024_11 %>% 
  filter(ZoneScheme == "ORO")

```

### checking the data fit into the WORDAC zones 
code from chatgpt

Define custom zone polygons
```{r}
# Step 1: Create custom zone polygons
zones_df <- tibble(
  ZONE = c("Northern", "Central", "Southern"),
  geometry = list(
    st_polygon(list(rbind(  # Northern
      c(138.0, -38.0),
      c(140.0, -38.0),
      c(140.0, -37.0),
      c(138.0, -37.0),
      c(138.0, -38.0)
    ))),
    st_polygon(list(rbind(  # Central
      c(140.0, -40.0),
      c(143.5, -40.0),
      c(143.5, -38.0),
      c(140.0, -38.0),
      c(140.0, -40.0)
    ))),
    st_polygon(list(rbind(  # Southern
      c(143.5, -43.5),
      c(145.0, -43.5),
      c(145.0, -40.5),
      c(143.5, -40.5),
      c(143.5, -43.5)
    )))
  )
)

zones_sf <- st_sf(zones_df, crs = 4326)

# Step 2: Prepare sample data
# Replace NA zones with "Unknown"
WORDAC_1 <- WORDAC_1 %>%
  mutate(ZONE = if_else(is.na(ZONE), "Unknown", ZONE))

# Convert to sf object with coordinates
samples_sf <- WORDAC_1 %>%
  filter(!is.na(LONGITUDE), !is.na(LATITUDE)) %>%
  mutate(ZONE_original = ZONE) %>%
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326)

# Step 3: Spatial join with zones
samples_reassigned <- st_join(samples_sf, zones_sf, join = st_within)

# Step 4: Rename to avoid duplicate names
# First, check actual column names:
zone_cols <- names(samples_reassigned)[grepl("ZONE", names(samples_reassigned))]
print(zone_cols)
zone_cols

# Then rename based on detected names:
# (adjust these if your suffixes differ)
samples_reassigned <- samples_reassigned %>%
  dplyr::rename(
    ZONE_from_data = ZONE.x,         # Your original zone (before spatial join)
    ZONE_reassigned = ZONE.y         # New zone from polygon
  )

# Step 5: Plot the reassigned sample points and zone polygons
ggplot() +
  geom_sf(data = zones_sf, aes(fill = ZONE), alpha = 0.3, color = "black") +
  geom_sf(data = samples_reassigned, aes(color = ZONE_reassigned), size = 2, alpha = 0.8) +
  theme_minimal() +
  labs(
    title = "Sample Locations with Reassigned Zones",
    fill = "Zone Polygons",
    color = "Reassigned Zone"
  ) +
  theme(legend.position = "bottom")
```

### adding onto data
```{r}
anyDuplicated(WORDAC_1$PORTSAMPLE_ID)

zone_lookup <- samples_reassigned %>%
  st_drop_geometry() %>%
  select(PORTSAMPLE_ID, ZONE = ZONE_reassigned)  # Rename to avoid overwriting too soon

WORDAC_1 <- WORDAC_1 %>%
  mutate(ZONE_original = ZONE) %>%     # Save old ZONE
  select(-ZONE) %>% # so you don't get suffixes and confusion
  left_join(zone_lookup, by = "PORTSAMPLE_ID")  # New ZONE replaces old one
```

## Plotting Length data

* by sex - y
* by sex and month 
* by sex and zone - y
* by month for each zone
### Length Histograms by Sex

Females as expected are on average larger than males
-histograms suggest different cohorts - possibly difference in zones
- stats show that indeterminate fish are on averages smaller than F fish (-3.2 cm) and M fish (-1.9). 

```{r}
summarySE(WORDAC_1, measurevar = "LENGTH_CM", groupvars = c("SEX")) 

WORDAC_1 %>%
  dplyr::group_by(SEX) %>%
  dplyr::summarise(meanSL = mean(LENGTH_CM, na.rm = TRUE))
```

Length by SEX

```{r}
ggplot(
  data = WORDAC_1 ,
  aes(x = LENGTH_CM, fill = SEX)) +
  geom_histogram(binwidth = 0.5, position = "identity", alpha = 0.5) +
  geom_density(aes(y = ..count.., color = SEX), size = .1, adjust = .5, fill = NA) +
  xlab("Standard Length (cm) by year") +
  ylab("Count") +
  #facet_grid(rows = vars(year), cols = vars(SEX), scales = "free_y")+
  facet_grid(cols = vars(SEX), scales = "free_y")+
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1") +
  labs(fill = "Sex", color = "Sex") +
  theme_minimal()
```
Thought perhaps the small fish may not have been easy to sex and that 'indeterminate' fish may have been obviously smaller - not super obvious - as many large fish are also classed as indeterminate.

Add in a juvenile column
```{r}
WORDAC_1 <- WORDAC_1 %>% 
  mutate(Juv = ifelse(LENGTH_CM < 28,"J","M")) 

JuvFish <- WORDAC_1 %>% 
  select(ZONE,year,month, SEX,LENGTH_CM,Juv) %>%
  group_by(ZONE,year,month,Juv) %>% 
  dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
  dplyr::group_by(ZONE,year,month) %>% 
    dplyr::mutate(prop = n / sum(n)) %>%
  dplyr::ungroup() %>% 
  
  
  
  
  

  dplyr::group_by(ZONE, SEX, LENGTH_CM) %>%
  dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
  
  # 2. Get total per ZONE (not per facet!)
  dplyr::group_by(ZONE) %>%
  dplyr::mutate(prop = n / sum(n)) %>%
  dplyr::ungroup()
  
```

Check for statistical differences in sex

Thought perhaps the small fish may not have been easy to sex and that 'indeterminate' fish may have been obviously smaller - not super obvious - as many large fish are also classed as indeterminate.

Sex and interaction is significant

Tukey = all groups are diff from eachother. Both F and M are longer than Indeterminate. Suggests some evidence that smaller fish/imature fish were hard to sex?
```{r}
anova_model <- aov(LENGTH_CM ~ SEX, data = WORDAC_1)
summary(anova_model)
TukeyHSD(anova_model)
```

visualise results
```{r}
ggplot(WORDAC_1, aes(x = SEX, y = LENGTH_CM, fill = SEX)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  labs(x = "Sex", y = "Length (cm)")

# double check b/c plot is not convincing...
WORDAC_1 %>%
  dplyr::group_by(SEX) %>%
  dplyr::summarise(
    mean_length = mean(LENGTH_CM, na.rm = TRUE),
    n = n()
  )
```


### Length by sex and zone
ANOVA and Tukey test
 - Southern fish are relatively MUCH smaller than BOTH Northern and Central fish 
    *(-6.2cm and -4.91 cm respectively)
 - Northern fish are larger than Central fish 
    *(1.29cm)

Length by sex and zone - prop per zone
```{r}
lf_zone_prop <- WORDAC_1 %>%
  dplyr::filter(SEX %in% c("F", "M")) %>%
  dplyr::group_by(ZONE, SEX, LENGTH_CM) %>%
  dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
  
  # 2. Get total per ZONE (not per facet!)
  dplyr::group_by(ZONE) %>%
  dplyr::mutate(prop = n / sum(n)) %>%
  dplyr::ungroup()

# 3. Sample size labels (total per ZONE × SEX)
n_labels <- WORDAC_1 %>%
  dplyr::filter(SEX %in% c("F", "M")) %>%
  dplyr::group_by(ZONE, SEX) %>%
  dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
  dplyr::mutate(label = paste0("n = ", n))

# 4. Plot
ggplot(lf_zone_prop, aes(x = LENGTH_CM, y = prop, fill = SEX)) +
  geom_col(width = 0.5, color = "black", alpha = 0.6) +

  # Add sample size labels in top-right of each facet
  geom_text(
    data = n_labels,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.3,
    inherit.aes = FALSE,
    size = 3.5
  ) +

  facet_grid(rows = vars(ZONE), cols = vars(SEX), scales = "free_y") +
  xlab("Standard Length (cm)") +
  ylab("Proportion (relative to total fish in zone)") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal(base_size = 12) +
  theme(strip.text = element_text(size = 10, face = "bold"))

```

test significance - anova
```{r}
Anova_Zone <- aov(LENGTH_CM ~ ZONE * SEX, data = WORDAC_1 %>% filter(SEX %in% c("F", "M")))
summary(Anova_Zone)
TukeyHSD(Anova_Zone)
```
visulise differences
```{r}
mean_ci <- WORDAC_1 %>%
  filter(SEX %in% c("F", "M")) %>%
  dplyr::group_by(ZONE, SEX) %>%
  dplyr::summarise(
    mean = mean(LENGTH_CM, na.rm = TRUE),
    sd = sd(LENGTH_CM, na.rm = TRUE),
    n = n(),
    se = sd / sqrt(n),
    ci = se * 1.96,  # 95% confidence interval
    .groups = "drop"
  )

# 2. Plot
ggplot(mean_ci, aes(x = ZONE, y = mean, fill = SEX)) +
  geom_col(position = position_dodge(0.9), width = 0.7, color = "black", alpha = 0.8) +
  geom_errorbar(
    aes(ymin = mean - ci, ymax = mean + ci),
    position = position_dodge(0.9),
    width = 0.2
  ) +
  labs(
    title = "Mean Standard Length by Zone and Sex",
    x = "Zone",
    y = "Mean Standard Length (cm)",
    fill = "Sex"
  ) +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    axis.title = element_text(size = 13),
    legend.position = "right"
  )
```
```{r}

```

By year

Length by sex and year and ZONE 
```{r}
lf_zone_year_prop <- WORDAC_1 %>%
  dplyr::filter(SEX %in% c("F", "M")) %>%
  mutate(sex_group = factor(SEX, levels = c("F", "M"))) %>% 
  dplyr::group_by(ZONE,year, SEX, LENGTH_CM) %>%
  dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
  
  # 2. Get total per ZONE and year (not per facet!)
  dplyr::group_by(ZONE, year, SEX) %>%
  dplyr::mutate(prop = n / sum(n)) %>%
  dplyr::ungroup()

# 3. Sample size labels (total per ZONE x year × SEX)
n_labels <- WORDAC_1 %>%
  dplyr::filter(SEX %in% c("F", "M")) %>%
  dplyr::group_by(ZONE, year, SEX) %>%
  dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
  dplyr::mutate(label = paste0("n = ", n))

# 4. Plot
ggplot(lf_zone_year_prop, aes(x = LENGTH_CM, y = prop, fill = SEX)) +
  geom_col(
    position = position_dodge(width = 1),
    width = 0.7,
    #color = "clear",
    alpha = 0.7
  ) +
  scale_fill_manual(values = c("F" = "red", "M" = "blue"), name = "Sex") +

  # Add sample size labels in top-right of each facet
  geom_text(
    data = n_labels,
    aes(x = Inf, y = Inf, label = label),
    hjust = 1.1, vjust = 1.3,
    inherit.aes = FALSE,
    size = 2.5
  ) +
  
  #facet_grid(rows = vars(ZONE), cols = vars(year), scales = "free_y") +
  facet_grid(year ~ SEX + ZONE, scales = "free_y") +  # facet by year rows and sex+month columns
  xlab("Standard Length (cm)") +
  ylab("Proportion") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal(base_size = 12) +
  theme(strip.text = element_text(size = 10, face = "bold"))
```
Look for differences by month = look at each zone seperatley

do this by each zone

***** not sure if prop are calculated correctly
check 
```{r}
WORDAC_1 %>%
  filter(ZONE == "Central", year == 2020, month == 5, SEX == "F") %>%
  filter(!is.na(LENGTH_CM)) %>%
  mutate(LENGTH_BIN = floor(LENGTH_CM)) %>%
  dplyr::group_by(LENGTH_BIN) %>%
  dplyr::summarise(count = n()) %>%
  dplyr::summarise(total_binned = sum(count))

WORDAC_1 %>%
  filter(ZONE == "Central", year == 2020, month == 5, SEX == "F") %>%
  filter(!is.na(LENGTH_CM)) %>%
  dplyr::summarise(total = n())
```
```{r}
# 1. Bin length values (e.g. round down to nearest cm)
lf_zone_year_prop <- WORDAC_1 %>%
  filter(SEX %in% c("F", "M")) %>%
  dplyr::mutate(LENGTH_BIN = floor(LENGTH_CM)) %>%  # <- NEW binning step
  dplyr::group_by(ZONE, year, month, SEX, LENGTH_BIN) %>%
  dplyr::summarise(count = n(), .groups = "drop") %>%
  dplyr::group_by(ZONE, year, month, SEX) %>%
  mutate(prop = count / sum(count)) %>%
  ungroup()

# 2. Create sample size labels
n_labels <- WORDAC_1 %>%
  filter(SEX %in% c("F", "M")) %>%
  dplyr::group_by(ZONE, year, SEX, month) %>%
  dplyr::summarise(label = paste0("n=", n()), .groups = "drop")

# 3. Generate one plot per zone
zones <- unique(lf_zone_year_prop$ZONE)

for (zone_i in zones) {
  p <- ggplot(
    lf_zone_year_prop %>% filter(ZONE == zone_i),
    aes(x = LENGTH_BIN, y = prop, fill = SEX)
  ) +
    geom_col(position = position_dodge(width = 1), width = 0.7, alpha = 0.7) +
    scale_fill_manual(values = c("F" = "red", "M" = "blue"), name = "Sex") +
    
    # Add sample size labels
    geom_text(
      data = n_labels %>% filter(ZONE == zone_i),
      aes(x = Inf, y = Inf, label = label),
      hjust = 1.1, vjust = 1.3,
      inherit.aes = FALSE,
      size = 2.5
    ) +
    
    # Facet by year × (SEX + month)
    facet_grid(year ~ SEX + month, scales = "free_y") +
    
    # Labels and styling
    xlab("Standard Length (cm, binned)") +
    ylab("Proportion (%)") +
    ggtitle(paste("ZONE:", zone_i)) +
    scale_y_continuous(labels = percent_format(accuracy = .1)) +  # <- Percent axis
    #scale_y_continuous()+
    theme_minimal(base_size = 10) +
    theme(
      strip.text = element_text(size = 6, face = "bold"),
      axis.text.x = element_text(angle = 90, hjust = 1),
      axis.text.y = element_text(angle = 0)
    )
  
  print(p)
}
```

 
Length by sex, year and month for each Zone
```{r}
WORDAC_1 %>%
filter(SEX %in% c("F", "M")) %>%
  ggplot(aes(x = LENGTH_CM, fill = SEX)) +
   geom_histogram(binwidth = 0.5, position = "identity", alpha = 0.5) +
   geom_density(aes(y = ..count.., color = SEX), size = 0.3, adjust = 0.5, fill = NA) +
   xlab("Standard Length (cm)") +
  ylab("Count") +
   facet_grid(ZONE + year ~ SEX + month, scales = "free_y") +
   scale_fill_brewer(palette = "Set1") +
   scale_color_brewer(palette = "Set1") +
   labs(fill = "Sex", color = "Sex") +
   theme_minimal(base_size = 11) +
   theme(
   strip.text = element_text(size = 8, face = "bold"),
   axis.text.x = element_text(angle = 0, vjust = 0.5),
)
```


Length by sex and year
```{r}
ggplot(
  data = WORDAC_1 %>% filter(SEX %in% c("F", "M")),
  aes(x = LENGTH_CM, fill = SEX)) +
  geom_histogram(binwidth = 0.5, position = "identity", alpha = 0.5) +
  geom_density(aes(y = ..count.., color = SEX), size = .1, adjust = .5, fill = NA) +
  xlab("Standard Length (cm) by") +
  ylab("Count") +
  facet_grid(rows = vars(year), cols = vars(SEX), scales = "free_y")+
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1") +
  labs(fill = "Sex", color = "Sex") +
  theme_minimal()
```

Length by sex, year and month
```{r}
ggplot(
  data = WORDAC_1 %>% filter(SEX %in% c("F", "M")),
  aes(x = LENGTH_CM, fill = SEX)
) +
  geom_histogram(binwidth = 0.5, position = "identity", alpha = 0.5) +
  geom_density(aes(y = ..count.., color = SEX), size = .1, adjust = .5, fill = NA) +
  xlab("Standard Length (cm)") +
  ylab("Count") +
  facet_grid(year ~ SEX + month, scales = "free_y") +  # facet by year rows and sex+month columns
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1") +
  labs(fill = "Sex", color = "Sex") +
  theme_minimal()
```

```{r}
LF_Tally_Sex <- WORDAC_1 %>%
  #filter(Sex =="F" |Sex =="M") %>%
  group_by(SEX) %>%
  add_tally() %>%
  group_by(SEX, LENGTH_CM) %>%
  dplyr::summarise(prop = n() / n[1])


LF_sex <- ggplot(LF_Tally_Sex, aes(LENGTH_CM, prop, fill = SEX)) +
  geom_col(position = "identity", alpha = 0.6) +
  scale_x_continuous(breaks = seq(0,65,10),limits = c(0,65))+
  scale_y_continuous(breaks = seq(0,0.1,0.1),limits = c(0,0.1),
sec.axis = dup_axis(name = NULL))+
  facet_wrap(~ SEX, ncol = 1) +  # <--- one plot per Sex
   theme_bw(base_size = 8)+
  ylab("proportion")+
  xlab("Length (cm)")+
  theme(legend.position = "bottom",
  strip.text = element_text(size = 7, face = "plain"),  # <-- Thinner headings
  axis.title = element_text(size = 8),
  axis.text = element_text(size = 7),
  legend.text = element_text(size = 7),
  legend.title = element_text(size = 7),
  legend.key.height = unit(0.3, "cm"),  # Optional: reduce legend key height
  legend.key.width = unit(0.3, "cm")    # Optional: reduce legend key width
  )
# # 
plot(LF_sex)
# # ggsave(filename = "Data/Roughy_SESSF/LF_sex.emf",device = emf)
# # #ggsave(filename = "Data/Roughy_SESSF/LF_RoughyDL.jpg")
```


## Length frequencies by zones

ANOVA and Tukey test
 - Southern fish are MUCH smaller than BOTH Northern and Central fish 
    *(-6.4cm and -5.4 cm respectively)
 - Northern fish are larger than Central fish 
    *(~1cm)
 - In Southern Zone M and F fish are not significantly different 
      *(Southern:M-Southern:F -0.7583570 -1.6015240  0.0848100 0.1064615)

Need to break it down by years within each ZONE to see cohorts

combine the null and 0 zones
```{r}
WORDAC_1 <- WORDAC_1 %>%
  mutate(ZONE = ifelse(ZONE == 0, NA, ZONE)) %>% 
  mutate(ZONE = ifelse(ZONE == "null", NA, ZONE)) # didn't work

WORDAC_1 <- WORDAC_1 %>% mutate(ZONE = ifelse(ZONE == "(null)", NA, ZONE)) # worked


unique(WORDAC_1$ZONE)


```

```{r}
LF_Tally <- WORDAC_1 %>%
   filter(Sex =="F" |Sex =="M") %>%
   group_by(ZONE,Sex) %>%
   add_tally() %>%
group_by(ZONE,Sex,Lgt_cm) %>%
   dplyr::summarise(prop = n() / n[1])

 LF <- ggplot(LF_Tally, aes(Lgt_cm, prop, fill = Sex)) +
   geom_col(position = "identity", alpha = 0.6) +
   scale_x_continuous(breaks = seq(0,65,10),limits = c(0,65))+
   scale_y_continuous(breaks = seq(0,0.15,0.1),limits = c(0,0.15),
 sec.axis = dup_axis(name = NULL))+
   facet_wrap(~ paste(ZONE), ncol = 4)+
    theme_bw(base_size = 8)+
   ylab("proportion")+
   xlab("Length (cm)")+
   theme(legend.position = "bottom",
   strip.text = element_text(size = 7, face = "plain"),  # <-- Thinner headings
   axis.title = element_text(size = 8),
   axis.text = element_text(size = 7),
   legend.text = element_text(size = 7),
   legend.title = element_text(size = 7),
   legend.key.height = unit(0.3, "cm"),  # Optional: reduce legend key height
   legend.key.width = unit(0.3, "cm")    # Optional: reduce legend key width
   )

 plot(LF)
 ggsave(filename = "Data/Roughy_SESSF/LF_RouhgyDL.emf",device = emf)
 ggsave(filename = "Data/Roughy_SESSF/LF_RoughyDL.jpg")
```


average SL by ZONE 
```{r}
WORDAC_1 %>%
  dplyr::group_by(SEX,ZONE) %>%
  dplyr::summarise(meanSL = mean(LENGTH_CM, na.rm = TRUE))

```

anova
```{r}
Anova_Zone <- aov(LENGTH_CM ~ ZONE * SEX, data = WORDAC_1 %>% filter(SEX %in% c("F", "M")))
summary(Anova_Zone)
TukeyHSD(Anova_Zone)
```

LF by Zone and sex 
```{r}
ggplot(WORDAC_1, aes(x = LENGTH_CM, fill = SEX)) +
  geom_histogram(binwidth = 1) +   # adjust binwidth as needed
  xlab("length (cm)") +
  ylab("Count") +
  facet_wrap(~ZONE)+
   theme_minimal()
```


## Average Standard Length for each zone by year

```{r}
WORDAC_1 %>%
  filter(SEX %in% c("F", "M")) %>%
  dplyr::group_by(ZONE,year, SEX) %>%
  dplyr::summarise(mean_SL = mean(LENGTH_CM, na.rm = TRUE),
            se = sd(LENGTH_CM, na.rm = TRUE) / sqrt(n())) %>%
  ggplot(aes(x = factor(year), y = mean_SL, color = SEX, group = SEX)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_SL - se, ymax = mean_SL + se), width = 0.2) +
  xlab("Year") + ylab("Mean Standard Length (cm)") +
  theme_minimal() +
  labs(title = "Yearly Standard Length by Sex")+
  facet_wrap(~ZONE)

WORDAC_1 %>%
  filter(SEX %in% c("F", "M")) %>%
  dplyr::group_by(ZONE, year, SEX) %>%
  dplyr::summarise(
    mean_SL = mean(LENGTH_CM, na.rm = TRUE),
    se = sd(LENGTH_CM, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = factor(year), y = mean_SL, color = SEX, group = SEX)) +
  geom_line(size = 1) +
  geom_point(size = 2.5) +
  geom_errorbar(
    aes(ymin = mean_SL - se, ymax = mean_SL + se),
    width = 0.2,
    size = 0.4
  ) +
  facet_wrap(~ ZONE, scales = "free_y") +
  scale_color_brewer(palette = "Set1") +
  xlab("Year") +
  ylab("Mean Standard Length (cm)") +
  labs(
    title = "Yearly Mean Standard Length by Sex and Zone",
    color = "Sex"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5)
  )
```


## Length frequencies for each ZONE

### Northern
- Histogram - two peaks obvious for F in years with high sampling
      *2024,2022, 2021,2020 - note that for 2021 there could be ~4 peaks?
      * not as obvious for M fish
      

```{r}
# Northern <- WORDAC_1 %>% 
#   filter(ZONE == "Northern")

ggplot(
  data = WORDAC_1 %>% filter(SEX %in% c("F", "M"), ZONE == "Northern"),
  aes(x = LENGTH_CM, fill = SEX)) +
  geom_histogram(binwidth = 0.5, position = "identity", alpha = 0.5) +
  geom_density(aes(y = ..count.., color = SEX), size = .1, adjust = .5, fill = NA) +
  xlab("Northern Zone Standard Length (cm)") +
  ylab("Count") +
  facet_grid(rows = vars(year), cols = vars(SEX), scales = "free_y")+
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1") +
  labs(fill = "Sex", color = "Sex") +
  theme_minimal()
```


### Central
- Histogram - up to 4 peaks 

```{r}
ggplot(
  data = WORDAC_1 %>% filter(SEX %in% c("F", "M"), ZONE == "Central"),
  aes(x = LENGTH_CM, fill = SEX)) +
  geom_histogram(binwidth = 0.5, position = "identity", alpha = 0.5) +
  geom_density(aes(y = ..count.., color = SEX), size = .1, adjust = .5, fill = NA) +
  xlab("Central Zone Standard Length (cm)") +
  ylab("Count") +
   facet_grid(rows = vars(year), cols = vars(SEX), scales = "free_y")+
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1") +
  labs(fill = "Sex", color = "Sex") +
  theme_minimal()
```


### Southern


```{r}
ggplot(
  data = WORDAC_1 %>% filter(SEX %in% c("F", "M"), ZONE == "Southern"),
  aes(x = LENGTH_CM, fill = SEX)) +
  geom_histogram(binwidth = 0.5, position = "identity", alpha = 0.5) +
  geom_density(aes(y = ..count.., color = SEX), size = .1, adjust = .5, fill = NA) +
  xlab("Southern Zone Standard Length (cm)") +
  ylab("Count") +
  facet_grid(rows = vars(year), cols = vars(SEX), scales = "free_y")+
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1") +
  labs(fill = "Sex", color = "Sex") +
  theme_minimal()
```
## Does looking by month make any difference?

### Northern
```{r}
ggplot(
  data = WORDAC_1 %>% filter(SEX %in% c("F", "M"), ZONE == "Northern"),
  aes(x = LENGTH_CM, fill = SEX)) +
  geom_histogram(binwidth = 0.5, position = "identity", alpha = 0.5) +
  geom_density(aes(y = ..count.., color = SEX), size = .1, adjust = .5, fill = NA) +
  xlab("Northern Zone Standard Length (cm)") +
  ylab("Count") +
  facet_grid(rows = vars(MonthName), cols = vars(year), scales = "free_y")+
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1") +
  labs(fill = "Sex", color = "Sex") +
  theme_minimal()
```

### Central
```{r}
ggplot(
  data = WORDAC_1 %>% filter(SEX %in% c("F", "M"), ZONE == "Central"),
  aes(x = LENGTH_CM, fill = SEX)) +
  geom_histogram(binwidth = 0.5, position = "identity", alpha = 0.5) +
  geom_density(aes(y = ..count.., color = SEX), size = .1, adjust = .5, fill = NA) +
  xlab("Central Zone Standard Length (cm)") +
  ylab("Count") +
  facet_grid(rows = vars(MonthName), cols = vars(year), scales = "free_y")+
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1") +
  labs(fill = "Sex", color = "Sex") +
  theme_minimal()
```


### Southern
```{r}
ggplot(
  data = WORDAC_1 %>% filter(SEX %in% c("F", "M"), ZONE == "Southern"),
  aes(x = LENGTH_CM, fill = SEX)) +
  geom_histogram(binwidth = 0.5, position = "identity", alpha = 0.5) +
  #geom_density(aes(y = ..count.., color = SEX), size = .1, adjust = .5, fill = NA) +
  xlab("Southern Zone Standard Length (cm)") +
  ylab("Count") +
  facet_grid(rows = vars(MonthName), cols = vars(year), scales = "free_y")+
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1") +
  labs(fill = "Sex", color = "Sex") +
  theme_minimal()
```



exploring Stats looking -  month differences
```{r}
WORDAC_1$year <- as.factor(WORDAC_1$year)
WORDAC_1$ZONE <- as.factor(WORDAC_1$ZONE)
Anova_Zone <- aov(LENGTH_CM ~ year * MonthName * SEX, data = WORDAC_1 %>% filter(SEX %in% c("F", "M"),ZONE == "Northern"))
summary(Anova_Zone)
TukeyHSD(Anova_Zone)

```


stats with chatgpt
```{r}
results <- WORDAC_1 %>%
  filter(SEX %in% c("F", "M")) %>%
  dplyr::group_by(ZONE, year) %>%
  dplyr::group_modify(~ {
    data_sub <- .
    
    # Drop rows with missing values
    data_sub <- data_sub %>%
      filter(!is.na(LENGTH_CM), !is.na(MonthName), !is.na(SEX))
    
    # Drop unused factor levels
    data_sub <- droplevels(data_sub)
    
    # Defensive: If no data left, skip
    if (nrow(data_sub) == 0) {
      message("No data: ", unique(.$ZONE), " ", unique(.$year))
      return(tibble(term = "MonthName", statistic = NA_real_, p.value = NA_real_,
                    ZONE = unique(.$ZONE), year = unique(.$year)))
    }

    # Ensure factors are truly re-factored
    data_sub$MonthName <- droplevels(as.factor(data_sub$MonthName))
    data_sub$SEX <- droplevels(as.factor(data_sub$SEX))

    if (nlevels(data_sub$MonthName) < 2) {
      message("MonthName has < 2 levels: ", unique(data_sub$ZONE), " ", unique(data_sub$year))
      return(tibble(term = "MonthName", statistic = NA_real_, p.value = NA_real_,
                    ZONE = unique(data_sub$ZONE), year = unique(data_sub$year)))
    }

    if (nlevels(data_sub$SEX) < 2) {
      message("SEX has < 2 levels: ", unique(data_sub$ZONE), " ", unique(data_sub$year))
      return(tibble(term = "MonthName", statistic = NA_real_, p.value = NA_real_,
                    ZONE = unique(data_sub$ZONE), year = unique(data_sub$year)))
    }

    # Fit the model safely
    fit <- tryCatch(
      aov(LENGTH_CM ~ MonthName + SEX, data = data_sub),
      error = function(e) {
        message("Model failed for ", unique(data_sub$ZONE), " ", unique(data_sub$year), ": ", e$message)
        return(NULL)
      }
    )

    if (is.null(fit)) {
      return(tibble(term = "MonthName", statistic = NA_real_, p.value = NA_real_,
                    ZONE = unique(data_sub$ZONE), year = unique(data_sub$year)))
    }

    # Extract result
    broom::tidy(fit) %>%
      filter(term == "MonthName") %>%
      mutate(ZONE = unique(data_sub$ZONE), year = unique(data_sub$year))
  }) %>%
  ungroup() %>%
  select(ZONE, year, term, statistic, p.value)

print(results)

```
