---
title: "2019 Orange Roughy Eastern Zone Spawaning Survey Figures and Tables"
author: "Caroline Sutton"
date: "03/12/2019"
output:
  html_document:
    code_folding: hide

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE,
  warning = FALSE,
  message = FALSE,
  results = 'asis',
  fig.align = 'center')

#Load libraries
library(tidyverse)
library(ggplot2)
library(scales)
library(ggpubr)
library(Rmisc)
library(dplyr)
library(devEMF)
library(tidyr)
library(kableExtra)
library(knitr)


#read in tidy data from results folder
Roughy <- read_csv("Results/OREastern_Tidy_87to2019.csv", col_types = list(ShotWeight_kg = col_double(), mark = col_character()))


#remame and select columns
Roughy1 <- Roughy %>% 
  mutate(Length =round(SLadj_cm)) %>% #round values for length in cm 2019 was to nearest mm
  mutate(Ground = Area_MS) %>% 
  drop_na(Ground) %>% 
  drop_na(Length) %>% 
  drop_na(sex) %>% 
  select(Ground, year, Shot, Date,sex,stage, Length,ShotWeight_kg)
```
# Introduction
Summary Data, Tables and Figures for Eastern Zone Roughy Survey 2019


## Table and Figures

**Tables**
---


```{r mytable, out.width='100%', echo = T}
Data_summary <- Roughy1 %>% 
  dplyr::group_by(year, Ground,sex) %>% 
  dplyr::summarise(n= n())

Data_summary <- spread(Data_summary, key = year,value = n)


knitr::kable(tail(Data_summary, n = 5), format = "html", caption = "Summary data: Number of Fish measured for orange roughy eastern zone spawning surveys by, year, Area, sex") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width = "100%")

```

## Table 2.3
Summary of shot dates and weights for St Helens and St Patricks, Eastern Zone, 2019 (* no fish collected from these shots. Note that bycatch data were not reported

```{r table2.3, out.width='100%', echo = T} 
OR2019 <- Roughy1 %>% 
  filter(year == 2019)
OR2019_sum <- OR2019 %>% 
  dplyr::group_by(Ground,Shot,Date, 'Shot Weight (kg)' = ShotWeight_kg,sex) %>% 
  dplyr::summarise(num_rows = n()) %>% 
  spread(sex,num_rows) %>% 
  mutate(`number of fish` = F + M)



knitr::kable(head(OR2019_sum, n=14), format = "html", caption = "Table 2.3 Summary of shot dates and weights for St Helens and St Patricks, Eastern Zone, 2019") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width  = "100%")
```

## Figure 2.1 
Average Standard Length (cm) of orange roughy Eastern Zone spawning population, showing males and females at St Helens and St Patricks for 1987 to 2019. Error bars are 95% confidence limits (updated from Kloser et al. 2016)
```{r myfig2.1, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
#average SL by Area and sex

roughy2_sum <- summarySE(Roughy1, measurevar = "Length", groupvars = c("year","Ground","sex")) 

  
pd <- position_dodge(0.6) # move position horizontally b/c of overlap

f2.1 <- ggplot(roughy2_sum, aes(x = year,y = Length, shape = Ground, colour = sex)) +
  geom_errorbar(aes(ymin = Length - ci, ymax= Length + ci),
                width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_line(position = pd) +
  xlab("Eastern Zone Spawning Population by year for July")+
  ylab("Average Standard Length (cm)")+
  scale_y_continuous(breaks = seq(26,38,1),limits = c(26,38))+
  scale_x_continuous(limits = c(1986,2020), breaks = seq(1986,2020,2))+
  theme_bw()+
  theme(legend.position = c(.95, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'))
plot(f2.1)
ggsave("Figures/f2.1.emf")

```

## Figure 2.2
Standard Length by sex and ground
```{r myfig2.2}
OR_LF <- OR2019 %>% # This is the ONE
  group_by(Ground, sex) %>% 
  add_tally() %>% 
  group_by(Ground, sex, Length) %>% 
  dplyr::summarise(prop = n() / n[1])

f2.2 <- ggplot(OR_LF, aes(Length, prop, fill = sex)) +
  geom_col(position = "identity", alpha = 0.6) +
  scale_x_continuous(breaks = seq(28,46,2),limits = c(28,46))+
  facet_wrap(~Ground)+
   theme_bw()+
  ylab("proportion")+
  xlab("Length (cm)")+
  theme(legend.position = "bottom")
plot(f2.2)
ggsave(filename = "Figures/f2.2.emf")
```


## Table 2.4 
Mode and Numbers of fish by Year, Ground, Sex
```{r Table2.4}
Roughy1 <- Roughy1 %>% 
  mutate(GbyS= paste(Ground,sex,sep = "_")) %>%  #new variable to combine Ground and sex 
  group_by(GbyS) %>% 
  mutate(sumsum = cumsum(Length))


#create function to find the mode
Mode <- function(x) {
    if (is.numeric(x)) {
        x_table <- table(x)
        return(as.numeric(names(x_table)[which.max(x_table)]))
        }
}
Roughy_mode <- Roughy1 %>% 
  dplyr::group_by(year,GbyS) %>%
  dplyr::summarise(Mode = Mode(x = Length))%>%
  spread(GbyS, Mode) 
#create counts
Roughy_nos <- Roughy1 %>% 
  dplyr::group_by(year, GbyS) %>%
  dplyr::summarise(Numbers = n())%>%
  spread(GbyS, Numbers)

#bind the mode and counts
R_Mode_Count <-  cbind(Roughy_mode,Roughy_nos)%>%
  select(-6)#remove the year1 colomn

  R_Mode_Count_order <- R_Mode_Count[,c(1,2,6,3,7,4,8,5,9)]  # order the columns
 
knitr::kable(R_Mode_Count_order, col.names = c("", "Nos. Fish","Mode", "Nos. Fish","Mode", "Nos. Fish","Mode", "Nos. Fish","Mode"), format = "html", caption = "Number of Fish measured and mode of standard lengths (cm) for male and femailes measured at St Helens and St Particks from 1987 to 2019") %>% 
  kableExtra::kable_styling("striped") %>%
  add_header_above(c(" ", "Female" = 2, "Male" = 2, "Female" = 2, "Male"=2)) %>%
    add_header_above(c(" ", "St Helens" = 4, "St Patricks" = 4)) %>%
    kableExtra::scroll_box(width = "100%")
```
  
## Figure 2.4
Cummunlative Frequency
```{r myfig2.3}

#using stat_ecdf to calculate the cummulative frequency

f2.3 <- ggplot(OR2019,aes(x=Length, colour= sex)) +
   stat_ecdf(geom = "smooth")+
  scale_x_continuous(breaks = seq(28,44,2),limits = c(28,44))+
  facet_wrap(~Ground)+
  ylab("cummulative frequency")+
  xlab("Length (cm)")+
  theme_bw()+
  theme(legend.position = "bottom")
plot(f2.3)
ggsave(filename = "Figures/f2.3.emf")
  
```

## Figure 2.4
Average Standard Length (cm) by shot for males and females for 2019
```{r myfig2.4, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
#average SL by Area and sex

OR2019_sum <- summarySE(OR2019, measurevar = "Length", groupvars = c("Shot","Date","Ground","sex")) %>% 
  mutate(Shot = as.character(Shot))

  
pd <- position_dodge(0) # move position horizontally b/c of overlap

f2.4 <- ggplot(OR2019_sum, aes(x = reorder(Shot, sort(as.numeric(Shot))),y = Length, colour = sex)) +
  geom_errorbar(aes(ymin = Length - ci, ymax= Length + ci),
                width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
 # geom_line(position = pd) +
  xlab("Shot Number")+
  ylab("Average Standard Length (cm)")+
  scale_y_continuous(breaks = seq(30,38,1),limits = c(30,38))+
  theme_bw()+
  theme(legend.position = c(.95, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'))+
  facet_grid(.~Ground +Date, scales = "free_x") # group by Ground&Date, allows only the shots relevant for ea ground
plot(f2.4)
ggsave("Figures/f2.4.emf")
```

## Figure 2.5
```{r myfig2.5}

OR2019_1 <- OR2019 %>%
  mutate(stage = as.character(stage))%>% 
  mutate(Shot = as.character(Shot)) %>% 
  group_by(sex)
  
  
f2.5 <- OR2019_1 %>% 
  ggplot(
    mapping = aes(x = Shot,
                  fill = stage))+
  geom_bar(position = "fill")+
  scale_fill_manual(values = c("yellow","orange","red","black"))+
  facet_grid(.~Ground + Date, scales = "free_x")

  
plot(f2.5)

```


# My Data School Experience

Data School has been a very useful and rewarding experience and has changed the way I do my job. I can now manage and document my data wrangling which was previously recorded in notebooks and reports.  This has been hugely transformative as much of my job involves collating data from different researchers and projects with varied field and data management practices. I am also very grateful for the encouragement, patience and kindness of the Data School Focus team who made the learning process fun and manageable.
