---
title: "2019 Orange Roughy Eastern Zone Spawaning Survey Figures and Tables"
author: "Caroline Sutton"
date: "03/12/2019"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE,
  warning = FALSE,
  message = FALSE,
  results = 'asis',
  fig.align = 'center')
#load libraries
library(tidyverse)
library(ggplot2)
library(scales)
library(ggpubr)
library(Rmisc)
library(dplyr)
library(devEMF)
library(tidyr)
<<<<<<< HEAD
library(kableExtra)
library(knitr)
=======
>>>>>>> 6b95d4186e1a430b6431f128e57f7869d5d3b5fa


#read in tidy data from results folder
Roughy <- read_csv("Results/OREastern_Tidy_87to2019.csv", col_types = list(ShotWeight_kg = col_double(), mark = col_character()))


#remame and select columns
Roughy1 <- Roughy %>% 
  mutate(Length =round(SLadj_cm)) %>% #round values for length in cm 2019 was to nearest mm
  mutate(Ground = Area_MS) %>% 
  drop_na(Ground) %>% 
  drop_na(Length) %>% 
  drop_na(sex) %>% 
  select(Ground, year, Shot, Date,sex, Length,ShotWeight_kg)
```

## Summary data: 1987 to 2019 

```{r mytable, out.width='100%', echo = T}
Data_summary <- Roughy1 %>% 
  dplyr::group_by(year, Ground,sex) %>% 
  dplyr::summarise(n= n())

Data_summary <- spread(Data_summary, key = year,value = n)


knitr::kable(tail(Data_summary, n = 5), format = "html", caption = "Summary data: Orange Roughy Eastern Zone Spawning Survey Data by, year, Area, sex") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width = "100%")
```

## Figures: Biological data 

### Table 2.3
Summary of shot dates and weights for St Helens and St Patricks, Eastern Zone, 2013 (* no fish collected from these shots. Note that bycatch data is not available

```{r table2.3, out.width='100%', echo = T} 
OR2019 <- Roughy1 %>% 
  filter(year == 2019)
OR2019_sum <- OR2019 %>% 
  dplyr::group_by(Ground,Shot,Date, 'Shot Weight (kg)' = ShotWeight_kg,sex) %>% 
  dplyr::summarise(num_rows = n()) %>% 
  spread(sex,num_rows) %>% 
  mutate(`number of fish` = F + M)



knitr::kable(head(OR2019_sum, n=14), format = "html", caption = "Table 2.3 Summary of shot dates and weights for St Helens and St Patricks, Eastern Zone, 2019") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width  = "100%")

```
### Figure 2.1 
Average Standard Length (cm) of orange roughy Eastern Zone spawning population, showing males and females at St Helens and St Patricks for 1987 to 2016. Error bars are 95% confidence limits (updated from Kloser et al. 2016)
```{r myfig1, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
#average SL by Area and sex

roughy2_sum <- summarySE(Roughy1, measurevar = "Length", groupvars = c("year","Ground","sex"))

pd <- position_dodge(0.6) # move position horizontally b/c of overlap
Avg_SL <- ggplot(roughy2_sum, aes(x = year,y = Length, shape = Ground, colour = sex)) +
  geom_errorbar(aes(ymin = Length - ci, ymax= Length + ci),
                width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_line(position = pd) +
  xlab("Year, Eastern Zone, July spawning population")+
  ylab("Average Standard Length (cm)")+
  scale_y_continuous(breaks = seq(26,38,1),limits = c(26,38))+
  scale_x_continuous(limits = c(1986,2020), breaks = seq(1986,2020,2))+
  theme_bw()+
  theme(legend.position = c(.95, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'))
plot(Avg_SL)
ggsave("Figures/AVg_SL.emf")

```
#### Figure 2.2
```{r myfig}
OR2019 %>% # This is the ONE
  group_by(Ground, sex) %>% 
  add_tally() %>% 
  group_by(Ground, sex, Length) %>% 
  dplyr::summarise(prop = n() / n[1]) %>% 
  ggplot(aes(Length, prop, fill = sex)) +
  geom_col(position = "identity", alpha = 0.6) +
  scale_x_continuous(breaks = seq(28,46,2),limits = c(28,46))+
  facet_wrap(~Ground)+
   theme_bw()+
  ylab("proportion")+
  xlab("Length (cm)")+
  theme(legend.position = "bottom")
```
### Figure 2.2
Length frequencies (cm) for the 2019 orange roughy Eastern Zone spawning population for males and females  at St Helens and St Patricks.  Data presented as total counts of females (F) and males (M) within each spawning ground.  Samples numbers are as follows: St Helens F = 856, St Helens M = 438, St Patricks F = 526 and St Patricks M = 689. Vertical line showing the median for all data.


```{r myfig2}

OR2019 %>% # This is the ONE
  group_by(Ground, sex) %>% 
  add_tally() %>% 
  group_by(Ground, sex, Length) %>% 
  dplyr::summarise(prop = n() / n[1]) %>% 
  ggplot(aes(Length, prop, fill = sex)) +
  geom_col(position = "identity", alpha = 0.6) +
  scale_x_continuous(breaks = seq(28,46,2),limits = c(28,46))+
  facet_wrap(~Ground)+
   theme_bw()+
  ylab("proportion")+
  xlab("Length (cm)")+
  theme(legend.position = "bottom")


```
### Figure2.3 Cummulative frequency

```{r}
OR2019 <- OR2019 %>% 
  mutate(GbyS= paste(Ground,sex,sep = "_")) %>%  #new variable to combine Ground and sex 
  group_by(GbyS) %>% 
  mutate(sumsum = cumsum(Length))


#using stat_ecdf to calculate the cummulative frequency

ggplot(OR2019,aes(x=Length, colour= sex)) +
   stat_ecdf(geom = "smooth")+
  scale_x_continuous(breaks = seq(28,44,2),limits = c(28,44))+
  facet_wrap(~Ground)+
  ylab("cummulative frequency")+
  xlab("Length (cm)")+
  theme_bw()+
  theme(legend.position = "bottom")
  
```
### Table 2 4 summarise all data by counts and mode  
Number of fish measured and mode for standard lengths for males and females measured at St Helens and St Patricks from 1987 to 2013, updated from Kloser et al. 2012. 
<<<<<<< HEAD
```{r mytable2.4, out.width='100%', echo = T}
Roughy1 <- Roughy1 %>% #combine Ground and Sex columns 
  mutate(GbyS = paste(Ground,sex, sep = "_"))
  
#create modes
=======
```{r table 4}

#combine Ground and Sex columns 
Roughy1 <- Roughy1 %>% 
  mutate(GbyS = paste(Ground,sex, sep = "_")) %>% 
  view(Roughy1)

#create modes
Roughy_mode <- Roughy1 %>% 
  dplyr::group_by(year,GbyS) %>%
  dplyr::summarise(Mode = Mode(x = Length))%>%
  spread(GbyS, Mode) %>% 
  view(Roughy_mode)
#create counts
Roughy_nos <- Roughy1 %>% 
  dplyr::group_by(year, GbyS) %>%
  dplyr::summarise(Numbers = n())%>%
  spread(GbyS, Numbers) %>%
  view(Roughy_nos)

#bind the mode and counts
R_Mode_Count <-  cbind(Roughy_mode,Roughy_nos)%>%
  select(-6)  #remove the year1 colomn
  R_Mode_Count_order <- R_Mode_Count[,c(1,2,6,3,7,4,8,5,9)] %>% # order the columns
  view(R_Mode_Count_order)
 

knitr::kable(tail(R_Mode_Count, n = 17), format = "html", caption = "Number of Fish measured and mode of standard lengths (cm) for male and femailes measured at St Helens and St Particks from 1987 to 2019") %>% 
  kableExtra::kable_styling("striped") %>%
    add_header_above(c(" ", "Female" = 2, "Male" = 2, "Female" = 2, "Male"=2)) %>%
    add_header_above(c(" ", "St Helens" = 4, "St Patricks" = 4)) %>%
    kableExtra::scroll_box(width = "100%")

```


Number of fish measured and mode for standard lengths for males and females measured at St Helens and St Patricks from 1987 to 2013, updated from Kloser et al. 2012. 
```{r table}
>>>>>>> 6b95d4186e1a430b6431f128e57f7869d5d3b5fa

#create function to find the mode
Mode <- function(x) {
    if (is.numeric(x)) {
        x_table <- table(x)
        return(as.numeric(names(x_table)[which.max(x_table)]))
        }
}
Roughy_mode <- Roughy1 %>% 
  dplyr::group_by(year,GbyS) %>%
  dplyr::summarise(Mode = Mode(x = Length))%>%
  spread(GbyS, Mode) 
#create counts
Roughy_nos <- Roughy1 %>% 
  dplyr::group_by(year, GbyS) %>%
  dplyr::summarise(Numbers = n())%>%
  spread(GbyS, Numbers)

#bind the mode and counts
R_Mode_Count <-  cbind(Roughy_mode,Roughy_nos)%>%
  select(-6)#remove the year1 colomn

  R_Mode_Count_order <- R_Mode_Count[,c(1,2,6,3,7,4,8,5,9)]  # order the columns
 
knitr::kable(R_Mode_Count_order, col.names = c("", "Nos. Fish","Mode", "Nos. Fish","Mode", "Nos. Fish","Mode", "Nos. Fish","Mode"), format = "html", caption = "Number of Fish measured and mode of standard lengths (cm) for male and femailes measured at St Helens and St Particks from 1987 to 2019") %>% 
  kableExtra::kable_styling("striped") %>%
  add_header_above(c(" ", "Female" = 2, "Male" = 2, "Female" = 2, "Male"=2)) %>%
    add_header_above(c(" ", "St Helens" = 4, "St Patricks" = 4)) %>%
    kableExtra::scroll_box(width = "100%")

```



