---
title: "2019 Orange Roughy Eastern Zone Spawaning Survey Figures and Tables"
author: "Caroline Sutton"
date: "03/12/2019"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE,
  warning = FALSE,
  message = FALSE,
  results = 'asis',
  fig.align = 'center')
#load libraries
library(tidyverse)
library(ggplot2)
library(scales)
library(ggpubr)
library(Rmisc)
library(dplyr)
library(devEMF)


#read in tidy data from results folder
Roughy <- read_csv("Results/OREastern_Tidy_87to2019.csv", col_types = list(ShotWeight_kg = col_double(), mark = col_character()))


#remame and select columns
Roughy1 <- Roughy %>% 
  mutate(Length = SLadj_cm) %>% 
  mutate(Ground = Area_MS) %>% 
  drop_na(Ground) %>% 
  drop_na(Length) %>% 
  drop_na(sex) %>% 
  select(Ground, year, Shot, Date,sex, Length,ShotWeight_kg)
```

## Summary data: 1987 to 2019 

```{r mytable, out.width='100%', echo = T}
Data_summary <- Roughy1 %>% 
  dplyr::group_by(year, Ground,sex) %>% 
  dplyr::summarise(n= n())

Data_summary <- spread(Data_summary, key = year,value = n)


knitr::kable(tail(Data_summary, n = 5), format = "html", caption = "Summary data: Orange Roughy Eastern Zone Spawning Survey Data by, year, Area, sex") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width = "100%")
```

## Figures: Biological data 

#### Table 2.3
Summary of shot dates and weights for St Helens and St Patricks, Eastern Zone, 2013 (* no fish collected from these shots. Note that bycatch data is not available

```{r table, out.width='100%', echo = T} 
OR2019 <- Roughy1 %>% 
  filter(year == 2019)
OR2019_sum <- OR2019 %>% 
  dplyr::group_by(Ground,Shot,Date, ShotWeight_kg,sex) %>% 
  dplyr::summarise(num_rows = n()) %>% 
  spread(sex,num_rows) %>% 
  mutate(`total measured` = F + M)



knitr::kable(head(OR2019_sum, n=14), format = "html", caption = "Table 2.3 Summary of shot dates and weights for St Helens and St Patricks, Eastern Zone, 2019") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width  = "100%")

```
#### Figure 2.1 
Average Standard Length (cm) of orange roughy Eastern Zone spawning population, showing males and females at St Helens and St Patricks for 1987 to 2016. Error bars are 95% confidence limits (updated from Kloser et al. 2016)
```{r myfig1, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}

#average SL by Area and sex

roughy2_sum <- summarySE(Roughy1, measurevar = "Length", groupvars = c("year","Ground","sex"))

pd <- position_dodge(0.6) # move position horizontally b/c of overlap
Avg_SL <- ggplot(roughy2_sum, aes(x = year,y = Length, shape = Ground, colour = sex)) +
  geom_errorbar(aes(ymin = Length - ci, ymax= Length + ci),
                width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_line(position = pd) +
  xlab("Year, Eastern Zone, July spawning population")+
  ylab("Average Standard Length (cm)")+
  scale_y_continuous(breaks = seq(27,38,1),limits = c(27,38))+
  scale_x_continuous(limits = c(1986,2020), breaks = seq(1986,2020,2))+
  theme_bw()+
  theme(legend.position = c(.95, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'))
plot(Avg_SL)
ggsave("Figures/AVg_SL.emf")
```
#### Figure 2.2
```{r myfig}

OR2019 %>% # This is the ONE
  group_by(Ground, sex) %>% 
  add_tally() %>% 
  group_by(Ground, sex, Length) %>% 
  dplyr::summarise(prop = n() / n[1]) %>% 
  ggplot(aes(Length, prop, fill = sex)) +
  geom_col(position = "identity", alpha = 0.6) +
  scale_x_continuous(breaks = seq(28,46,2),limits = c(28,46))+
  facet_wrap(~Ground)+
   theme_bw()+
  ylab("proportion")+
  xlab("Length (cm)")+
  theme(legend.position = "bottom")
```
#### Figure 2.2a
Length frequencies (cm) for the 2019 orange roughy Eastern Zone spawning population for males and females  at St Helens and St Patricks.  Data presented as total counts of females (F) and males (M) within each spawning ground.  Samples numbers are as follows: St Helens F = 856, St Helens M = 438, St Patricks F = 526 and St Patricks M = 689. Vertical line showing the median for all data.

```{r myfig2}
#filter(Roughy, Area_MS == "St Helens") %>% 
  
ggplot(OR2019, aes(Length, fill = sex)) +
  geom_histogram(binwidth=1, position = "identity",alpha =.7)+
  facet_grid(~Ground)+
  
  geom_vline(xintercept = 35.7, size = .01)+ #showing median
  scale_y_continuous()+
  scale_x_continuous(breaks = seq(28,44,2),limits = c(28,44))+
 #scale_y_continuous(limits = c(0,250), breaks = seq(0,250,25))+
  theme_bw()+
  ylab("count")+
  xlab("Length (cm)")
```
#### Figure2.2b
Length frequencies (cm) for the 2019 orange roughy Eastern Zone spawning population for males and females  at St Helens and St Patricks.  Data presented as proporation  contribution to each Length categrory for female fish (F) and male fish (M) at each Ground. Samples numbers are as follows: St Helens F = 856, St Helens M = 438, St Patricks F = 526 and St Patricks M = 689. Vertical line showing the median for all data.
```{r myfig3} 
  
  ggplot(OR2019, aes(x= Length, fill = sex)) +
  geom_histogram(aes(y=c(..count..[..group..==1]/sum(..count..[..group..==1]),
                         ..count..[..group..==2]/sum(..count..[..group..==2]))),
                 position = "dodge",alpha=.7, binwidth = 1)+
    ylab("proportion of total fish for each Ground")+
    xlab("Length (cm)") +
    scale_x_continuous(breaks = seq(28,44,2),limits = c(28,44))+
    facet_grid(.~Ground)+
    theme_bw()+
    theme(legend.position = "bottom")
    #geom_density(alpha = .5)
  
    
```

#### Figure 2.2c Creating LF using percentages
Length frequencies (cm) for the 2019 orange roughy Eastern Zone spawning population for males and females  at St Helens and St Patricks.  Data presented as proporation of number of females (F) and males (M) relative to the total catch within each spawning ground.  Samples numbers are as follows: St Helens F = 856, St Helens M = 438, St Patricks F = 526 and St Patricks M = 689. Vertical line showing the median for all data.
```{r myfig}   
OR2019 <- OR2019 %>% 
  mutate(Length = round(Length)) %>% #round length
  group_by(Ground,sex,Length) %>%  
  dplyr::mutate(total_L = n()) %>% # calculating total number of fish in each length category (by Ground and sex)
  group_by(Ground,sex) %>%  
  dplyr::mutate(total_gs = n())  # calculating the total number of fish for each group (by Ground and sex)
  
  
OR2019_prop <- OR2019 %>% 
  group_by(Ground, sex,Length) %>% 
  dplyr:: summarise(prop = n()/first(total_gs)) # summarise data for prop of total fish in each length categrory (by Gound and sex)

OR2019mean <- OR2019 %>% 
  group_by(Ground,sex) %>% 
  dplyr::summarise(mean = mean(Length))

  ggplot(OR2019_prop,aes(x=Length, y= prop, colour = sex))+ # create the LF plot 
    geom_point(size = 1)+
    geom_smooth()+
    ylab("proportion of fish")+
    xlab("Length (cm)") +
    scale_x_continuous(breaks = seq(28,44,2),limits = c(28,44)) +
    #scale_y_continuous(breaks = seq(0,0.25,.05), limits = c(0,0.25))+
    geom_vline(data=OR2019mean, aes(xintercept=mean,  colour=sex),
             linetype="dashed", size=.5)+
    facet_wrap(~Ground)+
    theme(legend.position = "bottom")
  
    #scale_y_continuous(breaks = seq(0,0.3,.05), limits = c(0,.3))+
    
    
    
    #geom_vline(data=sum2019, aes(xintercept=mean,  colour=sex, group = Groundx),
             #  linetype="dashed", size=1)
    
    
```


#### Figure 2.2d
smoothed length frequency using geom_density which isn't proporation 
```{r myfig4}

ggplot(OR2019,aes(x=Length, colour= sex),group = Ground)+
  geom_density()+
  facet_grid(~Ground)+
  scale_x_continuous(breaks = seq(28,44,2),limits = c(28,44))+
  geom_vline(data=OR2019mean, aes(xintercept=mean,  colour=sex, group = Ground),
               linetype="dashed", size=1)+
  ylab("density")+
  xlab("Length(cm)")+
theme_bw()

```
#### figure2.2explore
checking Alex's method for 2019
```{r}
OR2016 <- filter(Roughy1,year==2016)

ggplot(OR2016,aes(x=Length,fill=sex))+
  geom_histogram(aes(y=stat(density)),binwidth = 1)+
  facet_grid(~Ground)

ggplot(OR2016,aes(x=Length, colour= sex),group = Ground)+
  geom_density()+
  facet_grid(~Ground)

ggplot(OR2016, aes(Length, fill = sex)) +
  geom_histogram(binwidth=1, position = "identity",alpha =.7)+
  facet_grid(~Ground)

OR2016 %>% # This is the ONE
  group_by(Ground, sex) %>% 
  add_tally() %>% 
  group_by(Ground, sex, Length) %>% 
  dplyr::summarise(prop = n() / n[1]) %>% 
  ggplot(aes(Length, prop, fill = sex)) +
  geom_col(position = "identity", alpha = 0.6) +
  facet_wrap(~Ground)
```

### Figure2.3 Cummulative frequency

```{r}
OR2019 <- OR2019 %>% 
  mutate(GbyS= paste(Ground,sex,sep = "_")) %>%  #new variable to combine Ground and sex 
  group_by(GbyS) %>% 
  mutate(sumsum = cumsum(Length))


#using stat_ecdf to calculate the cummulative frequency

ggplot(OR2019,aes(x=Length, colour= sex)) +
   stat_ecdf(geom = "smooth")+
  scale_x_continuous(breaks = seq(28,44,2),limits = c(28,44))+
  facet_wrap(~Ground)+
  ylab("cummulative frequency")+
  xlab("Length (cm)")+
  theme_bw()+
  theme(legend.position = "bottom")
  
```
### Table 2 4 summarise all data by counts and mode  

Number of fish measured and mode for standard lengths for males and females measured at St Helens and St Patricks from 1987 to 2013, updated from Kloser et al. 2012. 
```{r table}

#create function to find the mode
Mode <- function(x) {
    if (is.numeric(x)) {
        x_table <- table(x)
        return(as.numeric(names(x_table)[which.max(x_table)]))
    }
}


# get mode of Length for data by year, Ground and sex  
Dataformode<- Roughy1 %>% 
  mutate(GbyS= paste(Ground,sex,sep = "_")) %>% 
  dplyr::group_by(year, GbyS) %>% #group
  dplyr::summarise(mode = Mode(x= Length), n= n())  # apply Mode function

library(data.table) ## v >= 1.9.6
Table_mode <- dcast(setDT(Dataformode), year ~ GbyS, value.var = c("n", "mode"))


```
