---
title: "Oragne Roughy Eastern Zone Survey Data"
subtitle: "Bioloigical Data from 1987 to 2016"
author: "Caroline Sutton"
affiliation: "CSIRO Oceans and Atmosphere"
photo: resources/img/photo.jpg

short_title: Orange Roughy, *Hoplostethus atlanticus*

output: 
  DSreport::project_summary:
    code_folding: show

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE,
  warning = FALSE,
  message = FALSE,
  results = 'asis',
  fig.align = 'center')

#Load libraries
library(tidyverse)
library(ggplot2)
library(scales)
library(ggpubr)
library(Rmisc)
library(dplyr)
library(devEMF)
library(tidyr)
library(kableExtra)
library(knitr)


#read in tidy data from results folder
Roughy <- read_csv("Results/OREastern_Tidy_87to2019.csv", col_types = list(ShotWeight_kg = col_double(), mark = col_character()))


#remame and select columns
Roughy1 <- Roughy %>% 
  mutate(Length =round(SLadj_cm)) %>% #round values for length in cm 2019 was to nearest mm
  mutate(Ground = Area_MS) %>% 
  drop_na(Ground) %>% 
  drop_na(Length) %>% 
  drop_na(sex) %>% 
  select(Ground, year, Shot, Date,sex, Length,ShotWeight_kg)
```
# Introduction
Summary Data, Tables and Figures for Eastern Zone Roughy Survey 2019


## Table and Figures

**Tables**
---


```{r mytable, out.width='100%', echo = T}
Data_summary <- Roughy1 %>% 
  dplyr::group_by(year, Ground,sex) %>% 
  dplyr::summarise(n= n())

Data_summary <- spread(Data_summary, key = year,value = n)


knitr::kable(tail(Data_summary, n = 5), format = "html", caption = "Summary data: Number of Fish measured for orange roughy eastern zone spawning surveys by, year, Area, sex") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width = "100%")

```

## Table 2.3
Summary of shot dates and weights for St Helens and St Patricks, Eastern Zone, 2019 (* no fish collected from these shots. Note that bycatch data were not reported

```{r table2.3, out.width='100%', echo = T} 
OR2019 <- Roughy1 %>% 
  filter(year == 2019)
OR2019_sum <- OR2019 %>% 
  dplyr::group_by(Ground,Shot,Date, 'Shot Weight (kg)' = ShotWeight_kg,sex) %>% 
  dplyr::summarise(num_rows = n()) %>% 
  spread(sex,num_rows) %>% 
  mutate(`number of fish` = F + M)



knitr::kable(head(OR2019_sum, n=14), format = "html", caption = "Table 2.3 Summary of shot dates and weights for St Helens and St Patricks, Eastern Zone, 2019") %>% 
  kableExtra::kable_styling("striped") %>%
  kableExtra::scroll_box(width  = "100%")
```
--


## Figure 2.1 
Average Standard Length (cm) of orange roughy Eastern Zone spawning population, showing males and females at St Helens and St Patricks for 1987 to 2016. Error bars are 95% confidence limits (updated from Kloser et al. 2016)
```{r myfig2.1, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
#average SL by Area and sex

roughy2_sum <- summarySE(Roughy1, measurevar = "Length", groupvars = c("year","Ground","sex")) 

  
pd <- position_dodge(0.6) # move position horizontally b/c of overlap

Avg_SL <- ggplot(roughy2_sum, aes(x = year,y = Length, shape = Ground, colour = sex)) +
  geom_errorbar(aes(ymin = Length - ci, ymax= Length + ci),
                width =.7, position = pd) +
  geom_point( position = pd, size = 2)+
  geom_line(position = pd) +
  xlab("Year, Eastern Zone, July spawning population")+
  ylab("Average Standard Length (cm)")+
  scale_y_continuous(breaks = seq(26,38,1),limits = c(26,38))+
  scale_x_continuous(limits = c(1986,2020), breaks = seq(1986,2020,2))+
  theme_bw()+
  theme(legend.position = c(.95, .05),
        legend.justification = c(1,0),
        legend.text =element_text(size=8),
        legend.title = element_text(size = 9, margin = margin(t=.05)),
        legend.spacing.y =unit(.05,'cm'))
plot(Avg_SL)
ggsave("Figures/AVg_SL.emf")

```
### Figure 2.2
```{r myfig2.2}
OR2019 %>% # This is the ONE
  group_by(Ground, sex) %>% 
  add_tally() %>% 
  group_by(Ground, sex, Length) %>% 
  dplyr::summarise(prop = n() / n[1]) %>% 
  ggplot(aes(Length, prop, fill = sex)) +
  geom_col(position = "identity", alpha = 0.6) +
  scale_x_continuous(breaks = seq(28,46,2),limits = c(28,46))+
  facet_wrap(~Ground)+
   theme_bw()+
  ylab("proportion")+
  xlab("Length (cm)")+
  theme(legend.position = "bottom")
```

### Figure2.3 Cummulative frequency

```{r}
OR2019 <- OR2019 %>% 
  mutate(GbyS= paste(Ground,sex,sep = "_")) %>%  #new variable to combine Ground and sex 
  group_by(GbyS) %>% 
  mutate(sumsum = cumsum(Length))


#using stat_ecdf to calculate the cummulative frequency

ggplot(OR2019,aes(x=Length, colour= sex)) +
   stat_ecdf(geom = "smooth")+
  scale_x_continuous(breaks = seq(28,44,2),limits = c(28,44))+
  facet_wrap(~Ground)+
  ylab("cummulative frequency")+
  xlab("Length (cm)")+
  theme_bw()+
  theme(legend.position = "bottom")
  
```




# My Data School Experience

Data School has been a very useful and rewarding experience and has changed the way I do my job. I can now manage and document my data wrangling which was previously recorded in notebooks and reports.  This has been hugely transformative as much of my job involves collating data from different researchers and projects with varied field and data management practices. I am also very grateful for the encouragement, patience and kindness of the Data School Focus team who made the learning process fun and manageable.
